### 2022-08-15 manuscript.Rmd
first data push

---
title: Reproducible and Programmatic Generation of Neuroimaging Visualizations
authors:
  - name: Sidhant Chopra
    thanks: Corresponding author
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
    email: sidhant.chopra@yale.com
  - name: Loïc Labache
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Elvisha Dhamala
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Edwina R Orchard
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
  - name: Avram Holmes
    department: Department of Psychology
    affiliation: Yale University
    location: CT, USA, 06511
keywords:
  - Neuroimaging visualization
  - Reproducibility 
  - Programmatic figures
  - Open Science
  - R
  - Python
  - Brain Visualization
abstract: Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results. Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI). There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB. Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process. This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments. Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data.
output: 
  bookdown::pdf_book:
    base_format: rticles::arxiv_article
    keep_tex: true
    latex_engine: lualatex
  bookdown::word_document2:
    toc: false
linestretch: 2
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{setspace}\doublespacing
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Introduction

The visualization of neuroimaging data is one of the primary ways in which we evaluate data quality, interpret results, and communicate findings. These visualizations are commonly produced using graphical user interface-based (GUI) tools where individual images are opened and, within each instance, display settings are manually changed until the desired output is reached. In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility. However, there now exists many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community (Table 1). These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments.

The present article covers three major advantages of using code-generated visualizations over GUI based tools: replicability, flexibility, and integration. First, by generating visualizations using code, you increase the replicability of your figures for yourself, your collaborators and your readers. Second, code-based tools provides more precise controls over display settings, while also benefiting from the advantages of programming, such as the ability to iteratively and rapidly generate multiple figures. Finally, being able to integrate figure generation into your analysis scripts reduces chances of errors and increases the accessibility of resulting analyses pipeline, with more advanced tools now allowing for the seamless integration of prose and code. Here, we review the advantages of learning and using programmatic neuroimaging visualizations, focusing on benefits to replicability, flexibility and integration. We conclude by introducing examples of R-packages which allow for visualization of statistics at region-of-interest (ROI), voxel, vertex, and edge level data, followed by a brief discussion of limitations and functionality gaps in code-based brain visualization tools.

# Replicability

In recent years, there have been multiple large-scale efforts empirically demonstrating the lack of reproducibility of findings from neuroimaging data (Poldrack et al., 2017). One common solution proposed for achieving robust and reliable discoveries has been to encourage scientific output which can be transparently evaluated and independently replicated. In practice, this typically entails openly sharing detailed methods, materials, code, and data. While there is a trend towards increasing transparency and code sharing of neuroimaging analyses, the sharing of code used to generate figures which include brain renderings and spatial maps has been relatively neglected. This gap in reproducibility is partly driven by the fact that brain figures are often created using a manual process that involves tinkering with sliders, buttons, and overlays on a GUI, concluding with a screenshot and sometimes beautification in image processing software like Illustrator or Photoshop. Such a process inherently makes neuroimaging visualizations difficult, if not impossible to replicate, even by the authors themselves.

Visualization scripts should reflect a core feature of open science. Given that brain figures regularly form the centerpiece of interpretation within papers, conference presentations, or news reports, making sure they can be reliably regenerated is crucial for knowledge generation. By writing and sharing code used to generate brain visualizations, a direct and traceable link is established between the underlying data and the corresponding scientific figure. While the code that produces a replicable figure doesn't necessarily reflect the validity of the scientific finding or the accuracy of the associated content, it allows for reproducibility, instilling transparency and robustness, while demonstrating a desire to further scientific knowledge. Some even consider publishing figures which cannot be replicated as closer to advertising, rather than science (Steel, 2013).

While some GUI-based tools have historically offered command-line access to generate replicable visualizations, they can lack both the flexibility to easily generate publication ready figures and the benefits, such as iteration, provided by your preferred programming environment. Likewise, other GUI-based tools offer replicability in the form of automatically generated batch scripts or in-built terminals, which are often idiosyncratic and lack documentation to make them easily usable or reproducible by those not familiar with the specific software.

# Flexibility

Being able to exactly replicate your figures via code has marked advantages beyond positive open science practices. In particular, the ability to reprogram inputs (such as statistical maps) and settings (such as color schemes, thresholds, and visual orientations) can streamline your entire scientific workflow. Changing inputs and settings via code allows for the easy production of multiple figures, such as those resulting from multiple analyses which require similar visualizations. A simple for-loop or copying and pasting the code with altered input and/or settings-of-interest can be a powerful method for exploring visualization options or rapidly creating multi-panel figures. Likewise, an arduous request from a reviewer or collaborator to alter the image processing or analysis becomes less of a burden when the associated figures can be re-generated with a few lines of code, as opposed to re-pasting and re-illustrating them manually. Having a code-base with modifiable inputs can mean that the generation of visualizations requires less time, energy and effort than image and instance specific GUI-based generation. This also makes it easier to generate consistent figures across subsequent projects. Keep in mind that the gains of writing code for your figures are cumulative, and in addition to improving your programming, you start to build a code-base for figure generation that you can continue to reuse and share throughout your scientific career.

Precise controls via code over visualization settings, such as color schemes, legend placement and camera angles, can provide you with much greater flexibility over visualizations. Nonetheless, part of the appeal of GUI-based tools is that the presets for such settings can provide a useful starting point and reduce the decision burden on novice users. However, similar presets are often available in the form of default settings across most code-based packages, negating the need for the user to manually enter each and every choice of setting required for creating an image. Most code-based tools also come with documentation, with R-packages on the *CRAN* or *Neuroconductor* (Muschelli et al., 2019) repositories requiring detailed guidance. Recent packages have started to include detailed documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, & Mejia, 2022; Mowinckel & Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer & Ecker, 2020) which provide examples of figures that can be used as starting points or templates for new users. As the popularity of code-sharing for figure increases, there will be a cornucopia of templates that can be used as the basis for new figures.

While brain visualizations are often thought of as the end results of analyses, they also form a vital part of quality control for imaging data. Tools to automatically detect artefacts, de-noise the data and generate derivatives are becoming more robust, but we are not yet at the stage where visualizing the data during processing is no longer necessary. Nonetheless, when working with large datasets such as Human Connectome Project (Van Essen et al., 2013) or UK BioBank (Sudlow et al., 2015), it is unfeasible to use traditional GUI-based tool to visually examine the data. The time it takes to open a single file and achieve the desired visualization settings vastly compounds when working with large datasets. Knowing how to programmatically generate brain visualizations can allow you to iterate your visualization code over each image of a large datasets making checking the quality of each data processing or analysis step accessible and achievable. The visual outputs of each iteration can be complied into accessible documents which can be easily scrolled, with more advanced usage allowing for the creation of interactive HTML reports, similar to those created by standardized data processing tools like *fmriprep* (Esteban et al., 2019). Increasing capacity to conduct visual quality control on larger datasets can increase the identification of processing errors and result in more reliable and valid findings from your data.

# Integrative and Interactive Reporting

Often programming languages such as R, Python and MATLAB are used for the analysis and non-brain visualizations in neuroimaging studies, but the brain visualizations resulting from these analyses are outsourced to separate GUI's of tools such as *FSLeyes*, *Freeview* or *ITK-snap*. Switching from your analysis environment to a GUI-based visualization process can be a cumbersome deviation from the scientific workflow. This can make debugging errors more difficult, as you have to regularly switch program to visually examine the results of any modifications or adjustments to prior analyses. Using the brain visualization tools that already exist within your chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analyses.

Increasingly popular software such as *R Markdown*, *Quarto* and *Jupyter Notebook* allow for the mixing of prose and code in a single script, resulting in fully reproducible and publication ready papers. By using code-based tools available within your preferred environment, brain visualizations can be directly integrated and embedded within a paper or report. For instance, a fully reproducible version of the current paper can be found on [GitHub](https://github.com/sidchop/Brain_comms_field_potential). Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g. *e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code computational environment linked to the associated article using 'compute capsules' from the cloud-based platform Code Ocean (Code Ocean, 2021).

Neuroimaging data are often spatially 3D and can have multiple time points, adding a 4th dimension (e.g., fMRI data). Thus, communicating findings or evaluating quality using static 2D slices is challenging, and may not be the best representation of the data, or the interpretations. While well-curated 3D renderings can help with spatial localisation (see Madan, 2015; Pernet & Madan, 2019), in the end, static images can only provide an incomplete representation of the data, and forces researchers to choose the "best" angle or slice to show, which often involves compromising one result to emphasize another. An added advantage of some of the code-based tools is that you can generate 'rich' media like interactive widgets -- figures or animations, which allow users to zoom, rotate and scroll through slices. Interacting with a figure in this way can improve scientific communication of findings. Linking to or even embedding these videos or interactive figures in papers can greatly enhance the communication of findings and make your paper more engaging for the reader. Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of your research with peers and a reach larger audiences (Li and Xie, 2020). This last point is becoming increasingly salient as marketing science on social media has become a core medium for spreading discoveries, science communication to the public, and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019). Overall, public engagement is one of the cornerstones of science, and the images we create are at the center of the process.

\
**Table 1. Examples of code-based neuroimaging visualizations tools that can be accessed directly within R, MATLAB and Python environments.**

|                        | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
|:-----------------------|:--------|:--------|:--------------|:--------|:------------|
| **R**                  |         |         |               |         |             |
| ANTsR                  | ***+*** | ***+*** | ***+***       |         |             |
| brainconn              |         |         |               | ***+*** |             |
| brainR                 | ***+*** |         | ***+***       |         |             |
| ciftitools             | ***+*** | ***+*** | ***+*****\*** |         |             |
| fsbrain                | ***+*** | ***+*** | ***+*****\*** |         |             |
| ggseg                  |         |         | ***+***       |         |             |
| neurobase              | ***+*** |         |               |         |             |
| oro.nifti              | ***+*** |         |               |         |             |
| **Python**             |         |         |               |         |             |
| ANTsPy                 | ***+*** | ***+*** | ***+***       |         |             |
| brainiak               | ***+*** |         |               |         |             |
| Brainplotlib           |         | ***+*** | ***+*****\*** |         |             |
| Brainspace/surfplot    |         | ***+*** | ***+*****\*** |         |             |
| DIPY                   | ***+*** |         |               |         | ***+***     |
| ENIGMA toolbox         |         |         | ***+***       |         |             |
| FSLeyes                | ***+*** | ***+*** | ***+***       |         |             |
| ggseg                  |         |         | ***+***       |         |             |
| graphpype              |         |         |               | ***+*** |             |
| MMVT                   |         | ***+*** | ***+***       | ***+*** |             |
| MNE                    | ***+*** | ***+*** | ***+***       |         |             |
| mrivis                 | ***+*** |         |               |         |             |
| NaNSlice               | ***+*** |         |               |         |             |
| netneurotools          |         | ***+*** | ***+*****\*** |         |             |
| nilearn                | ***+*** | ***+*** | ***+***       | ***+*** |             |
| niwidget               | ***+*** | ***+*** |               |         | ***+***     |
| Pycortex               | ***+*** | ***+*** | ***+*****\*** |         |             |
| pySurfer               |         | ***+*** | ***+*****\*** |         |             |
| surfice                | ***+*** | ***+*** | ***+***       | ***+*** | ***+***     |
| surfplot               |         | ***+*** | ***+***       |         |             |
| Visbrain               | ***+*** | ***+*** | ***+***       | ***+*** |             |
| **MATLAB**             |         |         |               |         |             |
| BrainNetViewer         | ***+*** |         | ***+***       | ***+*** |             |
| Brainspace             |         | ***+*** | ***+*****\*** |         |             |
| Brainstorm             | ***+*** | ***+*** | ***+***       |         |             |
| bspmview               | ***+*** |         | ***+***       |         |             |
| CandlabCore            | ***+*** |         | ***+***       |         |             |
| ECoG/fMRI Vis toolbox  |         | ***+*** | ***+*****\*** |         |             |
| ENIGMA toolbox         |         |         | ***+***       |         |             |
| FieldTrip              | ***+*** | ***+*** |               |         |             |
| Lead-DBS               | ***+*** |         | ***+***       |         |             |
| mni2fs                 |         | ***+*** |               |         |             |
| mrtools                | ***+*** | ***+*** | ***+*****\*** |         |             |
| plotSurfaceROIBoundary |         | ***+*** | ***+*****\*** |         |             |
| Vistasoft              | ***+*** | ***+*** | ***+***       |         | ***+***     |

Note: The tools listed contain functionality required to generate (at least close-to) publication-ready neuroimaging figures via user-entered code within R, MATLAB and Python environments. This list does not include cross-platform general purpose visualization software.\
\*Cortex only

# Examples of Neuroimaging Visualization Packages available in *R*

The following four sections provide brief examples of packages and functions available in R for visualizing voxel, vertex, ROI and edge-level data. A didactic introduction to these tools is provided in an accompanying [collection of beginner friendly R-Notebooks](https://github.com/sidchop/neuRo-vis-BrainHack-2021). This is not an exhaustive list of packages available for visualizing brain data in R, rather, the following sections and accompanying notebooks aim to give the reader a sense of the available options, and an entry point to using them.

```{=tex}
\newpage
\blandscape
```
```{r Figure1 , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center'}

library(patchwork)
library(ggplot2)

fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
figs <- list()

for (f in 1:length(fig_name)) { 
  knitr::plot_crop(paste0('data/R/',fig_name[f],'.png'))
  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/R/',fig_name[f],'.png')),
                                                             width=ggplot2::unit(1,"npc"),
                                                             height=ggplot2::unit(1,"npc")))
}

patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]])+plot_layout(heights = c(0.5,1,1)))

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations in R') &
    theme(plot.tag = element_text(face = 'bold'),
          plot.title = element_text(face="bold"))

```

*Figure 1. Examples of brain imaging visualization make in using different R-packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown. Made using the the `ortho2` function from the `neurobase` package. B) A voxel-level cortical parcellation overlaid on T1-weighted MRI data shown in 9-slice axial orientation. Made using the `overlay` function from the `neurobase` package. C) A CIFTI format surface ROI atlas with a corresponding statistic assigned to each regions, with both hemispheres displayed on a inflated template surface in lateral view. Made using the `view_xifti_surface` from the `ciftiTools` package. D) A coronal cross-sectional rendering of subcortical structures where a statistical value has been assigned to each region. Made using the `aseg` atlas from the `ggseg` package. E) A 3D rendering of 9 bilateral subcortical regions where a statistical value has been assigned to each region. Made using the `aseg` atlas from the `ggseg3d` package. F) Medial and lateral views of a ROI atlas displayed on inflated template cortical surface where a statistical value has been assigned to each region. Made using the `glasser` atlas from the `ggsegGlasser` package, which was plotted using `ggseg`. G) Lateral view of a CIFTI format vertex-level data displayed on a inflated template surface. Made using the `view_xifti_surface` function from the `ciftiTools` package. H) Medial and lateral views of vertex-level data displayed on a individuals white matter surface. Made using the `vis.subject.morph.standard` function from the `fsbrain` package. I) A weighted and undirected graph plotted on top, left and front views of a template of a brain rendering. Made using the `brainconn` function from the `brainconn` package. All code used to compile this figure, as well as the contents of each panel are provided in the Supplement.

\elandscape

## Voxel-Level Visualizations.

The `oro.nifti` and `neurobase` packages allow for the reading, writing, manipulation and visualization of voxel-level imaging data of either *nifti*, *analyze* or *afni* formats. In particular, the `image`, `slice` and `ortho2` functions allow users to display the desired number of slices in the desired orientations, while also providing precise control over overlaid images, color scales, legend placement and other aesthetic settings (Fig 1A-B). Smoothing and re-sampling of images is sometimes required for visualization purposes, which can be achieved via the `ANTsR` or `fslr` packages.

## Vertex-Level Visualizations.

The recently developed `fsbrain` package allows for the visualization of vertex-level and cortical ROI-level data which were derived using *FreeSurfer*. The package contains many flexible functions for visualizing individual-level and group-level measures such as cortical thickness, volume or surface area (Fig 1F). In addition to producing publication-ready figures, the visualization functions can allow for visual quality control of large datasets processed using FreeSurfer. The package also comes with extensive guides and documentations to assist users (Schäfer & Ecker, 2020).

CIFTI or 'grey-ordinate' data is becoming a popular format for structural and functional imaging data, as it combines vertex-level data of the cortex with voxel-level data for the cerebellum and sub-cortex. In R, CIFTI data can now be read, visualized and manipulated using the well-documented `ciftiTools` package (Fig 1C,1G; Pham, Muschelli, & Mejia, 2022). This package allows for both vertex-level visualization of the surface as well as voxel-level visualization of cerebellum and subcortex.

## ROI-Level Visualizations.

Statistical parameters can be mapped onto discreet cortical and sub-cortical regions using the `ggseg` and `ggseg3d` packages (Mowinckel & Vidal-Piñeiro, 2020). These packages flexibly generates aesthetic renderings of cortical, sub-cortical and white matter ROIs in 2D (Fig 1D,1F) and 3D (Fig 1E). Usually, ROI-level interpretations are dependent on a standardized atlas or parcellation scheme of the brain and accordingly, this package and its sister-packages (`ggsegExtra`) provide a large array of commonly used atlases and some functionality for users to contribute other atlases. Additionally, being part of the 'Grammar of Graphics' framework (Wickham, 2016) allows for integration with packages such as `gganimate`, enabling users to animate and visualize dynamic changes in spatial and temporal statistics across ROIs. Notably, the `ciftiTools` package also provides functions to visualize surface and voxel-level ROIs for the cortex and sub-cortex respectively (Fig 1E), flexibly allowing CIFTI format atlases and ROIs to be displayed.

## Edge-Level Visualizations.

Analyses which divide the brain into discrete regions and examine pair-wise dependencies between regional phenotypes are becoming increasingly common. Often the results of these dependency analyses are statistics relating to 'edges' or links between any two regions. The package `brainconn` allows users to plot brain regions as nodes of a network graph, and dependencies between regions as edges of that graph, following a standardized coordinate space in both 2D and interactive 3D (Figure 1I). The graphs can be weighted, with the strength of the edges represented with color or line thickness, as well as directed, where arrows will be used in place of edges. This package comes with node coordinates corresponding to multiple commonly used atlases and contains functionality for users to easily enter their own custom atlas coordinates.

# Limitations and Functionality Gaps

Most of the tools introduced above do not require a strong knowledge of programming, but there is still a steeper learning curve when compared to use a GUI. This is especially true when learning how to make fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning, for the purpose of a publication-ready figure. While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability. Relatedly, while some interactive image viewers which can be opened within an integrated development environment (e.g. Muschelli, 2016), for quick and interactive viewing of single images, GUI tool can be faster and more practical.

Often cerebellar and brain-stem regions are not well represented in software (e.g. Figure 1), potentially mirroring the cortico-centric sentiment that has prevailed in human neuroimaging research (Chin, Chang, & Holmes, 2022). Likewise, custom non-cortical atlases such as non-standard subcortical atlas schemes are not yet straight forward, and usually require multiple functions and packages to visualize. Finally, some neuroimaging related data types, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools.

As can be seen in Table 1, there are usually multiple packages within each programming environment which are capable of visualizing each data type. While this provide choice for advanced users, it can also lead to confusion for novice users who may not be familiar with the nuanced differences between tools. Future work should continue to consolidate brain visualization methods into unified beginner-friendly code-based tools which are capable of plotting multiple data types.

\newpage

# References

Poldrack, R. A., Baker, C. I., Durnez, J., Gorgolewski, K. J., Matthews, P. M., Munafò, M. R., ... & Yarkoni, T. (2017). Scanning the horizon: towards transparent and reproducible neuroimaging research. Nature reviews neuroscience, 18(2), 115-126.

Steel, G. (2013, September 3). Publishing research without data is simply advertising, not science. Open Knowledge Foundation. <https://blog.okfn.org/2013/09/03/publishing-research-without-data-is-simply-advertising-not-science/>

The Comprehensive R Archive Network. (n.d.). Retrieved April 21, 2022, from <https://cran.r-project.org/>

Muschelli, J., Gherman, A., Fortin, J. P., Avants, B., Whitcher, B., Clayden, J. D., ... & Crainiceanu, C. M. (2019). Neuroconductor: an R platform for medical imaging analysis. *Biostatistics*, *20*(2), 218-239.

Pham, D., Muschelli, J., & Mejia, A. (2022). ciftiTools: A package for reading, writing, visualizing, and manipulating CIFTI files in R. *NeuroImage*, 118877.

Mowinckel, A. M., & Vidal-Piñeiro, D. (2020). Visualization of brain statistics with R packages ggseg and ggseg3d. *Advances in Methods and Practices in Psychological Science*, *3*(4), 466-483.

Huntenburg, J., Abraham, A., Loula, J., Liem, F., Dadi, K., & Varoquaux, G. (2017). Loading and plotting of cortical surface representations in Nilearn. *Research Ideas and Outcomes*, *3*, e12342.

Wickham H. (2016). *ggplot2: Elegant Graphics for Data Analysis.* Springer-Verlag New York. ISBN 978-3-319-24277-4, <https://ggplot2.tidyverse.org>.

Muschelli, J. (2016). *papayar*. GitHub repository. <https://github.com/muschellij2/papayar>

David C. Van Essen, Stephen M. Smith, Deanna M. Barch, Timothy E.J. Behrens, Essa Yacoub, Kamil Ugurbil, for the WU-Minn HCP Consortium. (2013). The WU-Minn Human Connectome Project: An overview. NeuroImage 80(2013):62-79.

Sudlow, C., Gallacher, J., Allen, N., Beral, V., Burton, P., Danesh, J., ... & Collins, R. (2015). UK biobank: an open access resource for identifying the causes of a wide range of complex diseases of middle and old age. PLoS medicine, 12(3), e1001779.

Esteban, O., Markiewicz, C. J., Blair, R. W., Moodie, C. A., Isik, A. I., Erramuzpe, A., ... & Gorgolewski, K. J. (2019). fMRIPrep: a robust preprocessing pipeline for functional MRI. Nature methods, 16(1), 111-116.

Pernet, C., & Madan, C. R. (2019). Data visualization for inference in tomographic brain imaging.

Code Ocean. (2021, October 27). Computational Research Platform on. Retrieved April 21, 2022, from <https://codeocean.com/>

Madan, C. R. (2015). Creating 3D visualizations of MRI data: A brief guide. F1000Research, 4.

Baker M. (2015). Social media: A network boost. Nature 518: 263--265. 10.1038/nj7538-263a

Lee J.-S. (2019). How to use Twitter to further your research career. Nature. 10.1038/d41586-019-00535-w

Li, Y., & Xie, Y. (2020). Is a picture worth a thousand words? An empirical study of image content and social media engagement. Journal of Marketing Research, 57(1), 1-19.

Schäfer, T., & Ecker, C. (2020). fsbrain: an R package for the visualization of structural neuroimaging data. bioRxiv.

Chin, R., Chang, S. W., & Holmes, A. J. (2022). Beyond cortex: The evolution of the human brain. *Psychological Review*.

\newpage

# Supplement (Code used to generate Figures)

\singlespacing

```{r Figure1-sup , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center', eval=FALSE}

library(patchwork)
library(ggplot2)

fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
figs <- list()

for (f in 1:length(fig_name)) { 
  knitr::plot_crop(paste0('data/',fig_name[f],'.png'))
  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/',fig_name[f],'.png')),
                                                             width=ggplot2::unit(1,"npc"),
                                                             height=ggplot2::unit(1,"npc")))
}

patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]])+plot_layout(heights = c(0.7,1,1)))
patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations in R') &
    theme(plot.tag = element_text(face = 'bold'),
          plot.title = element_text(face="bold"))

```

```{r Fig1a, eval=FALSE}
#Figure 1A (Voxel)

library(neurobase)

#Load in a nifti (.nii.gz) file of a standardized template so use as a background. 
template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R

#Load in a nifti (.nii.gz) file of the statistic you wanted of overlay, with the same dimensions as the template.  
effect <- readnii("data/MNI152_effect_size.nii.gz")

#Threshold the statistic map so only values about a specified amount are displayed
effect[effect<0.3] <- NA

#Set the breaks/intervals you want on the color bar (e.g. from .3 to .6, by intervals of 0.005)
breaks = seq(.3,.6, by=0.005)

#[Optional] If you want to output the figure as a .png, open a png image device
png("data/fig1a.png")

#Use the ortho2 function from the neurobase package to make the figure
ortho2(x = template,   #Specify the background template
       y = effect,     #Specify the effect you want to overlay on the template
       crosshairs = F, #Remove the cross-hairs
       bg = "white",   #Make the background white 
       NA.x = T,       #Do not display NA values
       col.y= rev(viridis::inferno(n = 500)), #Select the color scale. In this case I have used the inferno scale from the viridis package. 
       xyz = c(70,50,80), #set the x y & z slice you want to visualize
       useRaster = T,     #Sometimes using Raster makes for clearer plots
       ycolorbar = TRUE,  #Add a color bar
       mfrow = c(1,1)) +  #layout of brain views (for this figure, we just want a axial view)
  colorbar(breaks, col = rev(viridis::inferno(n=length(breaks)-1)), #add a color bar with same color scale 
           text.col = "black", labels = TRUE, maxleft = 0.95)  #Set the specifications for the color bar

dev.off()

```

```{r Fig1b, eval=FALSE }
#Figure 1B (Voxel)

library(neurobase)
library(scales)

#Load in a nifti (.nii.gz) file to use as a background. 
t1 <- readnii('data/sub-001__t1_warped.nii.gz')

#Load in a nifti (.nii.gz) file to use as a overlay. 
atlas <-  readnii('data/sub-001_schaefer300n7_aseg_to_dwispace_gm_rois.nii.gz')

png('data/fig1b.png')
overlay(x=robust_window(t1), y=atlas, 
        plot.type = "single", 
        z=c(seq(30,110,10)), 
        col.y =  alpha(viridis::inferno(300), 0.6),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, 
        zlim.y =c(1,300))
dev.off()

```

```{r Fig1c , eval=FALSE}
#vertex 1
#Figure 1C
library(rgl)
library(ggplot2)
# Load the package and point to the Connectome Workbench
library(ciftiTools)

#Need connectome workbench to use ciftiTools - set path below
ciftiTools.setOption("wb_path", "/Applications/workbench/")

# Read in CIFTI file
cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
surfL_fname <- ciftiTools.files()$surf["left"]
surfR_fname <- ciftiTools.files()$surf["right"]

cii <- read_cifti(
  cifti_fname, brainstructures="all", 
  surfL_fname=surfL_fname,
  surfR_fname=surfR_fname)

#Plot cifti surface
view_xifti_surface(cii, 
                   hemisphere = "both", 
                   view = "lateral", 
                   idx = 1,
                   colors = viridis::plasma(n = 100),
                   zlim = c(1.2,1.5),  
                   legend_embed = T,
                   cex.title = 2)

rgl.snapshot("data/fig1c.png")
rgl.close()

```

```{r Fig1d , eval=FALSE}
#Vertex 2
#Figure 1X - Remove white space

library(fsbrain)
library(rgl)

#download_optional_data()
#download_fsaverage(accept_freesurfer_license = TRUE)

subjects_dir <- get_optional_data_filepath("subjects_dir")

subject_id <- 'subject1'

colourmap <- colorRampPalette(viridis::plasma(n = 1000))
fsbrain.set.default.figsize(700, 700);

brain = vis.subject.morph.native(subjects_dir, 
                                subject_id, 'sulc', 
                                cortex_only = T, 
                                views=NULL, 
                                draw_colorbar = FALSE,
                                makecmap_options = list('colFn'=colourmap))

img = export(brain, draw_colorbar = FALSE, output_img = "data/fig1d.png")

```

```{r Fig1e , eval=FALSE}
#ROI1 
#Figure 1e 
library(ciftiTools)
parc <- load_parc("Yeo_7")
view_xifti_surface(parc,  
                   legend_embed = T,
                   hemisphere = "right")

rgl.snapshot("data/fig1e.png")

rgl.close()

```

```{r Fig1f , eval=FALSE}
#ROI1 
#Figure 1X - combine with previous figure

set.seed(1993) #set a random seed (good practice for reproducibility)

parc <- load_parc("Schaefer_400")

ramdom_metric <- rnorm(400)

cii <- move_from_mwall(cii, NA)

parc_vec <- c(as.matrix(parc))

xii_metric <- c(NA, ramdom_metric)[parc_vec + 1]

xii1 <- select_xifti(cii, 1)

xii_metric <- newdata_xifti(xii1, xii_metric)

plot2 <- view_xifti_surface(xii_metric, 
                            colors = viridis::viridis(n = 400), 
                            borders = "black", 
                            hemisphere = "both",
                            view = 'lateral')

rgl.snapshot("data/fig1f.png")

rgl.close()

```

```{r Fig1g, eval=FALSE}
#ROI2

#devtools::install_github("LCBC-UiO/ggsegGlasser")
library(ggseg)
library(ggplot2)
library(ggsegGlasser)

set.seed(1993) #set a random seed (good practice for reproducibility)

base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))

colnames(base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas

base_atlas <- cbind(Effect, base_atlas)

cortex <- ggseg(atlas = glasser,
                .data = base_atlas,
                mapping=aes(fill=Effect), 
                position="stacked",
                colour="black",
                size=.2,
                show.legend = T,
                plot.background = "white") +
  scale_fill_viridis_b() + 
   theme_void() +
    theme(legend.position = "bottom")

sub_base_atlas <- as.data.frame(na.omit(cbind(aseg$data$region, aseg$data$hemi)))

colnames(sub_base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(sub_base_atlas)[1])

sub_base_atlas <- cbind(Effect, sub_base_atlas)

subcortex <- ggseg(atlas = aseg,
                   .data = sub_base_atlas,
                   mapping=aes(fill=Effect), 
                   position = "dispersed",
                   colour="black",
                   hemi =c('left', 'right'),
                   size=.2,
                   show.legend = T, 
                   plot.background = "white") + 
  scale_fill_viridis_b()

ggsave(filename = 'data/fig1g1.png', plot = cortex, device = 'png', bg = 'white')
ggsave(filename = 'data/fig1g2.png', plot = subcortex, device = 'png', bg = 'white')

```

```{r, eval=FALSE}

make_ggseg3d <- function(attribute, 
                         colour.pal = c("light yellow",
                                        "orange",
                                        "red",
                                        "dark red"), 
                         hide.colourbar=FALSE, 
                         output.png=FALSE, 
                         file.name="ggseg_3d.png") {
  # Inputs: 
  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left-Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Right-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Right-Amygdala, Right-Accumbens-area

  library(ggseg3d)
  library(tidyr)
  library(dplyr)
  #remove(aseg_3d)
  aseg_3d <- aseg_3d
  aseg_3d <- tidyr::unnest(aseg_3d, cols = c(ggseg_3d))

  attribute.ggseg3d <- c(rep(NA, 4), attribute[1:4], rep(NA, 3), attribute[5:7],
                         rep(NA, 5), attribute[8:14],  rep(NA, 6))

  data <- dplyr::mutate(aseg_3d, attribute =  attribute.ggseg3d)
  #remove NA regions
  aseg_3d[which(is.na(data$attribute)),] <- NA
  aseg_3d <- tidyr::drop_na(aseg_3d)

  data[which(is.na(data$attribute)),] <- NA
  data <- tidyr::drop_na(data)
  data$attribute[data$attribute==0]<-NA #make 0 valus NA so they are set as grey in ggseg3d

  scene=list(camera = list(eye = list(x = 0, y = 1, z = -2.25)),
             aspectratio = list(x=1.6,y=1.6,z=1.6))

  plot <- ggseg3d::ggseg3d(.data = data, 
                           atlas = aseg_3d, 
                           colour = "attribute", 
                           text = "attribute",
                           palette = colour.pal)
  plot <- remove_axes(plot)
  plot <- plotly::layout(plot, 
                         scene = scene, 
                         width = 600, height = 600)

  if(hide.colourbar==TRUE){
    plot <- plotly::hide_colorbar(plot)
  }

  if( output.png==TRUE){
    plotly::orca(plot, file =  file.name)
  }

  return(plot)

}

set.seed(1993)
volume <- sample(1:20, 14) #randomly generated data for 14 structures
make_ggseg3d(volume, 
             colour.pal = viridis::viridis(n=length(volume)), 
             output.png = T, file.name = "data/fig1g3.png")

```

```{r Fig1h , eval = F}
#Edge (also add 3d version)
#Fig1X

#devtools::install_github("sidchop/brainconn")
library(brainconn)

conmat <- example_weighted_undirected
degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]

brainconn(atlas ="schaefer300_n7", 
                   conmat=conmat, 
                   node.size = degree/2,
                   view="top", 
                   edge.color.weighted = T,
                   background.alpha = 0.8, 
                   show.legend = F, 
                   edge.color = scale_edge_colour_viridis())

ggsave(filename = 'data/fig1h.png', 
       plot = fig1h, 
       device = 'png',
       bg = 'white',
       height = 7, 
       units = "in")

brainconn3D(atlas ="schaefer300_n7", 
            conmat=conmat,
            show.legend = F)

```

### 2022-08-15 manuscript.tex
first data push

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

\title{Reproducible and Programmatic Generation of Neuroimaging Visualizations}

\author{
    Sidhant Chopra
    \thanks{Corresponding author}
   \\
    Department of Psychology \\
    Yale University \\
  CT, USA, 06511 \\
  \texttt{\href{mailto:sidhant.chopra@yale.com}{\nolinkurl{sidhant.chopra@yale.com}}} \\
   \And
    Loïc Labache
   \\
    Department of Psychology \\
    Yale University \\
  CT, USA, 06511 \\
  \texttt{} \\
   \And
    Elvisha Dhamala
   \\
    Department of Psychology \\
    Yale University \\
  CT, USA, 06511 \\
  \texttt{} \\
   \And
    Edwina R Orchard
   \\
    Department of Psychology \\
    Yale University \\
  CT, USA, 06511 \\
  \texttt{} \\
   \And
    Avram Holmes
   \\
    Department of Psychology \\
    Yale University \\
  CT, USA, 06511 \\
  \texttt{} \\
  }

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{pdflscape}
\newcommand{\blandscape}{\begin{landscape}}
\newcommand{\elandscape}{\end{landscape}}
\usepackage{setspace}\doublespacing
\begin{document}
\maketitle

\begin{abstract}
Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results. Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI). There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB. Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process. This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments. Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data.
\end{abstract}

\keywords{
    Neuroimaging visualization
   \and
    Reproducibility
   \and
    Programmatic figures
   \and
    Open Science
   \and
    R
   \and
    Python
   \and
    Brain Visualization
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The visualization of neuroimaging data is one of the primary ways in which we evaluate data quality, interpret results, and communicate findings. These visualizations are commonly produced using graphical user interface-based (GUI) tools where individual images are opened and, within each instance, display settings are manually changed until the desired output is reached. In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility. However, there now exists many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community (Table 1). These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments.

The present article covers three major advantages of using code-generated visualizations over GUI based tools: replicability, flexibility, and integration. First, by generating visualizations using code, you increase the replicability of your figures for yourself, your collaborators and your readers. Second, code-based tools provides more precise controls over display settings, while also benefiting from the advantages of programming, such as the ability to iteratively and rapidly generate multiple figures. Finally, being able to integrate figure generation into your analysis scripts reduces chances of errors and increases the accessibility of resulting analyses pipeline, with more advanced tools now allowing for the seamless integration of prose and code. Here, we review the advantages of learning and using programmatic neuroimaging visualizations, focusing on benefits to replicability, flexibility and integration. We conclude by introducing examples of R-packages which allow for visualization of statistics at region-of-interest (ROI), voxel, vertex, and edge level data, followed by a brief discussion of limitations and functionality gaps in code-based brain visualization tools.

\hypertarget{replicability}{%
\section{Replicability}\label{replicability}}

In recent years, there have been multiple large-scale efforts empirically demonstrating the lack of reproducibility of findings from neuroimaging data (Poldrack et al., 2017). One common solution proposed for achieving robust and reliable discoveries has been to encourage scientific output which can be transparently evaluated and independently replicated. In practice, this typically entails openly sharing detailed methods, materials, code, and data. While there is a trend towards increasing transparency and code sharing of neuroimaging analyses, the sharing of code used to generate figures which include brain renderings and spatial maps has been relatively neglected. This gap in reproducibility is partly driven by the fact that brain figures are often created using a manual process that involves tinkering with sliders, buttons, and overlays on a GUI, concluding with a screenshot and sometimes beautification in image processing software like Illustrator or Photoshop. Such a process inherently makes neuroimaging visualizations difficult, if not impossible to replicate, even by the authors themselves.

Visualization scripts should reflect a core feature of open science. Given that brain figures regularly form the centerpiece of interpretation within papers, conference presentations, or news reports, making sure they can be reliably regenerated is crucial for knowledge generation. By writing and sharing code used to generate brain visualizations, a direct and traceable link is established between the underlying data and the corresponding scientific figure. While the code that produces a replicable figure doesn't necessarily reflect the validity of the scientific finding or the accuracy of the associated content, it allows for reproducibility, instilling transparency and robustness, while demonstrating a desire to further scientific knowledge. Some even consider publishing figures which cannot be replicated as closer to advertising, rather than science (Steel, 2013).

While some GUI-based tools have historically offered command-line access to generate replicable visualizations, they can lack both the flexibility to easily generate publication ready figures and the benefits, such as iteration, provided by your preferred programming environment. Likewise, other GUI-based tools offer replicability in the form of automatically generated batch scripts or in-built terminals, which are often idiosyncratic and lack documentation to make them easily usable or reproducible by those not familiar with the specific software.

\hypertarget{flexibility}{%
\section{Flexibility}\label{flexibility}}

Being able to exactly replicate your figures via code has marked advantages beyond positive open science practices. In particular, the ability to reprogram inputs (such as statistical maps) and settings (such as color schemes, thresholds, and visual orientations) can streamline your entire scientific workflow. Changing inputs and settings via code allows for the easy production of multiple figures, such as those resulting from multiple analyses which require similar visualizations. A simple for-loop or copying and pasting the code with altered input and/or settings-of-interest can be a powerful method for exploring visualization options or rapidly creating multi-panel figures. Likewise, an arduous request from a reviewer or collaborator to alter the image processing or analysis becomes less of a burden when the associated figures can be re-generated with a few lines of code, as opposed to re-pasting and re-illustrating them manually. Having a code-base with modifiable inputs can mean that the generation of visualizations requires less time, energy and effort than image and instance specific GUI-based generation. This also makes it easier to generate consistent figures across subsequent projects. Keep in mind that the gains of writing code for your figures are cumulative, and in addition to improving your programming, you start to build a code-base for figure generation that you can continue to reuse and share throughout your scientific career.

Precise controls via code over visualization settings, such as color schemes, legend placement and camera angles, can provide you with much greater flexibility over visualizations. Nonetheless, part of the appeal of GUI-based tools is that the presets for such settings can provide a useful starting point and reduce the decision burden on novice users. However, similar presets are often available in the form of default settings across most code-based packages, negating the need for the user to manually enter each and every choice of setting required for creating an image. Most code-based tools also come with documentation, with R-packages on the \emph{CRAN} or \emph{Neuroconductor} (Muschelli et al., 2019) repositories requiring detailed guidance. Recent packages have started to include detailed documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, \& Mejia, 2022; Mowinckel \& Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer \& Ecker, 2020) which provide examples of figures that can be used as starting points or templates for new users. As the popularity of code-sharing for figure increases, there will be a cornucopia of templates that can be used as the basis for new figures.

While brain visualizations are often thought of as the end results of analyses, they also form a vital part of quality control for imaging data. Tools to automatically detect artefacts, de-noise the data and generate derivatives are becoming more robust, but we are not yet at the stage where visualizing the data during processing is no longer necessary. Nonetheless, when working with large datasets such as Human Connectome Project (Van Essen et al., 2013) or UK BioBank (Sudlow et al., 2015), it is unfeasible to use traditional GUI-based tool to visually examine the data. The time it takes to open a single file and achieve the desired visualization settings vastly compounds when working with large datasets. Knowing how to programmatically generate brain visualizations can allow you to iterate your visualization code over each image of a large datasets making checking the quality of each data processing or analysis step accessible and achievable. The visual outputs of each iteration can be complied into accessible documents which can be easily scrolled, with more advanced usage allowing for the creation of interactive HTML reports, similar to those created by standardized data processing tools like \emph{fmriprep} (Esteban et al., 2019). Increasing capacity to conduct visual quality control on larger datasets can increase the identification of processing errors and result in more reliable and valid findings from your data.

\hypertarget{integrative-and-interactive-reporting}{%
\section{Integrative and Interactive Reporting}\label{integrative-and-interactive-reporting}}

Often programming languages such as R, Python and MATLAB are used for the analysis and non-brain visualizations in neuroimaging studies, but the brain visualizations resulting from these analyses are outsourced to separate GUI's of tools such as \emph{FSLeyes}, \emph{Freeview} or \emph{ITK-snap}. Switching from your analysis environment to a GUI-based visualization process can be a cumbersome deviation from the scientific workflow. This can make debugging errors more difficult, as you have to regularly switch program to visually examine the results of any modifications or adjustments to prior analyses. Using the brain visualization tools that already exist within your chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analyses.

Increasingly popular software such as \emph{R Markdown}, \emph{Quarto} and \emph{Jupyter Notebook} allow for the mixing of prose and code in a single script, resulting in fully reproducible and publication ready papers. By using code-based tools available within your preferred environment, brain visualizations can be directly integrated and embedded within a paper or report. For instance, a fully reproducible version of the current paper can be found on \href{https://github.com/sidchop/Brain_comms_field_potential}{GitHub}. Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g.~\emph{e-Life}, \emph{Aperture Neuro}), with other journals like \emph{F1000Research} and \emph{GigaScience} even allowing on-demand re-running of code computational environment linked to the associated article using `compute capsules' from the cloud-based platform Code Ocean (Code Ocean, 2021).

Neuroimaging data are often spatially 3D and can have multiple time points, adding a 4th dimension (e.g., fMRI data). Thus, communicating findings or evaluating quality using static 2D slices is challenging, and may not be the best representation of the data, or the interpretations. While well-curated 3D renderings can help with spatial localisation (see Madan, 2015; Pernet \& Madan, 2019), in the end, static images can only provide an incomplete representation of the data, and forces researchers to choose the ``best'' angle or slice to show, which often involves compromising one result to emphasize another. An added advantage of some of the code-based tools is that you can generate `rich' media like interactive widgets -- figures or animations, which allow users to zoom, rotate and scroll through slices. Interacting with a figure in this way can improve scientific communication of findings. Linking to or even embedding these videos or interactive figures in papers can greatly enhance the communication of findings and make your paper more engaging for the reader. Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of your research with peers and a reach larger audiences (Li and Xie, 2020). This last point is becoming increasingly salient as marketing science on social media has become a core medium for spreading discoveries, science communication to the public, and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019). Overall, public engagement is one of the cornerstones of science, and the images we create are at the center of the process.

\hfill\break
\textbf{Table 1. Examples of code-based neuroimaging visualizations tools that can be accessed directly within R, MATLAB and Python environments.}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.3038}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1139}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1139}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1899}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1139}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1646}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Voxel
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Vertex
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
ROI
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Edge
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Streamlines
\end{minipage} \\
\midrule()
\endhead
\textbf{R} & & & & & \\
ANTsR & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & & \\
brainconn & & & & \textbf{\emph{+}} & \\
brainR & \textbf{\emph{+}} & & \textbf{\emph{+}} & & \\
ciftitools & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
fsbrain & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
ggseg & & & \textbf{\emph{+}} & & \\
neurobase & \textbf{\emph{+}} & & & & \\
oro.nifti & \textbf{\emph{+}} & & & & \\
\textbf{Python} & & & & & \\
ANTsPy & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & & \\
brainiak & \textbf{\emph{+}} & & & & \\
Brainplotlib & & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
Brainspace/surfplot & & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
DIPY & \textbf{\emph{+}} & & & & \textbf{\emph{+}} \\
ENIGMA toolbox & & & \textbf{\emph{+}} & & \\
FSLeyes & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & & \\
ggseg & & & \textbf{\emph{+}} & & \\
graphpype & & & & \textbf{\emph{+}} & \\
MMVT & & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & \\
MNE & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & & \\
mrivis & \textbf{\emph{+}} & & & & \\
NaNSlice & \textbf{\emph{+}} & & & & \\
netneurotools & & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
nilearn & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & \\
niwidget & \textbf{\emph{+}} & \textbf{\emph{+}} & & & \textbf{\emph{+}} \\
Pycortex & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
pySurfer & & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
surfice & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} \\
surfplot & & \textbf{\emph{+}} & \textbf{\emph{+}} & & \\
Visbrain & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & \\
\textbf{MATLAB} & & & & & \\
BrainNetViewer & \textbf{\emph{+}} & & \textbf{\emph{+}} & \textbf{\emph{+}} & \\
Brainspace & & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
Brainstorm & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & & \\
bspmview & \textbf{\emph{+}} & & \textbf{\emph{+}} & & \\
CandlabCore & \textbf{\emph{+}} & & \textbf{\emph{+}} & & \\
ECoG/fMRI Vis toolbox & & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
ENIGMA toolbox & & & \textbf{\emph{+}} & & \\
FieldTrip & \textbf{\emph{+}} & \textbf{\emph{+}} & & & \\
Lead-DBS & \textbf{\emph{+}} & & \textbf{\emph{+}} & & \\
mni2fs & & \textbf{\emph{+}} & & & \\
mrtools & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
plotSurfaceROIBoundary & & \textbf{\emph{+}} & \textbf{\emph{+}*} & & \\
Vistasoft & \textbf{\emph{+}} & \textbf{\emph{+}} & \textbf{\emph{+}} & & \textbf{\emph{+}} \\
\bottomrule()
\end{longtable}

Note: The tools listed contain functionality required to generate (at least close-to) publication-ready neuroimaging figures via user-entered code within R, MATLAB and Python environments. This list does not include cross-platform general purpose visualization software.\\
*Cortex only

\hypertarget{examples-of-neuroimaging-visualization-packages-available-in-r}{%
\section{\texorpdfstring{Examples of Neuroimaging Visualization Packages available in \emph{R}}{Examples of Neuroimaging Visualization Packages available in R}}\label{examples-of-neuroimaging-visualization-packages-available-in-r}}

The following four sections provide brief examples of packages and functions available in R for visualizing voxel, vertex, ROI and edge-level data. A didactic introduction to these tools is provided in an accompanying \href{https://github.com/sidchop/neuRo-vis-BrainHack-2021}{collection of beginner friendly R-Notebooks}. This is not an exhaustive list of packages available for visualizing brain data in R, rather, the following sections and accompanying notebooks aim to give the reader a sense of the available options, and an entry point to using them.

\newpage
\blandscape

\begin{center}\includegraphics[width=0.8\linewidth,height=0.8\textheight]{manuscript_files/figure-latex/Figure1 -1} \end{center}

\emph{Figure 1. Examples of brain imaging visualization make in using different R-packages.} A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown. Made using the the \texttt{ortho2} function from the \texttt{neurobase} package. B) A voxel-level cortical parcellation overlaid on T1-weighted MRI data shown in 9-slice axial orientation. Made using the \texttt{overlay} function from the \texttt{neurobase} package. C) A CIFTI format surface ROI atlas with a corresponding statistic assigned to each regions, with both hemispheres displayed on a inflated template surface in lateral view. Made using the \texttt{view\_xifti\_surface} from the \texttt{ciftiTools} package. D) A coronal cross-sectional rendering of subcortical structures where a statistical value has been assigned to each region. Made using the \texttt{aseg} atlas from the \texttt{ggseg} package. E) A 3D rendering of 9 bilateral subcortical regions where a statistical value has been assigned to each region. Made using the \texttt{aseg} atlas from the \texttt{ggseg3d} package. F) Medial and lateral views of a ROI atlas displayed on inflated template cortical surface where a statistical value has been assigned to each region. Made using the \texttt{glasser} atlas from the \texttt{ggsegGlasser} package, which was plotted using \texttt{ggseg}. G) Lateral view of a CIFTI format vertex-level data displayed on a inflated template surface. Made using the \texttt{view\_xifti\_surface} function from the \texttt{ciftiTools} package. H) Medial and lateral views of vertex-level data displayed on a individuals white matter surface. Made using the \texttt{vis.subject.morph.standard} function from the \texttt{fsbrain} package. I) A weighted and undirected graph plotted on top, left and front views of a template of a brain rendering. Made using the \texttt{brainconn} function from the \texttt{brainconn} package. All code used to compile this figure, as well as the contents of each panel are provided in the Supplement.

\end{landscape}

\hypertarget{voxel-level-visualizations.}{%
\subsection{Voxel-Level Visualizations.}\label{voxel-level-visualizations.}}

The \texttt{oro.nifti} and \texttt{neurobase} packages allow for the reading, writing, manipulation and visualization of voxel-level imaging data of either \emph{nifti}, \emph{analyze} or \emph{afni} formats. In particular, the \texttt{image}, \texttt{slice} and \texttt{ortho2} functions allow users to display the desired number of slices in the desired orientations, while also providing precise control over overlaid images, color scales, legend placement and other aesthetic settings (Fig 1A-B). Smoothing and re-sampling of images is sometimes required for visualization purposes, which can be achieved via the \texttt{ANTsR} or \texttt{fslr} packages.

\hypertarget{vertex-level-visualizations.}{%
\subsection{Vertex-Level Visualizations.}\label{vertex-level-visualizations.}}

The recently developed \texttt{fsbrain} package allows for the visualization of vertex-level and cortical ROI-level data which were derived using \emph{FreeSurfer}. The package contains many flexible functions for visualizing individual-level and group-level measures such as cortical thickness, volume or surface area (Fig 1F). In addition to producing publication-ready figures, the visualization functions can allow for visual quality control of large datasets processed using FreeSurfer. The package also comes with extensive guides and documentations to assist users (Schäfer \& Ecker, 2020).

CIFTI or `grey-ordinate' data is becoming a popular format for structural and functional imaging data, as it combines vertex-level data of the cortex with voxel-level data for the cerebellum and sub-cortex. In R, CIFTI data can now be read, visualized and manipulated using the well-documented \texttt{ciftiTools} package (Fig 1C,1G; Pham, Muschelli, \& Mejia, 2022). This package allows for both vertex-level visualization of the surface as well as voxel-level visualization of cerebellum and subcortex.

\hypertarget{roi-level-visualizations.}{%
\subsection{ROI-Level Visualizations.}\label{roi-level-visualizations.}}

Statistical parameters can be mapped onto discreet cortical and sub-cortical regions using the \texttt{ggseg} and \texttt{ggseg3d} packages (Mowinckel \& Vidal-Piñeiro, 2020). These packages flexibly generates aesthetic renderings of cortical, sub-cortical and white matter ROIs in 2D (Fig 1D,1F) and 3D (Fig 1E). Usually, ROI-level interpretations are dependent on a standardized atlas or parcellation scheme of the brain and accordingly, this package and its sister-packages (\texttt{ggsegExtra}) provide a large array of commonly used atlases and some functionality for users to contribute other atlases. Additionally, being part of the `Grammar of Graphics' framework (Wickham, 2016) allows for integration with packages such as \texttt{gganimate}, enabling users to animate and visualize dynamic changes in spatial and temporal statistics across ROIs. Notably, the \texttt{ciftiTools} package also provides functions to visualize surface and voxel-level ROIs for the cortex and sub-cortex respectively (Fig 1E), flexibly allowing CIFTI format atlases and ROIs to be displayed.

\hypertarget{edge-level-visualizations.}{%
\subsection{Edge-Level Visualizations.}\label{edge-level-visualizations.}}

Analyses which divide the brain into discrete regions and examine pair-wise dependencies between regional phenotypes are becoming increasingly common. Often the results of these dependency analyses are statistics relating to `edges' or links between any two regions. The package \texttt{brainconn} allows users to plot brain regions as nodes of a network graph, and dependencies between regions as edges of that graph, following a standardized coordinate space in both 2D and interactive 3D (Figure 1I). The graphs can be weighted, with the strength of the edges represented with color or line thickness, as well as directed, where arrows will be used in place of edges. This package comes with node coordinates corresponding to multiple commonly used atlases and contains functionality for users to easily enter their own custom atlas coordinates.

\hypertarget{limitations-and-functionality-gaps}{%
\section{Limitations and Functionality Gaps}\label{limitations-and-functionality-gaps}}

Most of the tools introduced above do not require a strong knowledge of programming, but there is still a steeper learning curve when compared to use a GUI. This is especially true when learning how to make fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning, for the purpose of a publication-ready figure. While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability. Relatedly, while some interactive image viewers which can be opened within an integrated development environment (e.g.~Muschelli, 2016), for quick and interactive viewing of single images, GUI tool can be faster and more practical.

Often cerebellar and brain-stem regions are not well represented in software (e.g.~Figure 1), potentially mirroring the cortico-centric sentiment that has prevailed in human neuroimaging research (Chin, Chang, \& Holmes, 2022). Likewise, custom non-cortical atlases such as non-standard subcortical atlas schemes are not yet straight forward, and usually require multiple functions and packages to visualize. Finally, some neuroimaging related data types, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools.

As can be seen in Table 1, there are usually multiple packages within each programming environment which are capable of visualizing each data type. While this provide choice for advanced users, it can also lead to confusion for novice users who may not be familiar with the nuanced differences between tools. Future work should continue to consolidate brain visualization methods into unified beginner-friendly code-based tools which are capable of plotting multiple data types.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

Poldrack, R. A., Baker, C. I., Durnez, J., Gorgolewski, K. J., Matthews, P. M., Munafò, M. R., \ldots{} \& Yarkoni, T. (2017). Scanning the horizon: towards transparent and reproducible neuroimaging research. Nature reviews neuroscience, 18(2), 115-126.

Steel, G. (2013, September 3). Publishing research without data is simply advertising, not science. Open Knowledge Foundation. \url{https://blog.okfn.org/2013/09/03/publishing-research-without-data-is-simply-advertising-not-science/}

The Comprehensive R Archive Network. (n.d.). Retrieved April 21, 2022, from \url{https://cran.r-project.org/}

Muschelli, J., Gherman, A., Fortin, J. P., Avants, B., Whitcher, B., Clayden, J. D., \ldots{} \& Crainiceanu, C. M. (2019). Neuroconductor: an R platform for medical imaging analysis. \emph{Biostatistics}, \emph{20}(2), 218-239.

Pham, D., Muschelli, J., \& Mejia, A. (2022). ciftiTools: A package for reading, writing, visualizing, and manipulating CIFTI files in R. \emph{NeuroImage}, 118877.

Mowinckel, A. M., \& Vidal-Piñeiro, D. (2020). Visualization of brain statistics with R packages ggseg and ggseg3d. \emph{Advances in Methods and Practices in Psychological Science}, \emph{3}(4), 466-483.

Huntenburg, J., Abraham, A., Loula, J., Liem, F., Dadi, K., \& Varoquaux, G. (2017). Loading and plotting of cortical surface representations in Nilearn. \emph{Research Ideas and Outcomes}, \emph{3}, e12342.

Wickham H. (2016). \emph{ggplot2: Elegant Graphics for Data Analysis.} Springer-Verlag New York. ISBN 978-3-319-24277-4, \url{https://ggplot2.tidyverse.org}.

Muschelli, J. (2016). \emph{papayar}. GitHub repository. \url{https://github.com/muschellij2/papayar}

David C. Van Essen, Stephen M. Smith, Deanna M. Barch, Timothy E.J. Behrens, Essa Yacoub, Kamil Ugurbil, for the WU-Minn HCP Consortium. (2013). The WU-Minn Human Connectome Project: An overview. NeuroImage 80(2013):62-79.

Sudlow, C., Gallacher, J., Allen, N., Beral, V., Burton, P., Danesh, J., \ldots{} \& Collins, R. (2015). UK biobank: an open access resource for identifying the causes of a wide range of complex diseases of middle and old age. PLoS medicine, 12(3), e1001779.

Esteban, O., Markiewicz, C. J., Blair, R. W., Moodie, C. A., Isik, A. I., Erramuzpe, A., \ldots{} \& Gorgolewski, K. J. (2019). fMRIPrep: a robust preprocessing pipeline for functional MRI. Nature methods, 16(1), 111-116.

Pernet, C., \& Madan, C. R. (2019). Data visualization for inference in tomographic brain imaging.

Code Ocean. (2021, October 27). Computational Research Platform on. Retrieved April 21, 2022, from \url{https://codeocean.com/}

Madan, C. R. (2015). Creating 3D visualizations of MRI data: A brief guide. F1000Research, 4.

Baker M. (2015). Social media: A network boost. Nature 518: 263--265. 10.1038/nj7538-263a

Lee J.-S. (2019). How to use Twitter to further your research career. Nature. 10.1038/d41586-019-00535-w

Li, Y., \& Xie, Y. (2020). Is a picture worth a thousand words? An empirical study of image content and social media engagement. Journal of Marketing Research, 57(1), 1-19.

Schäfer, T., \& Ecker, C. (2020). fsbrain: an R package for the visualization of structural neuroimaging data. bioRxiv.

Chin, R., Chang, S. W., \& Holmes, A. J. (2022). Beyond cortex: The evolution of the human brain. \emph{Psychological Review}.

\newpage

\hypertarget{supplement-code-used-to-generate-figures}{%
\section{Supplement (Code used to generate Figures)}\label{supplement-code-used-to-generate-figures}}

\singlespacing

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Figure 1A (Voxel)}

\FunctionTok{library}\NormalTok{(neurobase)}

\CommentTok{\#Load in a nifti (.nii.gz) file of a standardized template so use as a background. }
\NormalTok{template }\OtherTok{\textless{}{-}} \FunctionTok{readnii}\NormalTok{(}\StringTok{\textquotesingle{}data/MNI152\_T1\_1mm\_brain.nii.gz\textquotesingle{}}\NormalTok{) }\CommentTok{\#load the nifti file into R}

\CommentTok{\#Load in a nifti (.nii.gz) file of the statistic you wanted of overlay, with the same dimensions as the template.  }
\NormalTok{effect }\OtherTok{\textless{}{-}} \FunctionTok{readnii}\NormalTok{(}\StringTok{"data/MNI152\_effect\_size.nii.gz"}\NormalTok{) }

\CommentTok{\#Threshold the statistic map so only values about a specified amount are displayed}
\NormalTok{effect[effect}\SpecialCharTok{\textless{}}\FloatTok{0.3}\NormalTok{] }\OtherTok{\textless{}{-}} \ConstantTok{NA}

\CommentTok{\#Set the breaks/intervals you want on the color bar (e.g. from .3 to .6, by intervals of 0.005)}
\NormalTok{breaks }\OtherTok{=} \FunctionTok{seq}\NormalTok{(.}\DecValTok{3}\NormalTok{,.}\DecValTok{6}\NormalTok{, }\AttributeTok{by=}\FloatTok{0.005}\NormalTok{)}

\CommentTok{\#[Optional] If you want to output the figure as a .png, open a png image device}
\FunctionTok{png}\NormalTok{(}\StringTok{"data/fig1a.png"}\NormalTok{) }

\CommentTok{\#Use the ortho2 function from the neurobase package to make the figure}
\FunctionTok{ortho2}\NormalTok{(}\AttributeTok{x =}\NormalTok{ template,   }\CommentTok{\#Specify the background template}
       \AttributeTok{y =}\NormalTok{ effect,     }\CommentTok{\#Specify the effect you want to overlay on the template}
       \AttributeTok{crosshairs =}\NormalTok{ F, }\CommentTok{\#Remove the cross{-}hairs}
       \AttributeTok{bg =} \StringTok{"white"}\NormalTok{,   }\CommentTok{\#Make the background white }
       \AttributeTok{NA.x =}\NormalTok{ T,       }\CommentTok{\#Do not display NA values}
       \AttributeTok{col.y=} \FunctionTok{rev}\NormalTok{(viridis}\SpecialCharTok{::}\FunctionTok{inferno}\NormalTok{(}\AttributeTok{n =} \DecValTok{500}\NormalTok{)), }\CommentTok{\#Select the color scale. In this case I have used the inferno scale from the viridis package. }
       \AttributeTok{xyz =} \FunctionTok{c}\NormalTok{(}\DecValTok{70}\NormalTok{,}\DecValTok{50}\NormalTok{,}\DecValTok{80}\NormalTok{), }\CommentTok{\#set the x y \& z slice you want to visualize}
       \AttributeTok{useRaster =}\NormalTok{ T,     }\CommentTok{\#Sometimes using Raster makes for clearer plots}
       \AttributeTok{ycolorbar =} \ConstantTok{TRUE}\NormalTok{,  }\CommentTok{\#Add a color bar}
       \AttributeTok{mfrow =} \FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{)) }\SpecialCharTok{+}  \CommentTok{\#layout of brain views (for this figure, we just want a axial view)}
  \FunctionTok{colorbar}\NormalTok{(breaks, }\AttributeTok{col =} \FunctionTok{rev}\NormalTok{(viridis}\SpecialCharTok{::}\FunctionTok{inferno}\NormalTok{(}\AttributeTok{n=}\FunctionTok{length}\NormalTok{(breaks)}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{)), }\CommentTok{\#add a color bar with same color scale }
           \AttributeTok{text.col =} \StringTok{"black"}\NormalTok{, }\AttributeTok{labels =} \ConstantTok{TRUE}\NormalTok{, }\AttributeTok{maxleft =} \FloatTok{0.95}\NormalTok{)  }\CommentTok{\#Set the specifications for the color bar}

\FunctionTok{dev.off}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Figure 1B (Voxel)}

\FunctionTok{library}\NormalTok{(neurobase)}
\FunctionTok{library}\NormalTok{(scales)}

\CommentTok{\#Load in a nifti (.nii.gz) file to use as a background. }
\NormalTok{t1 }\OtherTok{\textless{}{-}} \FunctionTok{readnii}\NormalTok{(}\StringTok{\textquotesingle{}data/sub{-}001\_\_t1\_warped.nii.gz\textquotesingle{}}\NormalTok{) }

\CommentTok{\#Load in a nifti (.nii.gz) file to use as a overlay. }
\NormalTok{atlas }\OtherTok{\textless{}{-}}  \FunctionTok{readnii}\NormalTok{(}\StringTok{\textquotesingle{}data/sub{-}001\_schaefer300n7\_aseg\_to\_dwispace\_gm\_rois.nii.gz\textquotesingle{}}\NormalTok{)}

\FunctionTok{png}\NormalTok{(}\StringTok{\textquotesingle{}data/fig1b.png\textquotesingle{}}\NormalTok{)}
\FunctionTok{overlay}\NormalTok{(}\AttributeTok{x=}\FunctionTok{robust\_window}\NormalTok{(t1), }\AttributeTok{y=}\NormalTok{atlas, }
        \AttributeTok{plot.type =} \StringTok{"single"}\NormalTok{, }
        \AttributeTok{z=}\FunctionTok{c}\NormalTok{(}\FunctionTok{seq}\NormalTok{(}\DecValTok{30}\NormalTok{,}\DecValTok{110}\NormalTok{,}\DecValTok{10}\NormalTok{)), }
        \AttributeTok{col.y =}  \FunctionTok{alpha}\NormalTok{(viridis}\SpecialCharTok{::}\FunctionTok{inferno}\NormalTok{(}\DecValTok{300}\NormalTok{), }\FloatTok{0.6}\NormalTok{),}
        \AttributeTok{plane =} \StringTok{"axial"}\NormalTok{,  }
        \AttributeTok{useRaster =}\NormalTok{ T, }
        \AttributeTok{bg =} \StringTok{"white"}\NormalTok{,}
        \AttributeTok{NA.x=}\NormalTok{T, }
        \AttributeTok{zlim.y =}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{300}\NormalTok{))}
\FunctionTok{dev.off}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#vertex 1}
\CommentTok{\#Figure 1C}
\FunctionTok{library}\NormalTok{(rgl)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\CommentTok{\# Load the package and point to the Connectome Workbench}
\FunctionTok{library}\NormalTok{(ciftiTools)}

\CommentTok{\#Need connectome workbench to use ciftiTools {-} set path below}
\FunctionTok{ciftiTools.setOption}\NormalTok{(}\StringTok{"wb\_path"}\NormalTok{, }\StringTok{"/Applications/workbench/"}\NormalTok{)}

\CommentTok{\# Read in CIFTI file}
\NormalTok{cifti\_fname }\OtherTok{\textless{}{-}}\NormalTok{ ciftiTools}\SpecialCharTok{::}\FunctionTok{ciftiTools.files}\NormalTok{()}\SpecialCharTok{$}\NormalTok{cifti[}\StringTok{"dtseries"}\NormalTok{]}
\NormalTok{surfL\_fname }\OtherTok{\textless{}{-}} \FunctionTok{ciftiTools.files}\NormalTok{()}\SpecialCharTok{$}\NormalTok{surf[}\StringTok{"left"}\NormalTok{]}
\NormalTok{surfR\_fname }\OtherTok{\textless{}{-}} \FunctionTok{ciftiTools.files}\NormalTok{()}\SpecialCharTok{$}\NormalTok{surf[}\StringTok{"right"}\NormalTok{]}

\NormalTok{cii }\OtherTok{\textless{}{-}} \FunctionTok{read\_cifti}\NormalTok{(}
\NormalTok{  cifti\_fname, }\AttributeTok{brainstructures=}\StringTok{"all"}\NormalTok{, }
  \AttributeTok{surfL\_fname=}\NormalTok{surfL\_fname,}
  \AttributeTok{surfR\_fname=}\NormalTok{surfR\_fname)}

\CommentTok{\#Plot cifti surface}
\FunctionTok{view\_xifti\_surface}\NormalTok{(cii, }
                   \AttributeTok{hemisphere =} \StringTok{"both"}\NormalTok{, }
                   \AttributeTok{view =} \StringTok{"lateral"}\NormalTok{, }
                   \AttributeTok{idx =} \DecValTok{1}\NormalTok{,}
                   \AttributeTok{colors =}\NormalTok{ viridis}\SpecialCharTok{::}\FunctionTok{plasma}\NormalTok{(}\AttributeTok{n =} \DecValTok{100}\NormalTok{),}
                   \AttributeTok{zlim =} \FunctionTok{c}\NormalTok{(}\FloatTok{1.2}\NormalTok{,}\FloatTok{1.5}\NormalTok{),  }
                   \AttributeTok{legend\_embed =}\NormalTok{ T,}
                   \AttributeTok{cex.title =} \DecValTok{2}\NormalTok{)}

\FunctionTok{rgl.snapshot}\NormalTok{(}\StringTok{"data/fig1c.png"}\NormalTok{)}
\FunctionTok{rgl.close}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Vertex 2}
\CommentTok{\#Figure 1X {-} Remove white space}

\FunctionTok{library}\NormalTok{(fsbrain)}
\FunctionTok{library}\NormalTok{(rgl)}

\CommentTok{\#download\_optional\_data()}
\CommentTok{\#download\_fsaverage(accept\_freesurfer\_license = TRUE)}

\NormalTok{subjects\_dir }\OtherTok{\textless{}{-}} \FunctionTok{get\_optional\_data\_filepath}\NormalTok{(}\StringTok{"subjects\_dir"}\NormalTok{)}

\NormalTok{subject\_id }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}subject1\textquotesingle{}}

\NormalTok{colourmap }\OtherTok{\textless{}{-}} \FunctionTok{colorRampPalette}\NormalTok{(viridis}\SpecialCharTok{::}\FunctionTok{plasma}\NormalTok{(}\AttributeTok{n =} \DecValTok{1000}\NormalTok{))}
\FunctionTok{fsbrain.set.default.figsize}\NormalTok{(}\DecValTok{700}\NormalTok{, }\DecValTok{700}\NormalTok{);}

\NormalTok{brain }\OtherTok{=} \FunctionTok{vis.subject.morph.native}\NormalTok{(subjects\_dir, }
\NormalTok{                                subject\_id, }\StringTok{\textquotesingle{}sulc\textquotesingle{}}\NormalTok{, }
                                \AttributeTok{cortex\_only =}\NormalTok{ T, }
                                \AttributeTok{views=}\ConstantTok{NULL}\NormalTok{, }
                                \AttributeTok{draw\_colorbar =} \ConstantTok{FALSE}\NormalTok{,}
                                \AttributeTok{makecmap\_options =} \FunctionTok{list}\NormalTok{(}\StringTok{\textquotesingle{}colFn\textquotesingle{}}\OtherTok{=}\NormalTok{colourmap))}

\NormalTok{img }\OtherTok{=} \FunctionTok{export}\NormalTok{(brain, }\AttributeTok{draw\_colorbar =} \ConstantTok{FALSE}\NormalTok{, }\AttributeTok{output\_img =} \StringTok{"data/fig1d.png"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#ROI1 }
\CommentTok{\#Figure 1e }
\FunctionTok{library}\NormalTok{(ciftiTools)}
\NormalTok{parc }\OtherTok{\textless{}{-}} \FunctionTok{load\_parc}\NormalTok{(}\StringTok{"Yeo\_7"}\NormalTok{)}
\FunctionTok{view\_xifti\_surface}\NormalTok{(parc,  }
                   \AttributeTok{legend\_embed =}\NormalTok{ T,}
                   \AttributeTok{hemisphere =} \StringTok{"right"}\NormalTok{)}

\FunctionTok{rgl.snapshot}\NormalTok{(}\StringTok{"data/fig1e.png"}\NormalTok{)}

\FunctionTok{rgl.close}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#ROI1 }
\CommentTok{\#Figure 1X {-} combine with previous figure}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1993}\NormalTok{) }\CommentTok{\#set a random seed (good practice for reproducibility)}

\NormalTok{parc }\OtherTok{\textless{}{-}} \FunctionTok{load\_parc}\NormalTok{(}\StringTok{"Schaefer\_400"}\NormalTok{)}

\NormalTok{ramdom\_metric }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\DecValTok{400}\NormalTok{)}

\NormalTok{cii }\OtherTok{\textless{}{-}} \FunctionTok{move\_from\_mwall}\NormalTok{(cii, }\ConstantTok{NA}\NormalTok{)}

\NormalTok{parc\_vec }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{as.matrix}\NormalTok{(parc))}

\NormalTok{xii\_metric }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\ConstantTok{NA}\NormalTok{, ramdom\_metric)[parc\_vec }\SpecialCharTok{+} \DecValTok{1}\NormalTok{]}

\NormalTok{xii1 }\OtherTok{\textless{}{-}} \FunctionTok{select\_xifti}\NormalTok{(cii, }\DecValTok{1}\NormalTok{)}

\NormalTok{xii\_metric }\OtherTok{\textless{}{-}} \FunctionTok{newdata\_xifti}\NormalTok{(xii1, xii\_metric)}

\NormalTok{plot2 }\OtherTok{\textless{}{-}} \FunctionTok{view\_xifti\_surface}\NormalTok{(xii\_metric, }
                            \AttributeTok{colors =}\NormalTok{ viridis}\SpecialCharTok{::}\FunctionTok{viridis}\NormalTok{(}\AttributeTok{n =} \DecValTok{400}\NormalTok{), }
                            \AttributeTok{borders =} \StringTok{"black"}\NormalTok{, }
                            \AttributeTok{hemisphere =} \StringTok{"both"}\NormalTok{,}
                            \AttributeTok{view =} \StringTok{\textquotesingle{}lateral\textquotesingle{}}\NormalTok{)}

\FunctionTok{rgl.snapshot}\NormalTok{(}\StringTok{"data/fig1f.png"}\NormalTok{)}

\FunctionTok{rgl.close}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#ROI2}

\CommentTok{\#devtools::install\_github("LCBC{-}UiO/ggsegGlasser")}
\FunctionTok{library}\NormalTok{(ggseg)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{library}\NormalTok{(ggsegGlasser)}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1993}\NormalTok{) }\CommentTok{\#set a random seed (good practice for reproducibility)}

\NormalTok{base\_atlas }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(}\FunctionTok{na.omit}\NormalTok{(}\FunctionTok{cbind}\NormalTok{(glasser}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{region, glasser}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{hemi)))}

\FunctionTok{colnames}\NormalTok{(base\_atlas) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"region"}\NormalTok{, }\StringTok{"hemi"}\NormalTok{)}

\NormalTok{Effect }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{dim}\NormalTok{(base\_atlas)[}\DecValTok{1}\NormalTok{]) }\CommentTok{\#generate a random numbers for each roi in the atlas}

\NormalTok{base\_atlas }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(Effect, base\_atlas)}

\NormalTok{cortex }\OtherTok{\textless{}{-}} \FunctionTok{ggseg}\NormalTok{(}\AttributeTok{atlas =}\NormalTok{ glasser,}
                \AttributeTok{.data =}\NormalTok{ base\_atlas,}
                \AttributeTok{mapping=}\FunctionTok{aes}\NormalTok{(}\AttributeTok{fill=}\NormalTok{Effect), }
                \AttributeTok{position=}\StringTok{"stacked"}\NormalTok{,}
                \AttributeTok{colour=}\StringTok{"black"}\NormalTok{,}
                \AttributeTok{size=}\NormalTok{.}\DecValTok{2}\NormalTok{,}
                \AttributeTok{show.legend =}\NormalTok{ T,}
                \AttributeTok{plot.background =} \StringTok{"white"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_fill\_viridis\_b}\NormalTok{() }\SpecialCharTok{+} 
   \FunctionTok{theme\_void}\NormalTok{() }\SpecialCharTok{+}
    \FunctionTok{theme}\NormalTok{(}\AttributeTok{legend.position =} \StringTok{"bottom"}\NormalTok{)}

\NormalTok{sub\_base\_atlas }\OtherTok{\textless{}{-}} \FunctionTok{as.data.frame}\NormalTok{(}\FunctionTok{na.omit}\NormalTok{(}\FunctionTok{cbind}\NormalTok{(aseg}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{region, aseg}\SpecialCharTok{$}\NormalTok{data}\SpecialCharTok{$}\NormalTok{hemi)))}

\FunctionTok{colnames}\NormalTok{(sub\_base\_atlas) }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\StringTok{"region"}\NormalTok{, }\StringTok{"hemi"}\NormalTok{)}

\NormalTok{Effect }\OtherTok{\textless{}{-}} \FunctionTok{rnorm}\NormalTok{(}\FunctionTok{dim}\NormalTok{(sub\_base\_atlas)[}\DecValTok{1}\NormalTok{])}

\NormalTok{sub\_base\_atlas }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(Effect, sub\_base\_atlas)}

\NormalTok{subcortex }\OtherTok{\textless{}{-}} \FunctionTok{ggseg}\NormalTok{(}\AttributeTok{atlas =}\NormalTok{ aseg,}
                   \AttributeTok{.data =}\NormalTok{ sub\_base\_atlas,}
                   \AttributeTok{mapping=}\FunctionTok{aes}\NormalTok{(}\AttributeTok{fill=}\NormalTok{Effect), }
                   \AttributeTok{position =} \StringTok{"dispersed"}\NormalTok{,}
                   \AttributeTok{colour=}\StringTok{"black"}\NormalTok{,}
                   \AttributeTok{hemi =}\FunctionTok{c}\NormalTok{(}\StringTok{\textquotesingle{}left\textquotesingle{}}\NormalTok{, }\StringTok{\textquotesingle{}right\textquotesingle{}}\NormalTok{),}
                   \AttributeTok{size=}\NormalTok{.}\DecValTok{2}\NormalTok{,}
                   \AttributeTok{show.legend =}\NormalTok{ T, }
                   \AttributeTok{plot.background =} \StringTok{"white"}\NormalTok{) }\SpecialCharTok{+} 
  \FunctionTok{scale\_fill\_viridis\_b}\NormalTok{() }

\FunctionTok{ggsave}\NormalTok{(}\AttributeTok{filename =} \StringTok{\textquotesingle{}data/fig1g1.png\textquotesingle{}}\NormalTok{, }\AttributeTok{plot =}\NormalTok{ cortex, }\AttributeTok{device =} \StringTok{\textquotesingle{}png\textquotesingle{}}\NormalTok{, }\AttributeTok{bg =} \StringTok{\textquotesingle{}white\textquotesingle{}}\NormalTok{)}
\FunctionTok{ggsave}\NormalTok{(}\AttributeTok{filename =} \StringTok{\textquotesingle{}data/fig1g2.png\textquotesingle{}}\NormalTok{, }\AttributeTok{plot =}\NormalTok{ subcortex, }\AttributeTok{device =} \StringTok{\textquotesingle{}png\textquotesingle{}}\NormalTok{, }\AttributeTok{bg =} \StringTok{\textquotesingle{}white\textquotesingle{}}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{make\_ggseg3d }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(attribute, }
                         \AttributeTok{colour.pal =} \FunctionTok{c}\NormalTok{(}\StringTok{"light yellow"}\NormalTok{,}
                                        \StringTok{"orange"}\NormalTok{,}
                                        \StringTok{"red"}\NormalTok{,}
                                        \StringTok{"dark red"}\NormalTok{), }
                         \AttributeTok{hide.colourbar=}\ConstantTok{FALSE}\NormalTok{, }
                         \AttributeTok{output.png=}\ConstantTok{FALSE}\NormalTok{, }
                         \AttributeTok{file.name=}\StringTok{"ggseg\_3d.png"}\NormalTok{) \{}
  \CommentTok{\# Inputs: }
  \CommentTok{\#\textquotesingle{} attribute = This must be a numeric vector in this: Left{-}Thalamus, Left{-}Caudate, Left{-}Putamen, Left{-}Pallidum Left{-}Hippocampus, Left{-}Amygdala, Left{-}Accumbens{-}area, Right{-}Thalamus{-}Proper, Right{-}Caudate, Right{-}Putamen, Right{-}Pallidum, Right{-}Hippocampus, Right{-}Amygdala, Right{-}Accumbens{-}area}

  \FunctionTok{library}\NormalTok{(ggseg3d)}
  \FunctionTok{library}\NormalTok{(tidyr)}
  \FunctionTok{library}\NormalTok{(dplyr)}
  \CommentTok{\#remove(aseg\_3d)}
\NormalTok{  aseg\_3d }\OtherTok{\textless{}{-}}\NormalTok{ aseg\_3d}
\NormalTok{  aseg\_3d }\OtherTok{\textless{}{-}}\NormalTok{ tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(aseg\_3d, }\AttributeTok{cols =} \FunctionTok{c}\NormalTok{(ggseg\_3d))}

\NormalTok{  attribute.ggseg3d }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\FunctionTok{rep}\NormalTok{(}\ConstantTok{NA}\NormalTok{, }\DecValTok{4}\NormalTok{), attribute[}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{], }\FunctionTok{rep}\NormalTok{(}\ConstantTok{NA}\NormalTok{, }\DecValTok{3}\NormalTok{), attribute[}\DecValTok{5}\SpecialCharTok{:}\DecValTok{7}\NormalTok{],}
                         \FunctionTok{rep}\NormalTok{(}\ConstantTok{NA}\NormalTok{, }\DecValTok{5}\NormalTok{), attribute[}\DecValTok{8}\SpecialCharTok{:}\DecValTok{14}\NormalTok{],  }\FunctionTok{rep}\NormalTok{(}\ConstantTok{NA}\NormalTok{, }\DecValTok{6}\NormalTok{))}

\NormalTok{  data }\OtherTok{\textless{}{-}}\NormalTok{ dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(aseg\_3d, }\AttributeTok{attribute =}\NormalTok{  attribute.ggseg3d)}
  \CommentTok{\#remove NA regions}
\NormalTok{  aseg\_3d[}\FunctionTok{which}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(data}\SpecialCharTok{$}\NormalTok{attribute)),] }\OtherTok{\textless{}{-}} \ConstantTok{NA}
\NormalTok{  aseg\_3d }\OtherTok{\textless{}{-}}\NormalTok{ tidyr}\SpecialCharTok{::}\FunctionTok{drop\_na}\NormalTok{(aseg\_3d)}

\NormalTok{  data[}\FunctionTok{which}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(data}\SpecialCharTok{$}\NormalTok{attribute)),] }\OtherTok{\textless{}{-}} \ConstantTok{NA}
\NormalTok{  data }\OtherTok{\textless{}{-}}\NormalTok{ tidyr}\SpecialCharTok{::}\FunctionTok{drop\_na}\NormalTok{(data)}
\NormalTok{  data}\SpecialCharTok{$}\NormalTok{attribute[data}\SpecialCharTok{$}\NormalTok{attribute}\SpecialCharTok{==}\DecValTok{0}\NormalTok{]}\OtherTok{\textless{}{-}}\ConstantTok{NA} \CommentTok{\#make 0 valus NA so they are set as grey in ggseg3d}

\NormalTok{  scene}\OtherTok{=}\FunctionTok{list}\NormalTok{(}\AttributeTok{camera =} \FunctionTok{list}\NormalTok{(}\AttributeTok{eye =} \FunctionTok{list}\NormalTok{(}\AttributeTok{x =} \DecValTok{0}\NormalTok{, }\AttributeTok{y =} \DecValTok{1}\NormalTok{, }\AttributeTok{z =} \SpecialCharTok{{-}}\FloatTok{2.25}\NormalTok{)),}
             \AttributeTok{aspectratio =} \FunctionTok{list}\NormalTok{(}\AttributeTok{x=}\FloatTok{1.6}\NormalTok{,}\AttributeTok{y=}\FloatTok{1.6}\NormalTok{,}\AttributeTok{z=}\FloatTok{1.6}\NormalTok{))}

\NormalTok{  plot }\OtherTok{\textless{}{-}}\NormalTok{ ggseg3d}\SpecialCharTok{::}\FunctionTok{ggseg3d}\NormalTok{(}\AttributeTok{.data =}\NormalTok{ data, }
                           \AttributeTok{atlas =}\NormalTok{ aseg\_3d, }
                           \AttributeTok{colour =} \StringTok{"attribute"}\NormalTok{, }
                           \AttributeTok{text =} \StringTok{"attribute"}\NormalTok{,}
                           \AttributeTok{palette =}\NormalTok{ colour.pal)}
\NormalTok{  plot }\OtherTok{\textless{}{-}} \FunctionTok{remove\_axes}\NormalTok{(plot)}
\NormalTok{  plot }\OtherTok{\textless{}{-}}\NormalTok{ plotly}\SpecialCharTok{::}\FunctionTok{layout}\NormalTok{(plot, }
                         \AttributeTok{scene =}\NormalTok{ scene, }
                         \AttributeTok{width =} \DecValTok{600}\NormalTok{, }\AttributeTok{height =} \DecValTok{600}\NormalTok{) }

  \ControlFlowTok{if}\NormalTok{(hide.colourbar}\SpecialCharTok{==}\ConstantTok{TRUE}\NormalTok{)\{}
\NormalTok{    plot }\OtherTok{\textless{}{-}}\NormalTok{ plotly}\SpecialCharTok{::}\FunctionTok{hide\_colorbar}\NormalTok{(plot)}
\NormalTok{  \}}

  \ControlFlowTok{if}\NormalTok{( output.png}\SpecialCharTok{==}\ConstantTok{TRUE}\NormalTok{)\{}
\NormalTok{    plotly}\SpecialCharTok{::}\FunctionTok{orca}\NormalTok{(plot, }\AttributeTok{file =}\NormalTok{  file.name)}
\NormalTok{  \}}

  \FunctionTok{return}\NormalTok{(plot)}

\NormalTok{\}}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1993}\NormalTok{)}
\NormalTok{volume }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{20}\NormalTok{, }\DecValTok{14}\NormalTok{) }\CommentTok{\#randomly generated data for 14 structures}
\FunctionTok{make\_ggseg3d}\NormalTok{(volume, }
             \AttributeTok{colour.pal =}\NormalTok{ viridis}\SpecialCharTok{::}\FunctionTok{viridis}\NormalTok{(}\AttributeTok{n=}\FunctionTok{length}\NormalTok{(volume)), }
             \AttributeTok{output.png =}\NormalTok{ T, }\AttributeTok{file.name =} \StringTok{"data/fig1g3.png"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#Edge (also add 3d version)}
\CommentTok{\#Fig1X}

\CommentTok{\#devtools::install\_github("sidchop/brainconn")}
\FunctionTok{library}\NormalTok{(brainconn)}

\NormalTok{conmat }\OtherTok{\textless{}{-}}\NormalTok{ example\_weighted\_undirected}
\NormalTok{degree }\OtherTok{\textless{}{-}} \FunctionTok{rowSums}\NormalTok{(conmat)[}\FunctionTok{which}\NormalTok{(}\FunctionTok{rowSums}\NormalTok{(conmat)}\SpecialCharTok{!=}\DecValTok{0}\NormalTok{)]}

\FunctionTok{brainconn}\NormalTok{(}\AttributeTok{atlas =}\StringTok{"schaefer300\_n7"}\NormalTok{, }
                   \AttributeTok{conmat=}\NormalTok{conmat, }
                   \AttributeTok{node.size =}\NormalTok{ degree}\SpecialCharTok{/}\DecValTok{2}\NormalTok{,}
                   \AttributeTok{view=}\StringTok{"top"}\NormalTok{, }
                   \AttributeTok{edge.color.weighted =}\NormalTok{ T,}
                   \AttributeTok{background.alpha =} \FloatTok{0.8}\NormalTok{, }
                   \AttributeTok{show.legend =}\NormalTok{ F, }
                   \AttributeTok{edge.color =} \FunctionTok{scale\_edge\_colour\_viridis}\NormalTok{()) }

\FunctionTok{ggsave}\NormalTok{(}\AttributeTok{filename =} \StringTok{\textquotesingle{}data/fig1h.png\textquotesingle{}}\NormalTok{, }
       \AttributeTok{plot =}\NormalTok{ fig1h, }
       \AttributeTok{device =} \StringTok{\textquotesingle{}png\textquotesingle{}}\NormalTok{,}
       \AttributeTok{bg =} \StringTok{\textquotesingle{}white\textquotesingle{}}\NormalTok{,}
       \AttributeTok{height =} \DecValTok{7}\NormalTok{, }
       \AttributeTok{units =} \StringTok{"in"}\NormalTok{)}

\FunctionTok{brainconn3D}\NormalTok{(}\AttributeTok{atlas =}\StringTok{"schaefer300\_n7"}\NormalTok{, }
            \AttributeTok{conmat=}\NormalTok{conmat,}
            \AttributeTok{show.legend =}\NormalTok{ F)}
\end{Highlighting}
\end{Shaded}

\bibliographystyle{unsrt}
\bibliography{}

\end{document}

### 2022-08-15 manuscript.Rmd
c++ compline edit
python fig draft added

@@
 Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data.
+
 output: 
@@
 By using code-based tools available within your preferred environment, brain visualizations can be directly integrated and embedded within a paper or report.
-For instance, a fully reproducible version of the current paper can be found on [GitHub](https://github.com/sidchop/Brain_comms_field_potential).
+For instance, a fully reproducible version of the current paper can be found on [GitHub](https://github.com/sidchop/RepoNeuroVis).
 Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g.
@@
 |                                                        | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
-|:-----------------------|:--------|:--------|:--------------|:--------|:------------|
+|:-----------|:-----------|:-----------|:-----------|:-----------|:-----------|
 | **R**                                                  |         |         |               |         |             |
-| ANTsR                  | ***+*** | ***+*** | ***+***       |         |             |
-| brainconn              |         |         |               | ***+*** |             |
-| brainR                 | ***+*** |         | ***+***       |         |             |
-| ciftitools             | ***+*** | ***+*** | ***+*****\*** |         |             |
-| fsbrain                | ***+*** | ***+*** | ***+*****\*** |         |             |
+| [ANTsR](https://github.com/ANTsX/ANTsR)                | ***+*** | ***+*** | ***+***       |         |             |
+| [brainconn](https://github.com/sidchop/brainconn)      |         |         |               | ***+*** |             |
+| [brainR](https://github.com/muschellij2/brainR)        | ***+*** |         | ***+***       |         |             |
+| [ciftitools](https://github.com/mandymejia/ciftiTools) | ***+*** | ***+*** | ***+*****\*** |         |             |
+| [fsbrain](https://github.com/dfsp-spirit/fsbrain)      | ***+*** | ***+*** | ***+*****\*** |         |             |
 | ggseg                                                  |         |         | ***+***       |         |             |
@@
 
-# Examples of Neuroimaging Visualization Packages available in *R*
+# Examples of Neuroimaging Visualization Packages available in *R and Python*
 
-The following four sections provide brief examples of packages and functions available in R for visualizing voxel, vertex, ROI and edge-level data.
-A didactic introduction to these tools is provided in an accompanying [collection of beginner friendly R-Notebooks](https://github.com/sidchop/neuRo-vis-BrainHack-2021).
-This is not an exhaustive list of packages available for visualizing brain data in R, rather, the following sections and accompanying notebooks aim to give the reader a sense of the available options, and an entry point to using them.
+The following four sections provide brief examples of packages and functions available in R and Python for visualizing voxel, vertex, ROI and edge-level data.
+This is not an exhaustive list of packages available for visualizing brain data in R and Python, rather, the following sections aim to give the reader a sense of the available options, and an entry point to using them.
+A didactic introduction to the R based tools is provided in an accompanying [collection of beginner friendly R-Notebooks](https://github.com/sidchop/neuRo-vis-BrainHack-2021).
 
@@
 patchwork + plot_annotation(tag_levels = 'A',
-                            title = 'Examples of brain visualizations in R') &
+                            title = 'Examples of brain visualizations made in R') &
     theme(plot.tag = element_text(face = 'bold'),
@@
 *Figure 1.
-Examples of brain imaging visualization make in using different R-packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown.
+Examples of brain imaging visualization make in using different R packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown.
 Made using the the `ortho2` function from the `neurobase` package.
@@
 All code used to compile this figure, as well as the contents of each panel are provided in the Supplement.
+
+\elandscape
+
+```{=tex}
+\newpage
+\blandscape
+```
+```{r Figure2 , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center'}
+
+library(patchwork)
+library(ggplot2)
+
+fig_name <- c('fig1', 'fig2', 'fig3', 'fig4', 'fig5', 'fig6', 'fig7')
+figs <- list()
+
+for (f in 1:length(fig_name)) { 
+  knitr::plot_crop(paste0('data/python/',fig_name[f],'.png'))
+  figs[[f]] <- ggplot() +
+    annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/python/',fig_name[f],'.png')),
+                                                             width=ggplot2::unit(1,"npc"),
+                                                             height=ggplot2::unit(1,"npc")))
+}
+
+patchwork2 <- (figs[[1]]+figs[[3]])/figs[[2]] | (figs[[6]]/figs[[4]]/figs[[5]])
+
+  figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]])+plot_layout(heights = c(0.5,1,1)))
+
+patchwork + plot_annotation(tag_levels = 'A',
+                            title = 'Examples of brain visualizations made in Pythoin') &
+    theme(plot.tag = element_text(face = 'bold'),
+          plot.title = element_text(face="bold"))
+
+```
+
+*Figure 1. Examples of brain imaging visualization make in using different Python packages.* A)
 

### 2022-08-15 readme.Rmd
python fig draft added

---
title: "Reproducible and Programmatic Generation of Neuroimaging Visualizations"
output: github_document
---

## Reference

c.	Chopra, S., Labache, L., Dhamala, E., Orchard, E., Holmes. A. [Reproducible and programmatic generation of neuroimaging visualizatio](https://github.com/sidchop/RepoNeuroVis/blob/main/manuscript.pdf). Preprint.

------------------------------------------------------------------------

## Background

Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results. Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI). There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB. Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process. This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments. Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data..

------------------------------------------------------------------------

## Figures made in R: \

```{r, echo = FALSE,warning=F, error=F, out.width='100%'}
knitr::include_graphics("data/readme_figures/fig1.png")
```

## Figures made in Python: \

```{r, echo = FALSE,warning=F, error=F, out.width='100%'}
knitr::include_graphics("data/readme_figures/fig2.png")
```

------------------------------------------------------------------------

## Questions

Please contact me (Sidhant Chopra) as [sidhant.chopra\@yale.edu](mailto:sidhant.chopra@yale.edu){.email} and/or [sidhant.chopra4\@gmail.com](mailto:sidhant.chopra4@gmail.com){.email}

### 2022-08-15 readme.Rmd
rd up
tabel added to rdme

@@
 
-c.	Chopra, S., Labache, L., Dhamala, E., Orchard, E., Holmes.
+**Chopra, S.**, Labache, L., Dhamala, E., Orchard, E., Holmes.
 A.
@@
 
+## Examples of code-based neuroimaging visualizations tools that can be accessed directly within R, MATLAB and Python environments.
+
+|                                                        | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
+|:-----------|:-----------|:-----------|:-----------|:-----------|:-----------|
+| **R**                                                  |         |         |               |         |             |
+| [ANTsR](https://github.com/ANTsX/ANTsR)                | ***+*** | ***+*** | ***+***       |         |             |
+| [brainconn](https://github.com/sidchop/brainconn)      |         |         |               | ***+*** |             |
+| [brainR](https://github.com/muschellij2/brainR)        | ***+*** |         | ***+***       |         |             |
+| [ciftitools](https://github.com/mandymejia/ciftiTools) | ***+*** | ***+*** | ***+*****\*** |         |             |
+| [fsbrain](https://github.com/dfsp-spirit/fsbrain)      | ***+*** | ***+*** | ***+*****\*** |         |             |
+| ggseg                                                  |         |         | ***+***       |         |             |
+| neurobase                                              | ***+*** |         |               |         |             |
+| oro.nifti                                              | ***+*** |         |               |         |             |
+| **Python**                                             |         |         |               |         |             |
+| ANTsPy                                                 | ***+*** | ***+*** | ***+***       |         |             |
+| brainiak                                               | ***+*** |         |               |         |             |
+| Brainplotlib                                           |         | ***+*** | ***+*****\*** |         |             |
+| Brainspace/surfplot                                    |         | ***+*** | ***+*****\*** |         |             |
+| DIPY                                                   | ***+*** |         |               |         | ***+***     |
+| ENIGMA toolbox                                         |         |         | ***+***       |         |             |
+| FSLeyes                                                | ***+*** | ***+*** | ***+***       |         |             |
+| ggseg                                                  |         |         | ***+***       |         |             |
+| graphpype                                              |         |         |               | ***+*** |             |
+| MMVT                                                   |         | ***+*** | ***+***       | ***+*** |             |
+| MNE                                                    | ***+*** | ***+*** | ***+***       |         |             |
+| mrivis                                                 | ***+*** |         |               |         |             |
+| NaNSlice                                               | ***+*** |         |               |         |             |
+| netneurotools                                          |         | ***+*** | ***+*****\*** |         |             |
+| nilearn                                                | ***+*** | ***+*** | ***+***       | ***+*** |             |
+| niwidget                                               | ***+*** | ***+*** |               |         | ***+***     |
+| Pycortex                                               | ***+*** | ***+*** | ***+*****\*** |         |             |
+| pySurfer                                               |         | ***+*** | ***+*****\*** |         |             |
+| surfice                                                | ***+*** | ***+*** | ***+***       | ***+*** | ***+***     |
+| surfplot                                               |         | ***+*** | ***+***       |         |             |
+| Visbrain                                               | ***+*** | ***+*** | ***+***       | ***+*** |             |
+| **MATLAB**                                             |         |         |               |         |             |
+| BrainNetViewer                                         | ***+*** |         | ***+***       | ***+*** |             |
+| Brainspace                                             |         | ***+*** | ***+*****\*** |         |             |
+| Brainstorm                                             | ***+*** | ***+*** | ***+***       |         |             |
+| bspmview                                               | ***+*** |         | ***+***       |         |             |
+| CandlabCore                                            | ***+*** |         | ***+***       |         |             |
+| ECoG/fMRI Vis toolbox                                  |         | ***+*** | ***+*****\*** |         |             |
+| ENIGMA toolbox                                         |         |         | ***+***       |         |             |
+| FieldTrip                                              | ***+*** | ***+*** |               |         |             |
+| Lead-DBS                                               | ***+*** |         | ***+***       |         |             |
+| mni2fs                                                 |         | ***+*** |               |         |             |
+| mrtools                                                | ***+*** | ***+*** | ***+*****\*** |         |             |
+| plotSurfaceROIBoundary                                 |         | ***+*** | ***+*****\*** |         |             |
+| Vistasoft                                              | ***+*** | ***+*** | ***+***       |         | ***+***     |
+
+Note: The tools listed contain functionality required to generate (at least close-to) publication-ready neuroimaging figures via user-entered code within R, MATLAB and Python environments.
+This list does not include cross-platform general purpose visualization software.\
+\*Cortex only
+
+------------------------------------------------------------------------
+
 ## Questions

### 2022-08-27 manuscript.Rmd
py fig added

@@
 These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments.
+The present article argues for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
+We also providing didactic examples of visualizations and associated code as a reference point.
 
-The present article covers three major advantages of using code-generated visualizations over GUI based tools: replicability, flexibility, and integration.
-First, by generating visualizations using code, you increase the replicability of your figures for yourself, your collaborators and your readers.
-Second, code-based tools provides more precise controls over display settings, while also benefiting from the advantages of programming, such as the ability to iteratively and rapidly generate multiple figures.
+In brief, by generating visualizations using code, you increase the replicability of your figures for yourself, your collaborators and your readers.
+Moreover, code-based tools provides more precise controls over display settings, while also benefiting from the advantages of programming, such as the ability to iteratively and rapidly generate multiple figures.
 Finally, being able to integrate figure generation into your analysis scripts reduces chances of errors and increases the accessibility of resulting analyses pipeline, with more advanced tools now allowing for the seamless integration of prose and code.
@@
 | ENIGMA toolbox                                         |         |         | ***+***       |         |             |
-| FSLeyes                                                | ***+*** | ***+*** | ***+***       |         |             |
+| FSLeyes                                                | ***+*** | ***+*** | ***+***       |         | *+*         |
 | ggseg                                                  |         |         | ***+***       |         |             |
@@
 
-The following four sections provide brief examples of packages and functions available in R and Python for visualizing voxel, vertex, ROI and edge-level data.
-This is not an exhaustive list of packages available for visualizing brain data in R and Python, rather, the following sections aim to give the reader a sense of the available options, and an entry point to using them.
-A didactic introduction to the R based tools is provided in an accompanying [collection of beginner friendly R-Notebooks](https://github.com/sidchop/neuRo-vis-BrainHack-2021).
+The following four sections provide brief examples of well documented and beginner friendly packages and functions available in R and Python for visualizing voxel, vertex, ROI and edge-level data.
+This is not an exhaustive list of packages available for visualizing brain data in R and Python (See Table 1), rather, the following sections aim to give the reader a sense of the available options, and an entry point to using them.
+All code used to compile the figures, as well as the contents of each panel are provided in the Supplement or can be accessed in the accompanying Git repository.
 
@@
 ```
-```{r Figure2 , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center'}
+```{r Figure2 , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center', eval=FALSE}
 
@@
 patchwork + plot_annotation(tag_levels = 'A',
-                            title = 'Examples of brain visualizations made in Pythoin') &
+                            title = 'Examples of brain visualizations made in Python') &
     theme(plot.tag = element_text(face = 'bold'),
@@
 
-The `oro.nifti` and `neurobase` packages allow for the reading, writing, manipulation and visualization of voxel-level imaging data of either *nifti*, *analyze* or *afni* formats.
+**R.** The `oro.nifti` and `neurobase` packages allow for the reading, writing, manipulation and visualization of voxel-level imaging data of either *nifti*, *analyze* or *afni* formats.
 In particular, the `image`, `slice` and `ortho2` functions allow users to display the desired number of slices in the desired orientations, while also providing precise control over overlaid images, color scales, legend placement and other aesthetic settings (Fig 1A-B).
@@
 
+**Python.** The `NiBable` library provides the ability to read in and covert many common neuroimaging file formats, including *nifti,* to data-structures which can be manipulated and visualized.
+While these data-structures often form 3D or 4D matrices, so can be visualized using commonly used general purpose visualization tools such as `matplotlib` , purpose built librires like `nilearn` provide functions like `plot_stat_map` and `plot_glass_brain` which are user friendly and can generate publication ready figures (Fig.
+X).
+`nilearn` also includes smoothing and re-sampling functions which can sometimes required for visualization purposes.
+
 ## Vertex-Level Visualizations.
 
-The recently developed `fsbrain` package allows for the visualization of vertex-level and cortical ROI-level data which were derived using *FreeSurfer*.
+**R.** The recently developed `fsbrain` package allows for the visualization of vertex-level and cortical ROI-level data which were derived using *FreeSurfer*.
 The package contains many flexible functions for visualizing individual-level and group-level measures such as cortical thickness, volume or surface area (Fig 1F).
@@
 
+**Python.** ciftify.
+
 ## ROI-Level Visualizations.
 
-Statistical parameters can be mapped onto discreet cortical and sub-cortical regions using the `ggseg` and `ggseg3d` packages (Mowinckel & Vidal-Piñeiro, 2020).
+**R.** Statistical parameters can be mapped onto discreet cortical and sub-cortical regions using the `ggseg` and `ggseg3d` packages (Mowinckel & Vidal-Piñeiro, 2020).
 These packages flexibly generates aesthetic renderings of cortical, sub-cortical and white matter ROIs in 2D (Fig 1D,1F) and 3D (Fig 1E).
@@
 Notably, the `ciftiTools` package also provides functions to visualize surface and voxel-level ROIs for the cortex and sub-cortex respectively (Fig 1E), flexibly allowing CIFTI format atlases and ROIs to be displayed.
+
+**Python.**
 
@@
 This package comes with node coordinates corresponding to multiple commonly used atlases and contains functionality for users to easily enter their own custom atlas coordinates.
+
+**Python.**
 
@@
   # Inputs: 
-  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left-Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Right-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Right-Amygdala, Right-Accumbens-area
+  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left
+  #' -Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Righ
+  #' t-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Righ
+  #' t-Amygdala, Right-Accumbens-area
 

### 2022-08-29 manuscript.Rmd
py fig fix

@@
 ---
-title: Reproducible and Programmatic Generation of Neuroimaging Visualizations
+title: Moving Towards Reproducible and Programmatic Generation of Neuroimaging Visualizations
 authors:
@@
   - Brain Visualization
-abstract: Neuroimaging visualization forms the centrepiece of quality control, and interpretation and communication of results.
+abstract: Neuroimaging visualization forms the centerpiece of quality control, and interpretation and communication of results.
 Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI).
-There now exist many well-documented code-based brain visualisation tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB.
+There now exist many well-documented code-based brain visualization tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB.
 Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process.
-This article reviews the advantages of learning and using programmatic neuroimaging visualization tools and provides a list of tools across programming environments.
-Finally, we introduce examples of specific R and Python tools for visualization of voxel, vertex, region-of-interest, and edge level data.
+Here we outline the advantages of learning and using programmatic neuroimaging visualization tools, include a list of these tools across programming environments and give examples of code and visualizations at the voxel, vertex, region-of-interest, and edge-level data.
 
@@
 In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility.
-However, there now exists many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community (Table 1).
+However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community (Table 1).
 These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments.
-The present article argues for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
+The present commentary argues for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
 We also providing didactic examples of visualizations and associated code as a reference point.
@@
 While there is a trend towards increasing transparency and code sharing of neuroimaging analyses, the sharing of code used to generate figures which include brain renderings and spatial maps has been relatively neglected.
-This gap in reproducibility is partly driven by the fact that brain figures are often created using a manual process that involves tinkering with sliders, buttons, and overlays on a GUI, concluding with a screenshot and sometimes beautification in image processing software like Illustrator or Photoshop.
+This gap in reproducibility is partly driven by the fact that brain figures are often created using a manual process that involves tinkering with sliders, buttons, and overlays on a GUI, concluding with a screenshot and sometimes beautification in image processing software like Illustrator, Photoshop or Inkscape.
 Such a process inherently makes neuroimaging visualizations difficult, if not impossible to replicate, even by the authors themselves.
@@
 Linking to or even embedding these videos or interactive figures in papers can greatly enhance the communication of findings and make your paper more engaging for the reader.
-Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of your research with peers and a reach larger audiences (Li and Xie, 2020).
+Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of your research with peers and reach larger audiences (Li and Xie, 2020).
 This last point is becoming increasingly salient as marketing science on social media has become a core medium for spreading discoveries, science communication to the public, and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019).
@@
 |                                                        | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
-|:-----------|:-----------|:-----------|:-----------|:-----------|:-----------|
+|:------------------|:----------|:----------|:----------|:----------|:----------|
 | **R**                                                  |         |         |               |         |             |
@@
 This list does not include cross-platform general purpose visualization software.\
-\*Cortex only
+***\****Cortex only
 
@@
 ```
-```{r Figure1 , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center'}
+
+```{r Figure1, echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center'}
 
@@
 
-fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
+fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g', 'fig1h', 'fig1i')
 figs <- list()
@@
 
-patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]])+plot_layout(heights = c(0.5,1,1)))
+patchwork <- figs[[1]]/figs[[2]]|((figs[[3]]/(figs[[4]]|figs[[5]]))/figs[[6]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[7]]/figs[[8]]/figs[[9]])+plot_layout(heights = c(0.5,1,1)))
 
@@
 ```
-```{r Figure2 , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center', eval=FALSE}
+
+```{r Figure2, echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center', eval=FALSE}
 
@@
 
-fig_name <- c('fig1', 'fig2', 'fig3', 'fig4', 'fig5', 'fig6', 'fig7')
+fig_name <- c('fig2a', 'fig2b', 'fig2c', 'fig2d', 'fig2e', 'fig2f')
 figs <- list()
@@
 
-patchwork2 <- (figs[[1]]+figs[[3]])/figs[[2]] | (figs[[6]]/figs[[4]]/figs[[5]])
+patchwork1 <- (figs[[1]] + figs[[3]]) / figs[[2]] + plot_layout(heights = c(1.2,0.8))
 
-  figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]])+plot_layout(heights = c(0.5,1,1)))
+patchwork2 <-  figs[[4]] / (plot_spacer() + figs[[5]] + plot_spacer()) / figs[[6]] + plot_layout(heights = c(1.2,0.8, 1.2))
 
@@
 
-## Voxel-Level Visualizations.
-
-**R.** The `oro.nifti` and `neurobase` packages allow for the reading, writing, manipulation and visualization of voxel-level imaging data of either *nifti*, *analyze* or *afni* formats.
-In particular, the `image`, `slice` and `ortho2` functions allow users to display the desired number of slices in the desired orientations, while also providing precise control over overlaid images, color scales, legend placement and other aesthetic settings (Fig 1A-B).
-Smoothing and re-sampling of images is sometimes required for visualization purposes, which can be achieved via the `ANTsR` or `fslr` packages.
-
-**Python.** The `NiBable` library provides the ability to read in and covert many common neuroimaging file formats, including *nifti,* to data-structures which can be manipulated and visualized.
-While these data-structures often form 3D or 4D matrices, so can be visualized using commonly used general purpose visualization tools such as `matplotlib` , purpose built librires like `nilearn` provide functions like `plot_stat_map` and `plot_glass_brain` which are user friendly and can generate publication ready figures (Fig.
-X).
-`nilearn` also includes smoothing and re-sampling functions which can sometimes required for visualization purposes.
-
-## Vertex-Level Visualizations.
-
-**R.** The recently developed `fsbrain` package allows for the visualization of vertex-level and cortical ROI-level data which were derived using *FreeSurfer*.
-The package contains many flexible functions for visualizing individual-level and group-level measures such as cortical thickness, volume or surface area (Fig 1F).
-In addition to producing publication-ready figures, the visualization functions can allow for visual quality control of large datasets processed using FreeSurfer.
-The package also comes with extensive guides and documentations to assist users (Schäfer & Ecker, 2020).
-
-CIFTI or 'grey-ordinate' data is becoming a popular format for structural and functional imaging data, as it combines vertex-level data of the cortex with voxel-level data for the cerebellum and sub-cortex.
-In R, CIFTI data can now be read, visualized and manipulated using the well-documented `ciftiTools` package (Fig 1C,1G; Pham, Muschelli, & Mejia, 2022).
-This package allows for both vertex-level visualization of the surface as well as voxel-level visualization of cerebellum and subcortex.
-
-**Python.** ciftify.
-
-## ROI-Level Visualizations.
-
-**R.** Statistical parameters can be mapped onto discreet cortical and sub-cortical regions using the `ggseg` and `ggseg3d` packages (Mowinckel & Vidal-Piñeiro, 2020).
-These packages flexibly generates aesthetic renderings of cortical, sub-cortical and white matter ROIs in 2D (Fig 1D,1F) and 3D (Fig 1E).
-Usually, ROI-level interpretations are dependent on a standardized atlas or parcellation scheme of the brain and accordingly, this package and its sister-packages (`ggsegExtra`) provide a large array of commonly used atlases and some functionality for users to contribute other atlases.
-Additionally, being part of the 'Grammar of Graphics' framework (Wickham, 2016) allows for integration with packages such as `gganimate`, enabling users to animate and visualize dynamic changes in spatial and temporal statistics across ROIs.
-Notably, the `ciftiTools` package also provides functions to visualize surface and voxel-level ROIs for the cortex and sub-cortex respectively (Fig 1E), flexibly allowing CIFTI format atlases and ROIs to be displayed.
-
-**Python.**
-
-## Edge-Level Visualizations.
-
-Analyses which divide the brain into discrete regions and examine pair-wise dependencies between regional phenotypes are becoming increasingly common.
-Often the results of these dependency analyses are statistics relating to 'edges' or links between any two regions.
-The package `brainconn` allows users to plot brain regions as nodes of a network graph, and dependencies between regions as edges of that graph, following a standardized coordinate space in both 2D and interactive 3D (Figure 1I).
-The graphs can be weighted, with the strength of the edges represented with color or line thickness, as well as directed, where arrows will be used in place of edges.
-This package comes with node coordinates corresponding to multiple commonly used atlases and contains functionality for users to easily enter their own custom atlas coordinates.
-
-**Python.**
-
 # Limitations and Functionality Gaps
 
-Most of the tools introduced above do not require a strong knowledge of programming, but there is still a steeper learning curve when compared to use a GUI.
+Most of the tools introduced above do not require a strong knowledge of programming, but there is still a steeper learning curve compared to using a GUI.
 This is especially true when learning how to make fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning, for the purpose of a publication-ready figure.
@@
 Figure 1), potentially mirroring the cortico-centric sentiment that has prevailed in human neuroimaging research (Chin, Chang, & Holmes, 2022).
-Likewise, custom non-cortical atlases such as non-standard subcortical atlas schemes are not yet straight forward, and usually require multiple functions and packages to visualize.
+Likewise, custom non-cortical atlases such as non-standard subcortical atlas schemes are not yet straightforward, and usually require multiple functions and packages to visualize.
 Finally, some neuroimaging related data types, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools.
@@
 
-# Supplement (Code used to generate Figures)
+# Supplement 1 (Code used to generate Figure 1 - R Visulaizations)
 
@@
 
-```{r Figure1-sup , echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center', eval=FALSE}
+```{r Fig1 compile, eval=FALSE}
+# This code block complies all the panels of the figure into one, largely using the patchwork package.
 
@@
 for (f in 1:length(fig_name)) { 
-  knitr::plot_crop(paste0('data/',fig_name[f],'.png'))
-  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/',fig_name[f],'.png')),
+  knitr::plot_crop(paste0('data/R/',fig_name[f],'.png'))
+  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/R/',fig_name[f],'.png')),
                                                              width=ggplot2::unit(1,"npc"),
@@
 
-patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]])+plot_layout(heights = c(0.7,1,1)))
+patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]]) + 
+  plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]]) + 
+                                           plot_layout(heights = c(0.5,1,1)))
+
 patchwork + plot_annotation(tag_levels = 'A',
-                            title = 'Examples of brain visualizations in R') &
+                            title = 'Examples of brain visualizations made in R') &
   theme(plot.tag = element_text(face = 'bold'),
@@
 
-```{r Fig1a, eval=FALSE}
+```{r Fig1A, eval=FALSE}
 #Figure 1A (Voxel)
@@
 #[Optional] If you want to output the figure as a .png, open a png image device
-png("data/fig1a.png")
+png("data/R/fig1a.png")
 
@@
        NA.x = T,       #Do not display NA values
-       col.y= rev(viridis::inferno(n = 500)), #Select the color scale. In this case I have used the inferno scale from the viridis package. 
+       col.y= rev(viridis::inferno(n = 500)), #Select the color scale. 
        xyz = c(70,50,80), #set the x y & z slice you want to visualize
@@
   colorbar(breaks, col = rev(viridis::inferno(n=length(breaks)-1)), #add a color bar with same color scale 
-           text.col = "black", labels = TRUE, maxleft = 0.95)  #Set the specifications for the color bar
+           text.col = "black", labels = FALSE, maxleft = 0.85)  #Set the specifications for the color bar
 
@@
 
-```{r Fig1b, eval=FALSE }
+```{r Fig1B, eval=FALSE}
 #Figure 1B (Voxel)
@@
 
-png('data/fig1b.png')
+png('data/R/fig1b.png')
 overlay(x=robust_window(t1), y=atlas, 
-        plot.type = "single", 
-        z=c(seq(30,110,10)), 
+        plot.type = "single", #Single view (axial only)
+        z=c(seq(30,110,10)), #set slices you want to view (30 to 110, with a gap of 10 slices). 
         col.y =  alpha(viridis::inferno(300), 0.6),
@@
         bg = "white",
-        NA.x=T, 
+        NA.x=T, #Make NA values on overlay transparent
         zlim.y =c(1,300))
@@
 
-```{r Fig1c , eval=FALSE}
-#vertex 1
-#Figure 1C
+```{r Fig1C, eval=FALSE}
+#Figure 1C (Cortical ROI using ciftiTools)
+
+# Load the package 
+library(ciftiTools)
 library(rgl)
-library(ggplot2)
-# Load the package and point to the Connectome Workbench
-library(ciftiTools)
 
@@
 
-# Read in CIFTI file
+#Define CIFTI file names for time series and surface files. 
 cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
@@
 
+# Read in CIFTI file
 cii <- read_cifti(
@@
 
-#Plot cifti surface
-view_xifti_surface(cii, 
-                   hemisphere = "both", 
-                   view = "lateral", 
-                   idx = 1,
-                   colors = viridis::plasma(n = 100),
-                   zlim = c(1.2,1.5),  
-                   legend_embed = T,
-                   cex.title = 2)
-
-rgl.snapshot("data/fig1c.png")
-rgl.close()
-
-```
-
-```{r Fig1d , eval=FALSE}
-#Vertex 2
-#Figure 1X - Remove white space
-
-library(fsbrain)
-library(rgl)
-
-#download_optional_data()
-#download_fsaverage(accept_freesurfer_license = TRUE)
-
-subjects_dir <- get_optional_data_filepath("subjects_dir")
-
-subject_id <- 'subject1'
-
-colourmap <- colorRampPalette(viridis::plasma(n = 1000))
-fsbrain.set.default.figsize(700, 700);
-
-brain = vis.subject.morph.native(subjects_dir, 
-                                subject_id, 'sulc', 
-                                cortex_only = T, 
-                                views=NULL, 
-                                draw_colorbar = FALSE,
-                                makecmap_options = list('colFn'=colourmap))
-
-img = export(brain, draw_colorbar = FALSE, output_img = "data/fig1d.png")
-
-```
-
-```{r Fig1e , eval=FALSE}
-#ROI1 
-#Figure 1e 
-library(ciftiTools)
-parc <- load_parc("Yeo_7")
-view_xifti_surface(parc,  
-                   legend_embed = T,
-                   hemisphere = "right")
-
-rgl.snapshot("data/fig1e.png")
-
-rgl.close()
-
-```
-
-```{r Fig1f , eval=FALSE}
-#ROI1 
-#Figure 1X - combine with previous figure
-
 set.seed(1993) #set a random seed (good practice for reproducibility)
 
-parc <- load_parc("Schaefer_400")
+parc <- load_parc("Schaefer_400") #ciftitools has a few atlas in cifti format already included
 
-ramdom_metric <- rnorm(400)
+ramdom_metric <- rnorm(400) #generate random data to plot on cortex
 
@@
 
-plot2 <- view_xifti_surface(xii_metric, 
+view_xifti_surface(xii_metric, 
                    colors = viridis::viridis(n = 400), 
@@
 
-rgl.snapshot("data/fig1f.png")
+rgl.snapshot("data/R/fig1c.png")
 
@@
 
-```{r Fig1g, eval=FALSE}
-#ROI2
+```{r Fig1D, eval=FALSE}
+#Figure 1D (Subcortical ROI using ggseg)
 
-#devtools::install_github("LCBC-UiO/ggsegGlasser")
 library(ggseg)
 library(ggplot2)
-library(ggsegGlasser)
-
-set.seed(1993) #set a random seed (good practice for reproducibility)
-
-base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))
-
-colnames(base_atlas) <- c("region", "hemi")
-
-Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas
-
-base_atlas <- cbind(Effect, base_atlas)
-
-cortex <- ggseg(atlas = glasser,
-                .data = base_atlas,
-                mapping=aes(fill=Effect), 
-                position="stacked",
-                colour="black",
-                size=.2,
-                show.legend = T,
-                plot.background = "white") +
-  scale_fill_viridis_b() + 
-   theme_void() +
-    theme(legend.position = "bottom")
 
@@
 sub_base_atlas <- cbind(Effect, sub_base_atlas)
-
 subcortex <- ggseg(atlas = aseg,
@@
                    plot.background = "white") + 
-  scale_fill_viridis_b()
+  scale_fill_viridis_b() +
+  theme(legend.text=element_text(size=15),
+        legend.title  = element_blank())
 
-ggsave(filename = 'data/fig1g1.png', plot = cortex, device = 'png', bg = 'white')
-ggsave(filename = 'data/fig1g2.png', plot = subcortex, device = 'png', bg = 'white')
-
+ggsave(filename = 'data/R/fig1d.png', plot = subcortex, device = 'png', bg = 'white')
 ```
 
-```{r, eval=FALSE}
+```{r Fig1E, eval=FALSE}
+#Figure 1E (Subcortical ROI using ggseg3D)
 
+#The function below plots 3d version of aseg. See "inputs" within function for more details. 
 make_ggseg3d <- function(attribute, 
@@
              colour.pal = viridis::viridis(n=length(volume)), 
-             output.png = T, file.name = "data/fig1g3.png")
+             output.png = T,
+             file.name = "data/R/fig1e.png")
 
@@
 
-```{r Fig1h , eval = F}
-#Edge (also add 3d version)
-#Fig1X
+```{r Fig1F, eval=FALSE}
+#Figure 1F (Cortical ROI using ggseg)
+
+library(ggseg)
+library(ggplot2)
+#devtools::install_github("LCBC-UiO/ggsegGlasser")
+library(ggsegGlasser)
+
+#create dataframe with region names and hemisphere for the glasser atlas
+base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))
+
+colnames(base_atlas) <- c("region", "hemi")
+
+set.seed(1993) #set a random seed (good practice for reproducibility)
+Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas
+
+base_atlas <- cbind(Effect, base_atlas)
+
+#Cortex
+cortex <- ggseg(atlas = glasser,
+                .data = base_atlas,
+                mapping=aes(fill=Effect), 
+                position="stacked",
+                colour="black",
+                size=.2,
+                show.legend = T,
+                plot.background = "white") +
+  scale_fill_viridis_b() + 
+  theme_void() +
+  theme(legend.position = "bottom",
+        legend.text=element_text(size=15),
+        legend.title  = element_blank())
+
+ggsave(filename = 'data/R/fig1f.png', plot = cortex, device = 'png', bg = 'white')
+
+```
+
+```{r Fig1G, eval=FALSE}
+#Figure 1G (Vertex using ciftiTools)
+
+library(rgl)
+library(ggplot2)
+
+# Load the package and point to the Connectome Workbench
+library(ciftiTools)
+
+#Need connectome workbench to use ciftiTools - set path below
+ciftiTools.setOption("wb_path", "/Applications/workbench/")
+
+#Define CIFTI file names for time series and surface files. 
+cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
+surfL_fname <- ciftiTools.files()$surf["left"]
+surfR_fname <- ciftiTools.files()$surf["right"]
+
+# Read in CIFTI file
+cii <- read_cifti(
+  cifti_fname, brainstructures="all", 
+  surfL_fname=surfL_fname,
+  surfR_fname=surfR_fname)
+
+#Plot mean time series on CIFTI surface
+view_xifti_surface(cii, 
+                   hemisphere = "both", 
+                   view = "lateral", 
+                   idx = 1,
+                   colors = viridis::plasma(n = 100),
+                   zlim = c(1.2,1.5),  
+                   legend_embed = T)
+
+rgl.snapshot("data/R/fig1g.png") #save output in RGL window as 
+rgl.close()
+
+```
+
+```{r Fig1H, eval=FALSE}
+#Figure 1H (Vertex using fsbrain)
+
+library(fsbrain)
+library(rgl)
+
+#download example freesurfer data which comes with the fsbrain package
+#download_optional_data() 
+#download_fsaverage(accept_freesurfer_license = TRUE)
+
+subjects_dir <- get_optional_data_filepath("subjects_dir")
+
+subject_id <- 'subject1'
+
+colourmap <- colorRampPalette(viridis::plasma(n = 1000)) #set colourmap
+
+# set figure size (optional)
+fsbrain.set.default.figsize(700, 700)
+
+# generate figure
+brain = vis.subject.morph.native(subjects_dir, 
+                                 subject_id, 'sulc', 
+                                 cortex_only = T, 
+                                 views=NULL, 
+                                 draw_colorbar = FALSE,
+                                 makecmap_options = list('colFn'=colourmap))
+
+img = export(brain, draw_colorbar = FALSE, output_img = "data/R/fig1h.png") #write out figure
+
+```
+
+```{r Fig1I, eval=FALSE}
+#Fig1I (Edges visualized using brainconn)
 
@@
 
-conmat <- example_weighted_undirected
+conmat <- example_weighted_undirected #one of the example matrices included with brainconn
 degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]
 
-brainconn(atlas ="schaefer300_n7", 
+fig1i <- brainconn(atlas ="schaefer300_n7", 
                    conmat=conmat, 
                    node.size = degree/2,
-                   view="top", 
+                   view="ortho", 
                    edge.color.weighted = T,
-                   background.alpha = 0.8, 
-                   show.legend = F, 
-                   edge.color = scale_edge_colour_viridis())
+                   background.alpha = 0.6, 
+                   edge.color = scale_edge_colour_viridis(option = "C"))
 
-ggsave(filename = 'data/fig1h.png', 
-       plot = fig1h, 
+ggsave(filename = 'data/R/fig1i.png', 
+       plot = fig1i, 
        device = 'png',
@@
 
-brainconn3D(atlas ="schaefer300_n7", 
-            conmat=conmat,
-            show.legend = F)
-
 ```

### 2022-08-31 manuscript.Rmd
clean up ignore
sup sep

@@
 
-patchwork1 <- (figs[[1]] + figs[[3]]) / figs[[2]] + plot_layout(heights = c(1.2,0.8))
+layoutA <- "
+AAAA
+#BB#
+CCCC
+"
+patchwork1 <- (((figs[[1]] + figs[[3]]) + plot_layout(widths  = c(0.8,1))) / figs[[2]])  + plot_layout(heights = c(1,0.5)) 
+patchwork1
 
-patchwork2 <-  figs[[4]] / (plot_spacer() + figs[[5]] + plot_spacer()) / figs[[6]] + plot_layout(heights = c(1.2,0.8, 1.2))
+patchwork2 <-  (figs[[4]] / figs[[5]] / figs[[6]]) + plot_layout(design = layoutA, widths = c(0.2,1,0.2))
+
+(patchwork1 | patchwork2) + plot_layout(widths = c(1,0.6))
+
+patchwork2 <-  (figs[[4]] / figs[[5]] / figs[[6]] + plot_layout(design = layoutA, widths = c(0.2,1,0.2)))
+
+patchwork2
 
@@
 \newpage
-
-# Supplement 1 (Code used to generate Figure 1 - R Visulaizations)
-
-\singlespacing
-
-```{r Fig1 compile, eval=FALSE}
-# This code block complies all the panels of the figure into one, largely using the patchwork package.
-
-library(patchwork)
-library(ggplot2)
-
-fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
-figs <- list()
-
-for (f in 1:length(fig_name)) { 
-  knitr::plot_crop(paste0('data/R/',fig_name[f],'.png'))
-  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/R/',fig_name[f],'.png')),
-                                                             width=ggplot2::unit(1,"npc"),
-                                                             height=ggplot2::unit(1,"npc")))
-}
-
-patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]]) + 
-  plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]]) + 
-                                           plot_layout(heights = c(0.5,1,1)))
-
-patchwork + plot_annotation(tag_levels = 'A',
-                            title = 'Examples of brain visualizations made in R') &
-  theme(plot.tag = element_text(face = 'bold'),
-        plot.title = element_text(face="bold"))
-
-```
-
-```{r Fig1A, eval=FALSE}
-#Figure 1A (Voxel)
-
-library(neurobase)
-
-#Load in a nifti (.nii.gz) file of a standardized template so use as a background. 
-template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R
-
-#Load in a nifti (.nii.gz) file of the statistic you wanted of overlay, with the same dimensions as the template.  
-effect <- readnii("data/MNI152_effect_size.nii.gz")
-
-#Threshold the statistic map so only values about a specified amount are displayed
-effect[effect<0.3] <- NA
-
-#Set the breaks/intervals you want on the color bar (e.g. from .3 to .6, by intervals of 0.005)
-breaks = seq(.3,.6, by=0.005)
-
-#[Optional] If you want to output the figure as a .png, open a png image device
-png("data/R/fig1a.png")
-
-#Use the ortho2 function from the neurobase package to make the figure
-ortho2(x = template,   #Specify the background template
-       y = effect,     #Specify the effect you want to overlay on the template
-       crosshairs = F, #Remove the cross-hairs
-       bg = "white",   #Make the background white 
-       NA.x = T,       #Do not display NA values
-       col.y= rev(viridis::inferno(n = 500)), #Select the color scale. 
-       xyz = c(70,50,80), #set the x y & z slice you want to visualize
-       useRaster = T,     #Sometimes using Raster makes for clearer plots
-       ycolorbar = TRUE,  #Add a color bar
-       mfrow = c(1,1)) +  #layout of brain views (for this figure, we just want a axial view)
-  colorbar(breaks, col = rev(viridis::inferno(n=length(breaks)-1)), #add a color bar with same color scale 
-           text.col = "black", labels = FALSE, maxleft = 0.85)  #Set the specifications for the color bar
-
-dev.off()
-
-```
-
-```{r Fig1B, eval=FALSE}
-#Figure 1B (Voxel)
-
-library(neurobase)
-library(scales)
-
-#Load in a nifti (.nii.gz) file to use as a background. 
-t1 <- readnii('data/sub-001__t1_warped.nii.gz')
-
-#Load in a nifti (.nii.gz) file to use as a overlay. 
-atlas <-  readnii('data/sub-001_schaefer300n7_aseg_to_dwispace_gm_rois.nii.gz')
-
-png('data/R/fig1b.png')
-overlay(x=robust_window(t1), y=atlas, 
-        plot.type = "single", #Single view (axial only)
-        z=c(seq(30,110,10)), #set slices you want to view (30 to 110, with a gap of 10 slices). 
-        col.y =  alpha(viridis::inferno(300), 0.6),
-        plane = "axial",  
-        useRaster = T, 
-        bg = "white",
-        NA.x=T, #Make NA values on overlay transparent
-        zlim.y =c(1,300))
-dev.off()
-
-```
-
-```{r Fig1C, eval=FALSE}
-#Figure 1C (Cortical ROI using ciftiTools)
-
-# Load the package 
-library(ciftiTools)
-library(rgl)
-
-#Need connectome workbench to use ciftiTools - set path below
-ciftiTools.setOption("wb_path", "/Applications/workbench/")
-
-#Define CIFTI file names for time series and surface files. 
-cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
-surfL_fname <- ciftiTools.files()$surf["left"]
-surfR_fname <- ciftiTools.files()$surf["right"]
-
-# Read in CIFTI file
-cii <- read_cifti(
-  cifti_fname, brainstructures="all", 
-  surfL_fname=surfL_fname,
-  surfR_fname=surfR_fname)
-
-set.seed(1993) #set a random seed (good practice for reproducibility)
-
-parc <- load_parc("Schaefer_400") #ciftitools has a few atlas in cifti format already included
-
-ramdom_metric <- rnorm(400) #generate random data to plot on cortex
-
-cii <- move_from_mwall(cii, NA)
-
-parc_vec <- c(as.matrix(parc))
-
-xii_metric <- c(NA, ramdom_metric)[parc_vec + 1]
-
-xii1 <- select_xifti(cii, 1)
-
-xii_metric <- newdata_xifti(xii1, xii_metric)
-
-view_xifti_surface(xii_metric, 
-                   colors = viridis::viridis(n = 400), 
-                   borders = "black", 
-                   hemisphere = "both",
-                   view = 'lateral')
-
-rgl.snapshot("data/R/fig1c.png")
-
-rgl.close()
-
-```
-
-```{r Fig1D, eval=FALSE}
-#Figure 1D (Subcortical ROI using ggseg)
-
-library(ggseg)
-library(ggplot2)
-
-sub_base_atlas <- as.data.frame(na.omit(cbind(aseg$data$region, aseg$data$hemi)))
-
-colnames(sub_base_atlas) <- c("region", "hemi")
-
-Effect <- rnorm(dim(sub_base_atlas)[1])
-
-sub_base_atlas <- cbind(Effect, sub_base_atlas)
-subcortex <- ggseg(atlas = aseg,
-                   .data = sub_base_atlas,
-                   mapping=aes(fill=Effect), 
-                   position = "dispersed",
-                   colour="black",
-                   hemi =c('left', 'right'),
-                   size=.2,
-                   show.legend = T, 
-                   plot.background = "white") + 
-  scale_fill_viridis_b() +
-  theme(legend.text=element_text(size=15),
-        legend.title  = element_blank())
-
-ggsave(filename = 'data/R/fig1d.png', plot = subcortex, device = 'png', bg = 'white')
-```
-
-```{r Fig1E, eval=FALSE}
-#Figure 1E (Subcortical ROI using ggseg3D)
-
-#The function below plots 3d version of aseg. See "inputs" within function for more details. 
-make_ggseg3d <- function(attribute, 
-                         colour.pal = c("light yellow",
-                                        "orange",
-                                        "red",
-                                        "dark red"), 
-                         hide.colourbar=FALSE, 
-                         output.png=FALSE, 
-                         file.name="ggseg_3d.png") {
-  # Inputs: 
-  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left
-  #' -Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Righ
-  #' t-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Righ
-  #' t-Amygdala, Right-Accumbens-area
-
-  library(ggseg3d)
-  library(tidyr)
-  library(dplyr)
-  #remove(aseg_3d)
-  aseg_3d <- aseg_3d
-  aseg_3d <- tidyr::unnest(aseg_3d, cols = c(ggseg_3d))
-
-  attribute.ggseg3d <- c(rep(NA, 4), attribute[1:4], rep(NA, 3), attribute[5:7],
-                         rep(NA, 5), attribute[8:14],  rep(NA, 6))
-
-  data <- dplyr::mutate(aseg_3d, attribute =  attribute.ggseg3d)
-  #remove NA regions
-  aseg_3d[which(is.na(data$attribute)),] <- NA
-  aseg_3d <- tidyr::drop_na(aseg_3d)
-
-  data[which(is.na(data$attribute)),] <- NA
-  data <- tidyr::drop_na(data)
-  data$attribute[data$attribute==0]<-NA #make 0 valus NA so they are set as grey in ggseg3d
-
-  scene=list(camera = list(eye = list(x = 0, y = 1, z = -2.25)),
-             aspectratio = list(x=1.6,y=1.6,z=1.6))
-
-  plot <- ggseg3d::ggseg3d(.data = data, 
-                           atlas = aseg_3d, 
-                           colour = "attribute", 
-                           text = "attribute",
-                           palette = colour.pal)
-  plot <- remove_axes(plot)
-  plot <- plotly::layout(plot, 
-                         scene = scene, 
-                         width = 600, height = 600)
-
-  if(hide.colourbar==TRUE){
-    plot <- plotly::hide_colorbar(plot)
-  }
-
-  if( output.png==TRUE){
-    plotly::orca(plot, file =  file.name)
-  }
-
-  return(plot)
-
-}
-
-set.seed(1993)
-volume <- sample(1:20, 14) #randomly generated data for 14 structures
-make_ggseg3d(volume, 
-             colour.pal = viridis::viridis(n=length(volume)), 
-             output.png = T,
-             file.name = "data/R/fig1e.png")
-
-```
-
-```{r Fig1F, eval=FALSE}
-#Figure 1F (Cortical ROI using ggseg)
-
-library(ggseg)
-library(ggplot2)
-#devtools::install_github("LCBC-UiO/ggsegGlasser")
-library(ggsegGlasser)
-
-#create dataframe with region names and hemisphere for the glasser atlas
-base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))
-
-colnames(base_atlas) <- c("region", "hemi")
-
-set.seed(1993) #set a random seed (good practice for reproducibility)
-Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas
-
-base_atlas <- cbind(Effect, base_atlas)
-
-#Cortex
-cortex <- ggseg(atlas = glasser,
-                .data = base_atlas,
-                mapping=aes(fill=Effect), 
-                position="stacked",
-                colour="black",
-                size=.2,
-                show.legend = T,
-                plot.background = "white") +
-  scale_fill_viridis_b() + 
-  theme_void() +
-  theme(legend.position = "bottom",
-        legend.text=element_text(size=15),
-        legend.title  = element_blank())
-
-ggsave(filename = 'data/R/fig1f.png', plot = cortex, device = 'png', bg = 'white')
-
-```
-
-```{r Fig1G, eval=FALSE}
-#Figure 1G (Vertex using ciftiTools)
-
-library(rgl)
-library(ggplot2)
-
-# Load the package and point to the Connectome Workbench
-library(ciftiTools)
-
-#Need connectome workbench to use ciftiTools - set path below
-ciftiTools.setOption("wb_path", "/Applications/workbench/")
-
-#Define CIFTI file names for time series and surface files. 
-cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
-surfL_fname <- ciftiTools.files()$surf["left"]
-surfR_fname <- ciftiTools.files()$surf["right"]
-
-# Read in CIFTI file
-cii <- read_cifti(
-  cifti_fname, brainstructures="all", 
-  surfL_fname=surfL_fname,
-  surfR_fname=surfR_fname)
-
-#Plot mean time series on CIFTI surface
-view_xifti_surface(cii, 
-                   hemisphere = "both", 
-                   view = "lateral", 
-                   idx = 1,
-                   colors = viridis::plasma(n = 100),
-                   zlim = c(1.2,1.5),  
-                   legend_embed = T)
-
-rgl.snapshot("data/R/fig1g.png") #save output in RGL window as 
-rgl.close()
-
-```
-
-```{r Fig1H, eval=FALSE}
-#Figure 1H (Vertex using fsbrain)
-
-library(fsbrain)
-library(rgl)
-
-#download example freesurfer data which comes with the fsbrain package
-#download_optional_data() 
-#download_fsaverage(accept_freesurfer_license = TRUE)
-
-subjects_dir <- get_optional_data_filepath("subjects_dir")
-
-subject_id <- 'subject1'
-
-colourmap <- colorRampPalette(viridis::plasma(n = 1000)) #set colourmap
-
-# set figure size (optional)
-fsbrain.set.default.figsize(700, 700)
-
-# generate figure
-brain = vis.subject.morph.native(subjects_dir, 
-                                 subject_id, 'sulc', 
-                                 cortex_only = T, 
-                                 views=NULL, 
-                                 draw_colorbar = FALSE,
-                                 makecmap_options = list('colFn'=colourmap))
-
-img = export(brain, draw_colorbar = FALSE, output_img = "data/R/fig1h.png") #write out figure
-
-```
-
-```{r Fig1I, eval=FALSE}
-#Fig1I (Edges visualized using brainconn)
-
-#devtools::install_github("sidchop/brainconn")
-library(brainconn)
-
-conmat <- example_weighted_undirected #one of the example matrices included with brainconn
-degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]
-
-fig1i <- brainconn(atlas ="schaefer300_n7", 
-                   conmat=conmat, 
-                   node.size = degree/2,
-                   view="ortho", 
-                   edge.color.weighted = T,
-                   background.alpha = 0.6, 
-                   edge.color = scale_edge_colour_viridis(option = "C"))
-
-ggsave(filename = 'data/R/fig1i.png', 
-       plot = fig1i, 
-       device = 'png',
-       bg = 'white',
-       height = 7, 
-       units = "in")
-
-```

### 2022-08-31 SupplementA_R_Figs.Rmd
sup sep

---
title: "Supplement 1 - code for R figures"
output: html_notebook
---

```{r Fig1A, eval=FALSE}
#Figure 1A (Voxel)

library(neurobase)

#Load in a nifti (.nii.gz) file of a standardized template so use as a background. 
template <- readnii('data/MNI152_T1_1mm_brain.nii.gz') #load the nifti file into R

#Load in a nifti (.nii.gz) file of the statistic you wanted of overlay, with the same dimensions as the template.  
effect <- readnii("data/MNI152_effect_size.nii.gz")

#Threshold the statistic map so only values about a specified amount are displayed
effect[effect<0.3] <- NA

#Set the breaks/intervals you want on the color bar (e.g. from .3 to .6, by intervals of 0.005)
breaks = seq(.3,.6, by=0.005)

#[Optional] If you want to output the figure as a .png, open a png image device
png("data/R/fig1a.png")

#Use the ortho2 function from the neurobase package to make the figure
ortho2(x = template,   #Specify the background template
       y = effect,     #Specify the effect you want to overlay on the template
       crosshairs = F, #Remove the cross-hairs
       bg = "white",   #Make the background white 
       NA.x = T,       #Do not display NA values
       col.y= rev(viridis::inferno(n = 500)), #Select the color scale. 
       xyz = c(70,50,80), #set the x y & z slice you want to visualize
       useRaster = T,     #Sometimes using Raster makes for clearer plots
       ycolorbar = TRUE,  #Add a color bar
       mfrow = c(1,1)) +  #layout of brain views (for this figure, we just want a axial view)
  colorbar(breaks, col = rev(viridis::inferno(n=length(breaks)-1)), #add a color bar with same color scale 
           text.col = "black", labels = FALSE, maxleft = 0.85)  #Set the specifications for the color bar

dev.off()

```

```{r Fig1B, eval=FALSE}
#Figure 1B (Voxel)

library(neurobase)
library(scales)

#Load in a nifti (.nii.gz) file to use as a background. 
t1 <- readnii('data/sub-001__t1_warped.nii.gz')

#Load in a nifti (.nii.gz) file to use as a overlay. 
atlas <-  readnii('data/sub-001_schaefer300n7_aseg_to_dwispace_gm_rois.nii.gz')

png('data/R/fig1b.png')
overlay(x=robust_window(t1), y=atlas, 
        plot.type = "single", #Single view (axial only)
        z=c(seq(30,110,10)), #set slices you want to view (30 to 110, with a gap of 10 slices). 
        col.y =  alpha(viridis::inferno(300), 0.6),
        plane = "axial",  
        useRaster = T, 
        bg = "white",
        NA.x=T, #Make NA values on overlay transparent
        zlim.y =c(1,300))
dev.off()

```

```{r Fig1C, eval=FALSE}
#Figure 1C (Cortical ROI using ciftiTools)

# Load the package 
library(ciftiTools)
library(rgl)

#Need connectome workbench to use ciftiTools - set path below
ciftiTools.setOption("wb_path", "/Applications/workbench/")

#Define CIFTI file names for time series and surface files. 
cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
surfL_fname <- ciftiTools.files()$surf["left"]
surfR_fname <- ciftiTools.files()$surf["right"]

# Read in CIFTI file
cii <- read_cifti(
  cifti_fname, brainstructures="all", 
  surfL_fname=surfL_fname,
  surfR_fname=surfR_fname)

set.seed(1993) #set a random seed (good practice for reproducibility)

parc <- load_parc("Schaefer_400") #ciftitools has a few atlas in cifti format already included

ramdom_metric <- rnorm(400) #generate random data to plot on cortex

cii <- move_from_mwall(cii, NA)

parc_vec <- c(as.matrix(parc))

xii_metric <- c(NA, ramdom_metric)[parc_vec + 1]

xii1 <- select_xifti(cii, 1)

xii_metric <- newdata_xifti(xii1, xii_metric)

view_xifti_surface(xii_metric, 
                   colors = viridis::viridis(n = 400), 
                   borders = "black", 
                   hemisphere = "both",
                   view = 'lateral')

rgl.snapshot("data/R/fig1c.png")

rgl.close()

```

```{r Fig1D, eval=FALSE}
#Figure 1D (Subcortical ROI using ggseg)

library(ggseg)
library(ggplot2)

sub_base_atlas <- as.data.frame(na.omit(cbind(aseg$data$region, aseg$data$hemi)))

colnames(sub_base_atlas) <- c("region", "hemi")

Effect <- rnorm(dim(sub_base_atlas)[1])

sub_base_atlas <- cbind(Effect, sub_base_atlas)
subcortex <- ggseg(atlas = aseg,
                   .data = sub_base_atlas,
                   mapping=aes(fill=Effect), 
                   position = "dispersed",
                   colour="black",
                   hemi =c('left', 'right'),
                   size=.2,
                   show.legend = T, 
                   plot.background = "white") + 
  scale_fill_viridis_b() +
  theme(legend.text=element_text(size=15),
        legend.title  = element_blank())

ggsave(filename = 'data/R/fig1d.png', plot = subcortex, device = 'png', bg = 'white')
```

```{r Fig1E, eval=FALSE}
#Figure 1E (Subcortical ROI using ggseg3D)

#The function below plots 3d version of aseg. See "inputs" within function for more details. 
make_ggseg3d <- function(attribute, 
                         colour.pal = c("light yellow",
                                        "orange",
                                        "red",
                                        "dark red"), 
                         hide.colourbar=FALSE, 
                         output.png=FALSE, 
                         file.name="ggseg_3d.png") {
  # Inputs: 
  #' attribute = This must be a numeric vector in this: Left-Thalamus, Left-Caudate, Left
  #' -Putamen, Left-Pallidum Left-Hippocampus, Left-Amygdala, Left-Accumbens-area, Righ
  #' t-Thalamus-Proper, Right-Caudate, Right-Putamen, Right-Pallidum, Right-Hippocampus, Righ
  #' t-Amygdala, Right-Accumbens-area

  library(ggseg3d)
  library(tidyr)
  library(dplyr)
  #remove(aseg_3d)
  aseg_3d <- aseg_3d
  aseg_3d <- tidyr::unnest(aseg_3d, cols = c(ggseg_3d))

  attribute.ggseg3d <- c(rep(NA, 4), attribute[1:4], rep(NA, 3), attribute[5:7],
                         rep(NA, 5), attribute[8:14],  rep(NA, 6))

  data <- dplyr::mutate(aseg_3d, attribute =  attribute.ggseg3d)
  #remove NA regions
  aseg_3d[which(is.na(data$attribute)),] <- NA
  aseg_3d <- tidyr::drop_na(aseg_3d)

  data[which(is.na(data$attribute)),] <- NA
  data <- tidyr::drop_na(data)
  data$attribute[data$attribute==0]<-NA #make 0 valus NA so they are set as grey in ggseg3d

  scene=list(camera = list(eye = list(x = 0, y = 1, z = -2.25)),
             aspectratio = list(x=1.6,y=1.6,z=1.6))

  plot <- ggseg3d::ggseg3d(.data = data, 
                           atlas = aseg_3d, 
                           colour = "attribute", 
                           text = "attribute",
                           palette = colour.pal)
  plot <- remove_axes(plot)
  plot <- plotly::layout(plot, 
                         scene = scene, 
                         width = 600, height = 600)

  if(hide.colourbar==TRUE){
    plot <- plotly::hide_colorbar(plot)
  }

  if( output.png==TRUE){
    plotly::orca(plot, file =  file.name)
  }

  return(plot)

}

set.seed(1993)
volume <- sample(1:20, 14) #randomly generated data for 14 structures
make_ggseg3d(volume, 
             colour.pal = viridis::viridis(n=length(volume)), 
             output.png = T,
             file.name = "data/R/fig1e.png")

```

```{r Fig1F, eval=FALSE}
#Figure 1F (Cortical ROI using ggseg)

library(ggseg)
library(ggplot2)
#devtools::install_github("LCBC-UiO/ggsegGlasser")
library(ggsegGlasser)

#create dataframe with region names and hemisphere for the glasser atlas
base_atlas <- as.data.frame(na.omit(cbind(glasser$data$region, glasser$data$hemi)))

colnames(base_atlas) <- c("region", "hemi")

set.seed(1993) #set a random seed (good practice for reproducibility)
Effect <- rnorm(dim(base_atlas)[1]) #generate a random numbers for each roi in the atlas

base_atlas <- cbind(Effect, base_atlas)

#Cortex
cortex <- ggseg(atlas = glasser,
                .data = base_atlas,
                mapping=aes(fill=Effect), 
                position="stacked",
                colour="black",
                size=.2,
                show.legend = T,
                plot.background = "white") +
  scale_fill_viridis_b() + 
  theme_void() +
  theme(legend.position = "bottom",
        legend.text=element_text(size=15),
        legend.title  = element_blank())

ggsave(filename = 'data/R/fig1f.png', plot = cortex, device = 'png', bg = 'white')

```

```{r Fig1G, eval=FALSE}
#Figure 1G (Vertex using ciftiTools)

library(rgl)
library(ggplot2)

# Load the package and point to the Connectome Workbench
library(ciftiTools)

#Need connectome workbench to use ciftiTools - set path below
ciftiTools.setOption("wb_path", "/Applications/workbench/")

#Define CIFTI file names for time series and surface files. 
cifti_fname <- ciftiTools::ciftiTools.files()$cifti["dtseries"]
surfL_fname <- ciftiTools.files()$surf["left"]
surfR_fname <- ciftiTools.files()$surf["right"]

# Read in CIFTI file
cii <- read_cifti(
  cifti_fname, brainstructures="all", 
  surfL_fname=surfL_fname,
  surfR_fname=surfR_fname)

#Plot mean time series on CIFTI surface
view_xifti_surface(cii, 
                   hemisphere = "both", 
                   view = "lateral", 
                   idx = 1,
                   colors = viridis::plasma(n = 100),
                   zlim = c(1.2,1.5),  
                   legend_embed = T)

rgl.snapshot("data/R/fig1g.png") #save output in RGL window as 
rgl.close()

```

```{r Fig1H, eval=FALSE}
#Figure 1H (Vertex using fsbrain)

library(fsbrain)
library(rgl)

#download example freesurfer data which comes with the fsbrain package
#download_optional_data() 
#download_fsaverage(accept_freesurfer_license = TRUE)

subjects_dir <- get_optional_data_filepath("subjects_dir")

subject_id <- 'subject1'

colourmap <- colorRampPalette(viridis::plasma(n = 1000)) #set colourmap

# set figure size (optional)
fsbrain.set.default.figsize(700, 700)

# generate figure
brain = vis.subject.morph.native(subjects_dir, 
                                 subject_id, 'sulc', 
                                 cortex_only = T, 
                                 views=NULL, 
                                 draw_colorbar = FALSE,
                                 makecmap_options = list('colFn'=colourmap))

img = export(brain, draw_colorbar = FALSE, output_img = "data/R/fig1h.png") #write out figure

```

```{r Fig1I, eval=FALSE}
#Fig1I (Edges visualized using brainconn)

#devtools::install_github("sidchop/brainconn")
library(brainconn)

conmat <- example_weighted_undirected #one of the example matrices included with brainconn
degree <- rowSums(conmat)[which(rowSums(conmat)!=0)]

fig1i <- brainconn(atlas ="schaefer300_n7", 
                   conmat=conmat, 
                   node.size = degree/2,
                   view="ortho", 
                   edge.color.weighted = T,
                   background.alpha = 0.6, 
                   edge.color = scale_edge_colour_viridis(option = "C"))

ggsave(filename = 'data/R/fig1i.png', 
       plot = fig1i, 
       device = 'png',
       bg = 'white',
       height = 7, 
       units = "in")

```

\singlespacing

```{r Fig1 compile, eval=FALSE}
# This code block complies all the panels of the figure into one, largely using the patchwork package.

library(patchwork)
library(ggplot2)

fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g1', 'fig1g2', 'fig1g3','fig1h', 'fig1i')
figs <- list()

for (f in 1:length(fig_name)) { 
  knitr::plot_crop(paste0('data/R/',fig_name[f],'.png'))
  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/R/',fig_name[f],'.png')),
                                                             width=ggplot2::unit(1,"npc"),
                                                             height=ggplot2::unit(1,"npc")))
}

patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]]) + 
  plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]]) + 
                                           plot_layout(heights = c(0.5,1,1)))

patchwork + plot_annotation(tag_levels = 'A',
                            title = 'Examples of brain visualizations made in R') &
  theme(plot.tag = element_text(face = 'bold'),
        plot.title = element_text(face="bold"))

```

### 2022-09-04 SupplementA_R_Figs.Rmd
svg check

@@
 ```{r Fig1 compile, eval=FALSE}
-# This code block complies all the panels of the figure into one, largely using the patchwork package.
-
 library(patchwork)
@@
 
-patchwork <- figs[[1]]/figs[[2]]|((figs[[6]]/(figs[[8]]|figs[[9]]))/figs[[7]]) + 
-  plot_layout(heights = c(0.8,0.8,1.5))|((figs[[3]]/figs[[4]]/figs[[10]]) + 
-                                           plot_layout(heights = c(0.5,1,1)))
+patchwork <- figs[[1]]/figs[[2]]|((figs[[3]]/(figs[[4]]|figs[[5]]))/figs[[6]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[7]]/figs[[8]]/figs[[9]])+plot_layout(heights = c(0.5,1,1)))
 
+svg("data/R/Figure1.svg")
 patchwork + plot_annotation(tag_levels = 'A',
@@
         plot.title = element_text(face="bold"))
+dev.off()
 
 ```
+
+```{r Figure2, eval=FALSE}
+
+library(patchwork)
+library(ggplot2)
+
+fig_name <- c('fig2a', 'fig2b', 'fig2c', 'fig2d', 'fig2e', 'fig2f')
+figs <- list()
+
+for (f in 1:length(fig_name)) { 
+  knitr::plot_crop(paste0('data/python/',fig_name[f],'.png'))
+  figs[[f]] <- ggplot() +
+    annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/python/',fig_name[f],'.png')),
+                                       width=ggplot2::unit(1,"npc"),
+                                       height=ggplot2::unit(1,"npc")))
+}
+
+layoutA <- "
+AAAA
+#BB#
+CCCC
+"
+patchworkABC <- ((figs[[1]] + figs[[3]] + plot_layout(widths  = c(0.6,1))) / figs[[2]])  + plot_layout(heights = c(1,0.5))
+
+patchworkDEF <-  (figs[[4]] / figs[[5]] / figs[[6]] + plot_layout(design = layoutA, widths = c(0.2,1,0.2)))
+
+patchwork <- (patchworkABC| patchworkDEF)  + plot_layout(widths = c(1,0.6))
+
+#svg("data/R/Figure2.svg")
+patchwork + plot_annotation(tag_levels = 'A',
+                            title = 'Examples of brain visualizations made in Python') &
+  theme(plot.tag = element_text(face = 'bold'),
+        plot.title = element_text(face="bold"))
+#dev.off()
+
+```

### 2022-09-04 manuscript.Rmd
svg check

@@
 
-library(patchwork)
-library(ggplot2)
-
-fig_name <- c('fig1a', 'fig1b', 'fig1c', 'fig1d', 'fig1e', 'fig1f', 'fig1g', 'fig1h', 'fig1i')
-figs <- list()
-
-for (f in 1:length(fig_name)) { 
-  knitr::plot_crop(paste0('data/R/',fig_name[f],'.png'))
-  figs[[f]] <- ggplot() + annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/R/',fig_name[f],'.png')),
-                                                             width=ggplot2::unit(1,"npc"),
-                                                             height=ggplot2::unit(1,"npc")))
-}
-
-patchwork <- figs[[1]]/figs[[2]]|((figs[[3]]/(figs[[4]]|figs[[5]]))/figs[[6]])+plot_layout(heights = c(0.8,0.8,1.5))|((figs[[7]]/figs[[8]]/figs[[9]])+plot_layout(heights = c(0.5,1,1)))
-
-patchwork + plot_annotation(tag_levels = 'A',
-                            title = 'Examples of brain visualizations made in R') &
-  theme(plot.tag = element_text(face = 'bold'),
-        plot.title = element_text(face="bold"))
+knitr::include_graphics("data/R/Figure1.svg")
 
@@
 
-library(patchwork)
-library(ggplot2)
-
-fig_name <- c('fig2a', 'fig2b', 'fig2c', 'fig2d', 'fig2e', 'fig2f')
-figs <- list()
-
-for (f in 1:length(fig_name)) { 
-  knitr::plot_crop(paste0('data/python/',fig_name[f],'.png'))
-  figs[[f]] <- ggplot() +
-    annotation_custom(grid::rasterGrob(png::readPNG(paste0('data/python/',fig_name[f],'.png')),
-                                       width=ggplot2::unit(1,"npc"),
-                                       height=ggplot2::unit(1,"npc")))
-}
-
-layoutA <- "
-AAAA
-#BB#
-CCCC
-"
-patchwork1 <- (((figs[[1]] + figs[[3]]) + plot_layout(widths  = c(0.8,1))) / figs[[2]])  + plot_layout(heights = c(1,0.5)) 
-patchwork1
-
-patchwork2 <-  (figs[[4]] / figs[[5]] / figs[[6]]) + plot_layout(design = layoutA, widths = c(0.2,1,0.2))
-
-(patchwork1 | patchwork2) + plot_layout(widths = c(1,0.6))
-
-patchwork2 <-  (figs[[4]] / figs[[5]] / figs[[6]] + plot_layout(design = layoutA, widths = c(0.2,1,0.2)))
-
-patchwork2
-
-patchwork + plot_annotation(tag_levels = 'A',
-                            title = 'Examples of brain visualizations made in Python') &
-  theme(plot.tag = element_text(face = 'bold'),
-        plot.title = element_text(face="bold"))
+knitr::include_graphics("data/python/Figure2.pdf")
 

### 2022-09-05 manuscript.Rmd
file c-up

@@
 |                                                        | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
-|:------------------|:----------|:----------|:----------|:----------|:----------|
+|:------------|:-----------|:-----------|:-----------|:-----------|:-----------|
 | **R**                                                  |         |         |               |         |             |
@@
 | [fsbrain](https://github.com/dfsp-spirit/fsbrain)      | ***+*** | ***+*** | ***+*****\*** |         |             |
-| ggseg                                                  |         |         | ***+***       |         |             |
-| neurobase                                              | ***+*** |         |               |         |             |
-| oro.nifti                                              | ***+*** |         |               |         |             |
+| [ggseg](https://github.com/ggseg/ggseg)                                                  |         |         | ***+***       |         |             |
+| [neurobase](https://github.com/muschellij2/neurobase)                                              | ***+*** |         |               |         |             |
+| [oro.nifti](https://github.com/muschellij2/oro.nifti)                                              | ***+*** |         |               |         |             |
 | **Python**                                             |         |         |               |         |             |
-| ANTsPy                                                 | ***+*** | ***+*** | ***+***       |         |             |
-| brainiak                                               | ***+*** |         |               |         |             |
-| Brainplotlib                                           |         | ***+*** | ***+*****\*** |         |             |
-| Brainspace/surfplot                                    |         | ***+*** | ***+*****\*** |         |             |
-| DIPY                                                   | ***+*** |         |               |         | ***+***     |
-| ENIGMA toolbox                                         |         |         | ***+***       |         |             |
-| FSLeyes                                                | ***+*** | ***+*** | ***+***       |         | *+*         |
-| ggseg                                                  |         |         | ***+***       |         |             |
-| graphpype                                              |         |         |               | ***+*** |             |
-| MMVT                                                   |         | ***+*** | ***+***       | ***+*** |             |
-| MNE                                                    | ***+*** | ***+*** | ***+***       |         |             |
-| mrivis                                                 | ***+*** |         |               |         |             |
-| NaNSlice                                               | ***+*** |         |               |         |             |
-| netneurotools                                          |         | ***+*** | ***+*****\*** |         |             |
-| nilearn                                                | ***+*** | ***+*** | ***+***       | ***+*** |             |
+| [ANTsPy](https://github.com/ANTsX/ANTsPy)                                                 | ***+*** | ***+*** | ***+***       |         |             |
+| [brainiak](https://github.com/brainiak/brainiak)                                               | ***+*** |         |               |         |             |
+| [Brainplotlib](https://github.com/feilong/brainplotlib)                                           |         | ***+*** | ***+*****\*** |         |             |
+| [Brainspace](https://github.com/MICA-MNI/BrainSpace)/[surfplot](https://github.com/danjgale/surfplot)                                    |         | ***+*** | ***+*****\*** |         |             |
+| [DIPY](https://github.com/dipy/dipy)                                                   | ***+*** |         |               |         | ***+***     |
+| [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)                                         |         |         | ***+***       |         |             |
+| [FSLeyes](https://github.com/pauldmccarthy/fsleyes)                                                | ***+*** | ***+*** | ***+***       |         | *+*         |
+| [ggseg](https://github.com/ggseg/python-ggseg)                                                  |         |         | ***+***       |         |             |
+| [graphpype](https://neuropycon.github.io/graphpype/)                                              |         |         |               | ***+*** |             |
+| [MMVT](https://github.com/pelednoam/mmvt)                                                   |         | ***+*** | ***+***       | ***+*** |             |
+| [MNE](https://github.com/mne-tools/mne-python)                                                    | ***+*** | ***+*** | ***+***       |         |             |
+| [mrivis](https://github.com/raamana/mrivis)                                                 | ***+*** |         |               |         |             |
+| [NaNSlice](https://github.com/spinicist/nanslice)                                               | ***+*** |         |               |         |             |
+| [netneurotools](https://github.com/netneurolab/netneurotools)                                          |         | ***+*** | ***+*****\*** |         |             |
+| [nilearn](https://github.com/nilearn/nilearn)                                               | ***+*** | ***+*** | ***+***       | ***+*** |             |
 | niwidget                                               | ***+*** | ***+*** |               |         | ***+***     |
@@
 ```
+```{r Figure1, echo=FALSE, out.height = '70%', out.width = '80%', fig.align='center'}
 
-```{r Figure1, echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center'}
-
-knitr::include_graphics("data/R/Figure1.svg")
+show_fig <- function(f)
+  {if (knitr::is_latex_output())
+  {
+    output = xfun::with_ext(f, 'pdf')
+    rsvg::rsvg_pdf(xfun::with_ext(f,'svg'), file=output)
+  } else {
+    output = xfun::with_ext(f, 'svg')
+  }
+  knitr::include_graphics(output)
+}
+show_fig("./data/R/Figure1.svg")
 
@@
 Made using the `brainconn` function from the `brainconn` package.
-All code used to compile this figure, as well as the contents of each panel are provided in the Supplement.
+All code used to compile this figure, as well as the contents of each panel are provided in an accompanying [online repository](https://github.com/sidchop/RepoNeuroVis).
 
@@
 ```
-
-```{r Figure2, echo=FALSE, out.height = '80%', out.width = '80%', fig.align='center', eval=FALSE}
-
-knitr::include_graphics("data/python/Figure2.pdf")
-
+```{r Figure2, echo=FALSE, out.height = '60%', out.width = '80%', fig.align='center', dev='cairo_pdf'}
+knitr::include_graphics("./data/python/Figure2.pdf")
 ```
 
-*Figure 1. Examples of brain imaging visualization make in using different Python packages.* A)
+*Figure 2.
+Examples of brain imaging visualization make in using different Python packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown.
+Made using the `plotting.plot_stat_map` function from `nilearn`.
+B) Voxel-level cortical parcellation overlaid on T1-weighted MRI data shown on axial slices.
+Made using `plotting.plot_roi` function from `nilearn`.
+C) Medial and lateral views of an cortex-wide ROI atlas, displayed on an inflated template surface where a random value has been assigned to each ROI.
+Made using `Plot` function from `surfplot`.
+D) Lateral and medial views of vertex-level data displayed in a 3D rendering on an individual's white matter surface.
+Made using plotting.view_surf function in from `nilearn`.
+3D rendering of 16 subcortical structures from the the Deiskan-Killiany atlas, where a random value has been assigned to each region.
+Made using the `plot_subcortical` function from the `ENIGMA TOOLBOX`.A weighted and undirected graph plotted on front, right, and top views of a schematic outline of a brain in standardized coordinate space.
+Made using plotting.plot_connectome function in `nilearn`.All code used to generate the contents of each panel, and compile this figure are provided in an accompanying [online repository](https://github.com/sidchop/RepoNeuroVis).
 

### 2022-09-05 manuscript.Rmd
table fix

@@
   - Brain Visualization
-abstract: Neuroimaging visualization forms the centerpiece of quality control, and interpretation and communication of results.
+abstract: Neuroimaging visualization forms the centerpiece of interpretation and communication of scientific results, as well as data quality control.
 Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI).
-There now exist many well-documented code-based brain visualization tools that allow users to programmatically generate publication-ready figures directly within environments such as R, Python and MATLAB.
-Compared to figures generated using GUIs, programmatic tools are more replicable, flexible, interactive, and integrated with the scientific process.
-Here we outline the advantages of learning and using programmatic neuroimaging visualization tools, include a list of these tools across programming environments and give examples of code and visualizations at the voxel, vertex, region-of-interest, and edge-level data.
+There now exist many well-documented code-based brain visualization tools that allow users to programmatically generate publication-ready figures directly within R, Python and MATLAB environments.
+Compared to figures generated using GUIs, programmatic figures are more replicable, flexible, interactive, and integrated with the scientific process.
+Here we outline the advantages of using programmatic neuroimaging visualization tools, provide a list of avalible tools across programming environments and include examples of voxel, vertex, region-of-interest and edge-level visualizations.
 
@@
 - \newcommand{\elandscape}{\end{landscape}}
-- \usepackage{setspace}\doublespacing
+- \usepackage{setspace}\onehalfspacing
 ---
@@
 The present commentary argues for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
-We also providing didactic examples of visualizations and associated code as a reference point.
-
-In brief, by generating visualizations using code, you increase the replicability of your figures for yourself, your collaborators and your readers.
-Moreover, code-based tools provides more precise controls over display settings, while also benefiting from the advantages of programming, such as the ability to iteratively and rapidly generate multiple figures.
-Finally, being able to integrate figure generation into your analysis scripts reduces chances of errors and increases the accessibility of resulting analyses pipeline, with more advanced tools now allowing for the seamless integration of prose and code.
-Here, we review the advantages of learning and using programmatic neuroimaging visualizations, focusing on benefits to replicability, flexibility and integration.
-We conclude by introducing examples of R-packages which allow for visualization of statistics at region-of-interest (ROI), voxel, vertex, and edge level data, followed by a brief discussion of limitations and functionality gaps in code-based brain visualization tools.
+We also providing didactic examples of visualizations and associated code as a point of reference.
 
@@
 In practice, this typically entails openly sharing detailed methods, materials, code, and data.
-While there is a trend towards increasing transparency and code sharing of neuroimaging analyses, the sharing of code used to generate figures which include brain renderings and spatial maps has been relatively neglected.
+While there is a trend towards increasing transparency and code sharing of neuroimaging analyses, the sharing of code used to generate figures such as brain renderings and spatial maps has been relatively neglected.
 This gap in reproducibility is partly driven by the fact that brain figures are often created using a manual process that involves tinkering with sliders, buttons, and overlays on a GUI, concluding with a screenshot and sometimes beautification in image processing software like Illustrator, Photoshop or Inkscape.
-Such a process inherently makes neuroimaging visualizations difficult, if not impossible to replicate, even by the authors themselves.
+Such a process inherently makes neuroimaging visualizations difficult, if not impossible to replicate at times, even by the authors themselves.
 
 Visualization scripts should reflect a core feature of open science.
-Given that brain figures regularly form the centerpiece of interpretation within papers, conference presentations, or news reports, making sure they can be reliably regenerated is crucial for knowledge generation.
+Given that brain figures regularly form the centerpiece of interpretation within papers, conference presentations, or news reports, making sure they can be reliably regenerated is crucial for knowledge generation and dissemination.
 By writing and sharing code used to generate brain visualizations, a direct and traceable link is established between the underlying data and the corresponding scientific figure.
-While the code that produces a replicable figure doesn't necessarily reflect the validity of the scientific finding or the accuracy of the associated content, it allows for reproducibility, instilling transparency and robustness, while demonstrating a desire to further scientific knowledge.
+While this code doesn't necessarily reflect the validity or accuracy of the scientific finding, it allows for reproducibility, instilling transparency and robustness, while demonstrating a desire to further scientific knowledge.
 Some even consider publishing figures which cannot be replicated as closer to advertising, rather than science (Steel, 2013).
@@
 
-Being able to exactly replicate your figures via code has marked advantages beyond positive open science practices.
+Being able to exactly replicate your figures via code has marked advantages beyond open science practices.
 In particular, the ability to reprogram inputs (such as statistical maps) and settings (such as color schemes, thresholds, and visual orientations) can streamline your entire scientific workflow.
@@
 Nonetheless, part of the appeal of GUI-based tools is that the presets for such settings can provide a useful starting point and reduce the decision burden on novice users.
-However, similar presets are often available in the form of default settings across most code-based packages, negating the need for the user to manually enter each and every choice of setting required for creating an image.
-Most code-based tools also come with documentation, with R-packages on the *CRAN* or *Neuroconductor* (Muschelli et al., 2019) repositories requiring detailed guidance.
-Recent packages have started to include detailed documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, & Mejia, 2022; Mowinckel & Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer & Ecker, 2020) which provide examples of figures that can be used as starting points or templates for new users.
+However, similar presets are often available in the form of default settings across most code-based packages, negating the need for the user to manually enter each and every choice required for creating an image.
+Most code-based tools also come with documentation, with R-packages on the *CRAN* or *Neuroconductor* (Muschelli et al., 2019) repositories strictly requiring detailed guidance.
+Recent packages have started to include detailed beginner-friendly documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, & Mejia, 2022; Mowinckel & Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer & Ecker, 2020) which provide examples of figures that can be used as starting points or templates for new users.
 As the popularity of code-sharing for figure increases, there will be a cornucopia of templates that can be used as the basis for new figures.
@@
 Tools to automatically detect artefacts, de-noise the data and generate derivatives are becoming more robust, but we are not yet at the stage where visualizing the data during processing is no longer necessary.
-Nonetheless, when working with large datasets such as Human Connectome Project (Van Essen et al., 2013) or UK BioBank (Sudlow et al., 2015), it is unfeasible to use traditional GUI-based tool to visually examine the data.
+Nonetheless, when working with large datasets such as Human Connectome Project (Van Essen et al., 2013) or UK BioBank (Sudlow et al., 2015), it is unfeasible to use traditional GUI-based tools to visually examine the data.
 The time it takes to open a single file and achieve the desired visualization settings vastly compounds when working with large datasets.
-Knowing how to programmatically generate brain visualizations can allow you to iterate your visualization code over each image of a large datasets making checking the quality of each data processing or analysis step accessible and achievable.
+Knowing how to programmatically generate brain visualizations can allow you to iterate your visualization code over each image of a large datasets making checking the quality of each data processing step achievable.
 The visual outputs of each iteration can be complied into accessible documents which can be easily scrolled, with more advanced usage allowing for the creation of interactive HTML reports, similar to those created by standardized data processing tools like *fmriprep* (Esteban et al., 2019).
-Increasing capacity to conduct visual quality control on larger datasets can increase the identification of processing errors and result in more reliable and valid findings from your data.
+Increasing capacity to conduct visual quality control on larger datasets will improve the identification of processing errors and result in more reliable and valid findings.
 
@@
 
-Often programming languages such as R, Python and MATLAB are used for the analysis and non-brain visualizations in neuroimaging studies, but the brain visualizations resulting from these analyses are outsourced to separate GUI's of tools such as *FSLeyes*, *Freeview* or *ITK-snap*.
+Often in neuroimaging studies, programming languages such as R, Python and MATLAB are used for stastical analysis and generating non-brain figures, but the brain figures are outsourced to separate GUI's based tools such as *FSLeyes*, *Freeview* or *ITK-snap*.
 Switching from your analysis environment to a GUI-based visualization process can be a cumbersome deviation from the scientific workflow.
 This can make debugging errors more difficult, as you have to regularly switch program to visually examine the results of any modifications or adjustments to prior analyses.
-Using the brain visualization tools that already exist within your chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analyses.
+Using the brain visualization tools that already exist within your chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analysis.
 
@@
 Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g.
-*e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code computational environment linked to the associated article using 'compute capsules' from the cloud-based platform Code Ocean (Code Ocean, 2021).
+*e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code linked to the associated article using 'compute capsules' from the cloud-based platform Code Ocean (Code Ocean, 2021).
 
-Neuroimaging data are often spatially 3D and can have multiple time points, adding a 4th dimension (e.g., fMRI data).
-Thus, communicating findings or evaluating quality using static 2D slices is challenging, and may not be the best representation of the data, or the interpretations.
+Neuroimaging data are often spatially 3D and can have multiple time points, adding a 4th dimension (e.g., functional imaging data).
+Thus, communicating findings or evaluating quality using static 2D slices is challenging, and may not be the best representation of the data, or the associted interpretations.
 While well-curated 3D renderings can help with spatial localisation (see Madan, 2015; Pernet & Madan, 2019), in the end, static images can only provide an incomplete representation of the data, and forces researchers to choose the "best" angle or slice to show, which often involves compromising one result to emphasize another.
-An added advantage of some of the code-based tools is that you can generate 'rich' media like interactive widgets -- figures or animations, which allow users to zoom, rotate and scroll through slices.
+An added advantage of some of the code-based tools is that you can generate 'rich' media like interactive figures or animations, which allow users to zoom, rotate and scroll through slices.
 Interacting with a figure in this way can improve scientific communication of findings.
@@
 Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of your research with peers and reach larger audiences (Li and Xie, 2020).
-This last point is becoming increasingly salient as marketing science on social media has become a core medium for spreading discoveries, science communication to the public, and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019).
+This last point is becoming increasingly salient as promoting science on social media has become a core medium for spreading discoveries, science communication to the public, and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019).
 Overall, public engagement is one of the cornerstones of science, and the images we create are at the center of the process.
@@
 |                                                                                                         | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
-|:------------|:-----------|:-----------|:-----------|:-----------|:-----------|
+|:-------------|:-----------|:-----------|:-----------|:-----------|:-----------|
 | **R**                                                                                                   |         |         |               |         |             |
@@
 |   [netneurotools](https://github.com/netneurolab/netneurotools)                                         |         | ***+*** | ***+*****\*** |         |             |
+|   [netplotbrain](https://github.com/wiheto/netplotbrain)                                                |         |         | ***+***       | ***+*** |             |
 |   [nilearn](https://github.com/nilearn/nilearn)                                                         | ***+*** | ***+*** | ***+***       | ***+*** |             |
-| niwidget                                               | ***+*** | ***+*** |               |         | ***+***     |
-| Pycortex                                               | ***+*** | ***+*** | ***+*****\*** |         |             |
-| pySurfer                                               |         | ***+*** | ***+*****\*** |         |             |
-| surfice                                                | ***+*** | ***+*** | ***+***       | ***+*** | ***+***     |
-| surfplot                                               |         | ***+*** | ***+***       |         |             |
-| Visbrain                                               | ***+*** | ***+*** | ***+***       | ***+*** |             |
+|   [niwidget](https://github.com/nipy/niwidgets)                                                         | ***+*** | ***+*** |               |         | ***+***     |
+|   [Pycortex](https://github.com/gallantlab/pycortex)                                                    | ***+*** | ***+*** | ***+*****\*** |         |             |
+|   [pySurfer](https://pysurfer.github.io/)                                                               |         | ***+*** | ***+*****\*** |         |             |
+|   [surfice](https://github.com/neurolabusc/surf-ice)                                                    | ***+*** | ***+*** | ***+***       | ***+*** | ***+***     |
+|   [Visbrain](https://github.com/EtienneCmb/visbrain)                                                    | ***+*** | ***+*** | ***+***       | ***+*** |             |
 | **MATLAB**                                                                                              |         |         |               |         |             |
-| BrainNetViewer                                         | ***+*** |         | ***+***       | ***+*** |             |
-| Brainspace                                             |         | ***+*** | ***+*****\*** |         |             |
-| Brainstorm                                             | ***+*** | ***+*** | ***+***       |         |             |
-| bspmview                                               | ***+*** |         | ***+***       |         |             |
-| CandlabCore                                            | ***+*** |         | ***+***       |         |             |
-| ECoG/fMRI Vis toolbox                                  |         | ***+*** | ***+*****\*** |         |             |
-| ENIGMA toolbox                                         |         |         | ***+***       |         |             |
-| FieldTrip                                              | ***+*** | ***+*** |               |         |             |
-| Lead-DBS                                               | ***+*** |         | ***+***       |         |             |
-| mni2fs                                                 |         | ***+*** |               |         |             |
-| mrtools                                                | ***+*** | ***+*** | ***+*****\*** |         |             |
-| plotSurfaceROIBoundary                                 |         | ***+*** | ***+*****\*** |         |             |
-| Vistasoft                                              | ***+*** | ***+*** | ***+***       |         | ***+***     |
+|   [BrainNetViewer](https://www.nitrc.org/projects/bnv/)                                                 | ***+*** |         | ***+***       | ***+*** |             |
+|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)                                                  |         | ***+*** | ***+*****\*** |         |             |
+|   [Brainstorm](https://github.com/brainstorm-tools/brainstorm3)                                         | ***+*** | ***+*** | ***+***       |         |             |
+|   [bspmview](https://www.bobspunt.com/software/bspmview/)                                               | ***+*** |         | ***+***       |         |             |
+|   [CandlabCore](https://github.com/canlab/CanlabCore)                                                   | ***+*** |         | ***+***       |         |             |
+|   [ECoG/fMRI Vis toolbox](https://github.com/Immiora/matlab-ecog-visualization)                         |         | ***+*** | ***+*****\*** |         |             |
+|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)                                                  |         |         | ***+***       |         |             |
+|   [FieldTrip](https://www.fieldtriptoolbox.org/)                                                        | ***+*** | ***+*** |               |         |             |
+|   [Lead-DBS](https://www.lead-dbs.org/)                                                                 | ***+*** |         | ***+***       |         |             |
+|   [mni2fs](https://github.com/dprice80/mni2fs)                                                          |         | ***+*** |               |         |             |
+|   [mrtools](https://github.com/justingardner/mrTools)                                                   | ***+*** | ***+*** | ***+*****\*** |         |             |
+|   [plotSurfaceROIBoundary](https://github.com/StuartJO/plotSurfaceROIBoundary)                          |         | ***+*** | ***+*****\*** |         |             |
+|   [Vistasoft](https://github.com/vistalab/vistasoft)                                                    | ***+*** | ***+*** | ***+***       |         | ***+***     |
 
@@
 This list does not include cross-platform general purpose visualization software.\
-***\****Cortex only
+***\**** Cortex only
 
@@
 
-The following four sections provide brief examples of well documented and beginner friendly packages and functions available in R and Python for visualizing voxel, vertex, ROI and edge-level data.
-This is not an exhaustive list of packages available for visualizing brain data in R and Python (See Table 1), rather, the following sections aim to give the reader a sense of the available options, and an entry point to using them.
-All code used to compile the figures, as well as the contents of each panel are provided in the Supplement or can be accessed in the accompanying Git repository.
+The following two figures provide examples of voxel, vertex, ROI and edge-level figures generated within R (Figure 1) and Python (Figure 2) using open source, well documented and beginner-friendly packages.
+These are not an exhaustive representation of packages available for visualizing brain data in R and Python (See Table 1), rather, the figures aim to give the reader a sense of the available options, and an entry point to using them.
+All code used to compile the figures, as well as the contents of each panel are provided an accompanying [online repository.](https://github.com/sidchop/RepoNeuroVis)
 
@@
 ```
-```{r Figure1, echo=FALSE, out.height = '70%', out.width = '80%', fig.align='center'}
+```{r Figure1, echo=FALSE, out.height = '60%', out.width = '70%', fig.align='center'}
 
-show_fig <- function(f)
+show_fig <- function(f) # function to convert and embed svg file
   {if (knitr::is_latex_output())
@@
 
-*Figure 1.
-Examples of brain imaging visualization make in using different R packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown.
+***Figure 1.
+Examples of brain imaging visualizations made using R.*** A) Voxel-level statistical map thresholded and overlaid over a T1-weighted template image, with a single axial slice shown.
 Made using the the `ortho2` function from the `neurobase` package.
-B) A voxel-level cortical parcellation overlaid on T1-weighted MRI data shown in 9-slice axial orientation.
+B) A voxel-level cortical parcellation overlaid on a individual T1-weighted image, shown in 9-slice axial orientation.
 Made using the `overlay` function from the `neurobase` package.
-C) A CIFTI format surface ROI atlas with a corresponding statistic assigned to each regions, with both hemispheres displayed on a inflated template surface in lateral view.
+C) A CIFTI format surface ROI atlas with a corresponding statistic assigned to each region, with both hemispheres displayed on a inflated template surface in lateral view.
 Made using the `view_xifti_surface` from the `ciftiTools` package.
-D) A coronal cross-sectional rendering of subcortical structures where a statistical value has been assigned to each region.
+D) A coronal cross-sectional rendering of subcortical structures where a value has been assigned to each region.
 Made using the `aseg` atlas from the `ggseg` package.
-E) A 3D rendering of 9 bilateral subcortical regions where a statistical value has been assigned to each region.
+E) A 3D rendering of 9 bilateral subcortical regions where a value has been assigned to each region.
 Made using the `aseg` atlas from the `ggseg3d` package.
-F) Medial and lateral views of a ROI atlas displayed on inflated template cortical surface where a statistical value has been assigned to each region.
+F) Medial and lateral views of a ROI atlas displayed on inflated cortical surface where a value has been assigned to each region.
 Made using the `glasser` atlas from the `ggsegGlasser` package, which was plotted using `ggseg`.
@@
 Made using the `vis.subject.morph.standard` function from the `fsbrain` package.
-I) A weighted and undirected graph plotted on top, left and front views of a template of a brain rendering.
+I) A weighted and undirected graph plotted on top, left and front views of a schematic outline of a brain in MNI coordinate space.
 Made using the `brainconn` function from the `brainconn` package.
@@
 
-*Figure 2.
-Examples of brain imaging visualization make in using different Python packages.* A) Voxel-based statistical map threshold and overlaid over a T1-weighted template data, with a single axial slice shown.
+***Figure 2.
+Examples of brain imaging visualizations made using Python.*** A) Voxel-based statistical map thresholded and overlaid over a T1-weighted template data, with a single axial slice shown.
 Made using the `plotting.plot_stat_map` function from `nilearn`.
-B) Voxel-level cortical parcellation overlaid on T1-weighted MRI data shown on axial slices.
+B) Voxel-level cortical parcellation overlaid on T1-weighted MRI data shown in four axial slices.
 Made using `plotting.plot_roi` function from `nilearn`.
-C) Medial and lateral views of an cortex-wide ROI atlas, displayed on an inflated template surface where a random value has been assigned to each ROI.
+C) Medial and lateral views of an cortex-wide ROI atlas, displayed on an inflated template surface where a statistical value has been randomly assigned to each ROI.
 Made using `Plot` function from `surfplot`.
@@
 Made using plotting.view_surf function in from `nilearn`.
-3D rendering of 16 subcortical structures from the the Deiskan-Killiany atlas, where a random value has been assigned to each region.
-Made using the `plot_subcortical` function from the `ENIGMA TOOLBOX`.A weighted and undirected graph plotted on front, right, and top views of a schematic outline of a brain in standardized coordinate space.
-Made using plotting.plot_connectome function in `nilearn`.All code used to generate the contents of each panel, and compile this figure are provided in an accompanying [online repository](https://github.com/sidchop/RepoNeuroVis).
+3D rendering of 16 subcortical structures from the the Deiskan-Killiany atlas, where a statistical value has been randomly assigned to each region.
+Made using the `plot_subcortical` function from the `ENIGMA TOOLBOX`.
+A weighted and undirected graph plotted on front, right, and top views of a schematic outline of a brain in MNI coordinate space.
+Made using plotting.plot_connectome function in `nilearn`.
+All code used to generate the contents of each panel, and compile this figure are provided in an accompanying [online repository](https://github.com/sidchop/RepoNeuroVis).
 
@@
 
-Most of the tools introduced above do not require a strong knowledge of programming, but there is still a steeper learning curve compared to using a GUI.
-This is especially true when learning how to make fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning, for the purpose of a publication-ready figure.
+While many code-based tools are well documented and do not require a strong knowledge of programming, there can still be a steeper learning curve for new users, compared to using a GUI.
+This is especially true for the purpose of a publication-ready figure, were fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning are needed.
 While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability.
-Relatedly, while some interactive image viewers which can be opened within an integrated development environment (e.g.
+Relatedly, while some interactive image viewers which can be opened within an integrated development environment like R-Studio (e.g.
 Muschelli, 2016), for quick and interactive viewing of single images, GUI tool can be faster and more practical.
@@
 Likewise, custom non-cortical atlases such as non-standard subcortical atlas schemes are not yet straightforward, and usually require multiple functions and packages to visualize.
-Finally, some neuroimaging related data types, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools.
+Finally, some neuroimaging derived data types, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools.
 

### 2022-09-05 manuscript.Rmd
title update

@@
 Compared to figures generated using GUIs, programmatic figures are more replicable, flexible, interactive, and integrated with the scientific process.
-Here we outline the advantages of using programmatic neuroimaging visualization tools, provide a list of avalible tools across programming environments and include examples of voxel, vertex, region-of-interest and edge-level visualizations.
+Here we outline the advantages of using programmatic neuroimaging visualization tools, provide a list of available tools across programming environments and include examples of voxel, vertex, region-of-interest and edge-level visualizations.
 

### 2022-09-10 manuscript.Rmd
.

@@
 These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments.
-The present commentary argues for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
+Here we review the rational for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
 We also providing didactic examples of visualizations and associated code as a point of reference.
@@
 While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability.
-Relatedly, while some interactive image viewers which can be opened within an integrated development environment like R-Studio (e.g.
+Relatedly, while some interactive image viewers can be opened within an integrated development environment like R-Studio (e.g.
 Muschelli, 2016), for quick and interactive viewing of single images, GUI tool can be faster and more practical.

### 2023-02-10 manuscript.Rmd
fig update

@@
 ---
-title: Moving Towards Reproducible and Programmatic Generation of Neuroimaging Visualizations
+title: A Practical Guide for Generating Reproducible and Programmatic Neuroimaging Visualizations
 authors:
@@
   - Brain Visualization
-abstract: Neuroimaging visualization forms the centerpiece of interpretation and communication of scientific results, as well as data quality control.
-Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUI).
+abstract: Neuroimaging visualizations form the centerpiece of the interpretation and communication of scientific results, and are also a cornerstone for data quality control.
+Often, these images and figures are produced by manually changing settings on Graphical User Interfaces (GUIs).
 There now exist many well-documented code-based brain visualization tools that allow users to programmatically generate publication-ready figures directly within R, Python and MATLAB environments.
-Compared to figures generated using GUIs, programmatic figures are more replicable, flexible, interactive, and integrated with the scientific process.
-Here we outline the advantages of using programmatic neuroimaging visualization tools, provide a list of available tools across programming environments and include examples of voxel, vertex, region-of-interest and edge-level visualizations.
-
+Here, we provide a rationale for the wide-spread adoption of code-generated brain visualizations by highlighting corresponding advantages in replicability, flexibility, and integration over GUI based tools.
+We then provide a practical guide outlining the steps required to generate these code-based brain visualizations.
+We also present a comprehensive table of tools currently available for programmatic brain visualizations and provide didactic examples of visualizations and associated code as a point of reference.
+Finally, we provide a web-app (https://sidchop.shinyapps.io/braincode/) which can generate simple code-templates as starting points for these visualizations.
 output: 
+  bookdown::word_document2:
+    toc: false
   bookdown::pdf_book:
@@
     latex_engine: lualatex
-  bookdown::word_document2:
-    toc: false
-linestretch: 2
+linestretch: 1.5
+geometry: "left=1.5cm, right=1.5cm, top=1.5cm, bottom=1.5cm"
 header-includes:
@@
 - \usepackage{setspace}\onehalfspacing
+- \usepackage[left]{lineno}
+- \modulolinenumbers[5]
+- \linenumbers
 ---
@@
 In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility.
-However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community (Table 1).
-These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code (Figure 1), especially within the R, Python and MATLAB environments.
-Here we review the rational for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
-We also providing didactic examples of visualizations and associated code as a point of reference.
+However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community.
+These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code, especially within the R, Python and MATLAB environments.
+Here, we first discuss the rational for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
+We then provide a practical guide outlining the steps required to make code-based brain visualizations and provide a web-app (<https://sidchop.shinyapps.io/braincode/>; Figure 4) which can generate simple code-templates as starting points for these visualizationds.
+We also present a comprehensive table of tools currently available for programmatic brain visualizations (Table 1) and provide didactic examples of visualizations and associated code as a point of reference (Figure 2-3).
+The focus of this guide is on human brain magnetic resonance imaging (MRI) data, but many of the principals discussed and tools provided will equally apply to visualizing data from other organs and imaging modalities such as EEG, MEG, PET and CT.
 
-# Replicability
+# Benefits of learning to generate code-based brain visualizations
+
+## Replicability
 
@@
 
-Visualization scripts should reflect a core feature of open science.
+Code used for data visualization should reflect a core feature of open science.
 Given that brain figures regularly form the centerpiece of interpretation within papers, conference presentations, or news reports, making sure they can be reliably regenerated is crucial for knowledge generation and dissemination.
@@
 
-While some GUI-based tools have historically offered command-line access to generate replicable visualizations, they can lack both the flexibility to easily generate publication ready figures and the benefits, such as iteration, provided by your preferred programming environment.
-Likewise, other GUI-based tools offer replicability in the form of automatically generated batch scripts or in-built terminals, which are often idiosyncratic and lack documentation to make them easily usable or reproducible by those not familiar with the specific software.
+Notably, some GUI-based tools have historically offered command-line access to generate replicable visualizations (e.g., FreeView, FSLeyes, surfice), making their use potentially equally replicable to purely code-based tools.
+Use of these specialized command-line interfaces can provide a useful middle ground of those who have little experience with coding environments.
+Nonetheless, these interfaces often still have a learning curve, but can lack other advantages, such as iteration, provided by the preferred programming environment (see Sections \@ref(flexibility) and \@ref(integrative-and-interactive-reporting)).
+Likewise, other GUI-based tools offer replicability in the form of automatically generated batch scripts (text files containing lines of specialized commands that can be re-executed) or in-built terminals, which are often idiosyncratic and may lack documentation to make them easily usable or replicable by those not familiar with the specific software.
 
-# Flexibility
+## Flexibility
 
-Being able to exactly replicate your figures via code has marked advantages beyond open science practices.
-In particular, the ability to reprogram inputs (such as statistical maps) and settings (such as color schemes, thresholds, and visual orientations) can streamline your entire scientific workflow.
+Being able to exactly replicate a figures via code has marked advantages beyond open science practices.
+In particular, the ability to reprogram inputs (such as statistical maps) and settings (such as color schemes, thresholds, and visual orientations) can streamline the entire scientific workflow.
 Changing inputs and settings via code allows for the easy production of multiple figures, such as those resulting from multiple analyses which require similar visualizations.
@@
 This also makes it easier to generate consistent figures across subsequent projects.
-Keep in mind that the gains of writing code for your figures are cumulative, and in addition to improving your programming, you start to build a code-base for figure generation that you can continue to reuse and share throughout your scientific career.
+Keep in mind that the gains of writing code for figures are cumulative, and in addition to improving programming skills, one can build a code-base for figure generation that can continue to be reused and shared throughout a scientific career.
 
-Precise controls via code over visualization settings, such as color schemes, legend placement and camera angles, can provide you with much greater flexibility over visualizations.
+Precise controls via code over visualization settings, such as color schemes, legend placement and camera angles, can provide much greater flexibility over visualizations.
 Nonetheless, part of the appeal of GUI-based tools is that the presets for such settings can provide a useful starting point and reduce the decision burden on novice users.
@@
 Most code-based tools also come with documentation, with R-packages on the *CRAN* or *Neuroconductor* (Muschelli et al., 2019) repositories strictly requiring detailed guidance.
-Recent packages have started to include detailed beginner-friendly documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, & Mejia, 2022; Mowinckel & Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer & Ecker, 2020) which provide examples of figures that can be used as starting points or templates for new users.
+Recent packages have started to include detailed beginner-friendly documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, & Mejia, 2022; Mowinckel & Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer & Ecker, 2020) which provide examples of figures that can be used as starting points or templates for new users (also see Section\@ref(brain-code-a-web-app-for-generating-code-templates-for-brain-visualizations)).
 As the popularity of code-sharing for figure increases, there will be a cornucopia of templates that can be used as the basis for new figures.
@@
 Tools to automatically detect artefacts, de-noise the data and generate derivatives are becoming more robust, but we are not yet at the stage where visualizing the data during processing is no longer necessary.
-Nonetheless, when working with large datasets such as Human Connectome Project (Van Essen et al., 2013) or UK BioBank (Sudlow et al., 2015), it is unfeasible to use traditional GUI-based tools to visually examine the data.
+Nonetheless, when working with large datasets such as Human Connectome Project (Van Essen et al., 2013) or UK BioBank (Sudlow et al., 2015), it is simply not feasible to use traditional GUI-based tools to visually examine the data.
 The time it takes to open a single file and achieve the desired visualization settings vastly compounds when working with large datasets.
-Knowing how to programmatically generate brain visualizations can allow you to iterate your visualization code over each image of a large datasets making checking the quality of each data processing step achievable.
-The visual outputs of each iteration can be complied into accessible documents which can be easily scrolled, with more advanced usage allowing for the creation of interactive HTML reports, similar to those created by standardized data processing tools like *fmriprep* (Esteban et al., 2019).
+Knowing how to programmatically generate brain visualizations can allow for iteration of visualization code over each image of a large datasets making checking the quality of each data processing step achievable.
+The visual outputs of each iteration can be complied into accessible documents which can be easily scrolled, with more advanced usage allowing for the creation of interactive HTML reports (see Section \@ref(generate-and-share-visualization)), similar to those created by standardized data processing tools like *fmriprep* (Esteban et al., 2019).
 Increasing capacity to conduct visual quality control on larger datasets will improve the identification of processing errors and result in more reliable and valid findings.
 
-# Integrative and Interactive Reporting
+## Integrative and Interactive Reporting
 
-Often in neuroimaging studies, programming languages such as R, Python and MATLAB are used for stastical analysis and generating non-brain figures, but the brain figures are outsourced to separate GUI's based tools such as *FSLeyes*, *Freeview* or *ITK-snap*.
-Switching from your analysis environment to a GUI-based visualization process can be a cumbersome deviation from the scientific workflow.
-This can make debugging errors more difficult, as you have to regularly switch program to visually examine the results of any modifications or adjustments to prior analyses.
-Using the brain visualization tools that already exist within your chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analysis.
-
+Often in neuroimaging studies, programming languages such as R, Python and MATLAB are used for statistical analysis and generating non-brain figures, but the brain figures are outsourced to separate GUI-based tools such as *FSLeyes*, *Freeview* or *ITK-snap*.
 Increasingly popular software such as *R Markdown*, *Quarto* and *Jupyter Notebook* allow for the mixing of prose and code in a single script, resulting in fully reproducible and publication ready papers.
-By using code-based tools available within your preferred environment, brain visualizations can be directly integrated and embedded within a paper or report.
+By using code-based tools available within the preferred environment, brain visualizations can be directly integrated and embedded within a paper or report.
 For instance, a fully reproducible version of the current paper can be found on [GitHub](https://github.com/sidchop/RepoNeuroVis).
 Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g.
-*e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code linked to the associated article using 'compute capsules' from the cloud-based platform Code Ocean (Code Ocean, 2021).
+*e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code linked to the associated article via cloud-based platforms like *Code Ocean* (Code Ocean, 2021).
 
 Neuroimaging data are often spatially 3D and can have multiple time points, adding a 4th dimension (e.g., functional imaging data).
-Thus, communicating findings or evaluating quality using static 2D slices is challenging, and may not be the best representation of the data, or the associted interpretations.
-While well-curated 3D renderings can help with spatial localisation (see Madan, 2015; Pernet & Madan, 2019), in the end, static images can only provide an incomplete representation of the data, and forces researchers to choose the "best" angle or slice to show, which often involves compromising one result to emphasize another.
-An added advantage of some of the code-based tools is that you can generate 'rich' media like interactive figures or animations, which allow users to zoom, rotate and scroll through slices.
+Thus, communicating findings or evaluating quality using static 2D slices is challenging, and may not be the best representation of the data, or the associated interpretations.
+While well-curated 3D renderings can help with spatial localisation (see Madan, 2015; Pernet & Madan, 2019), in the end, static images can only provide an incomplete representation of the data, and they force researchers to choose the "best" angle or slice to show, which often involves compromising one result to emphasize another.
+An added advantage of some of the code-based tools is the generation of 'rich' media like interactive figures or animations, which allow users to zoom, rotate and scroll through slices.
 Interacting with a figure in this way can improve scientific communication of findings.
-Linking to or even embedding these videos or interactive figures in papers can greatly enhance the communication of findings and make your paper more engaging for the reader.
-Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of your research with peers and reach larger audiences (Li and Xie, 2020).
-This last point is becoming increasingly salient as promoting science on social media has become a core medium for spreading discoveries, science communication to the public, and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019).
+Linking to or even embedding these videos or interactive figures in papers can greatly enhance the communication of findings and make papers more engaging for the reader.
+Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of research with peers and reach larger audiences (Li and Xie, 2020).
+This last point is becoming increasingly salient as promoting science on social media has become a core medium for spreading discoveries via science communication to the public (Mueller-Herbst, et al., 2020; Smith & Seitz, 2019; Huber et al., 2019), to the research community (Luc et al., 2021; Quintana, 2020a), and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019).
 Overall, public engagement is one of the cornerstones of science, and the images we create are at the center of the process.
 
-\
-**Table 1. Examples of code-based neuroimaging visualizations tools that can be accessed directly within R, MATLAB and Python environments.**
+# Generating Code-based Visualizations
 
-|                                                                                                         | Voxel   | Vertex  | ROI           | Edge    | Streamlines |
-|:-------------|:-----------|:-----------|:-----------|:-----------|:-----------|
-| **R**                                                                                                   |         |         |               |         |             |
-|   [ANTsR](https://github.com/ANTsX/ANTsR)                                                               | ***+*** | ***+*** | ***+***       |         |             |
-|   [brainconn](https://github.com/sidchop/brainconn)                                                     |         |         |               | ***+*** |             |
-|   [brainR](https://github.com/muschellij2/brainR)                                                       | ***+*** |         | ***+***       |         |             |
-|   [ciftitools](https://github.com/mandymejia/ciftiTools)                                                | ***+*** | ***+*** | ***+*****\*** |         |             |
-|   [fsbrain](https://github.com/dfsp-spirit/fsbrain)                                                     | ***+*** | ***+*** | ***+*****\*** |         |             |
-|   [ggseg](https://github.com/ggseg/ggseg)                                                               |         |         | ***+***       |         |             |
-|   [neurobase](https://github.com/muschellij2/neurobase)                                                 | ***+*** |         |               |         |             |
-|   [oro.nifti](https://github.com/muschellij2/oro.nifti)                                                 | ***+*** |         |               |         |             |
-| **Python**                                                                                              |         |         |               |         |             |
-|   [ANTsPy](https://github.com/ANTsX/ANTsPy)                                                             | ***+*** | ***+*** | ***+***       |         |             |
-|   [brainiak](https://github.com/brainiak/brainiak)                                                      | ***+*** |         |               |         |             |
-|   [Brainplotlib](https://github.com/feilong/brainplotlib)                                               |         | ***+*** | ***+*****\*** |         |             |
-|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)/[surfplot](https://github.com/danjgale/surfplot) |         | ***+*** | ***+*****\*** |         |             |
-|   [DIPY](https://github.com/dipy/dipy)                                                                  | ***+*** |         |               |         | ***+***     |
-|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)                                                  |         |         | ***+***       |         |             |
-|   [FSLeyes](https://github.com/pauldmccarthy/fsleyes)                                                   | ***+*** | ***+*** | ***+***       |         | *+*         |
-|   [ggseg](https://github.com/ggseg/python-ggseg)                                                        |         |         | ***+***       |         |             |
-|   [graphpype](https://neuropycon.github.io/graphpype/)                                                  |         |         |               | ***+*** |             |
-|   [MMVT](https://github.com/pelednoam/mmvt)                                                             |         | ***+*** | ***+***       | ***+*** |             |
-|   [MNE](https://github.com/mne-tools/mne-python)                                                        | ***+*** | ***+*** | ***+***       |         |             |
-|   [mrivis](https://github.com/raamana/mrivis)                                                           | ***+*** |         |               |         |             |
-|   [NaNSlice](https://github.com/spinicist/nanslice)                                                     | ***+*** |         |               |         |             |
-|   [netneurotools](https://github.com/netneurolab/netneurotools)                                         |         | ***+*** | ***+*****\*** |         |             |
-|   [netplotbrain](https://github.com/wiheto/netplotbrain)                                                |         |         | ***+***       | ***+*** |             |
-|   [nilearn](https://github.com/nilearn/nilearn)                                                         | ***+*** | ***+*** | ***+***       | ***+*** |             |
-|   [niwidget](https://github.com/nipy/niwidgets)                                                         | ***+*** | ***+*** |               |         | ***+***     |
-|   [Pycortex](https://github.com/gallantlab/pycortex)                                                    | ***+*** | ***+*** | ***+*****\*** |         |             |
-|   [pySurfer](https://pysurfer.github.io/)                                                               |         | ***+*** | ***+*****\*** |         |             |
-|   [surfice](https://github.com/neurolabusc/surf-ice)                                                    | ***+*** | ***+*** | ***+***       | ***+*** | ***+***     |
-|   [Visbrain](https://github.com/EtienneCmb/visbrain)                                                    | ***+*** | ***+*** | ***+***       | ***+*** |             |
-| **MATLAB**                                                                                              |         |         |               |         |             |
-|   [BrainNetViewer](https://www.nitrc.org/projects/bnv/)                                                 | ***+*** |         | ***+***       | ***+*** |             |
-|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)                                                  |         | ***+*** | ***+*****\*** |         |             |
-|   [Brainstorm](https://github.com/brainstorm-tools/brainstorm3)                                         | ***+*** | ***+*** | ***+***       |         |             |
-|   [bspmview](https://www.bobspunt.com/software/bspmview/)                                               | ***+*** |         | ***+***       |         |             |
-|   [CandlabCore](https://github.com/canlab/CanlabCore)                                                   | ***+*** |         | ***+***       |         |             |
-|   [ECoG/fMRI Vis toolbox](https://github.com/Immiora/matlab-ecog-visualization)                         |         | ***+*** | ***+*****\*** |         |             |
-|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)                                                  |         |         | ***+***       |         |             |
-|   [FieldTrip](https://www.fieldtriptoolbox.org/)                                                        | ***+*** | ***+*** |               |         |             |
-|   [Lead-DBS](https://www.lead-dbs.org/)                                                                 | ***+*** |         | ***+***       |         |             |
-|   [mni2fs](https://github.com/dprice80/mni2fs)                                                          |         | ***+*** |               |         |             |
-|   [mrtools](https://github.com/justingardner/mrTools)                                                   | ***+*** | ***+*** | ***+*****\*** |         |             |
-|   [plotSurfaceROIBoundary](https://github.com/StuartJO/plotSurfaceROIBoundary)                          |         | ***+*** | ***+*****\*** |         |             |
-|   [Vistasoft](https://github.com/vistalab/vistasoft)                                                    | ***+*** | ***+*** | ***+***       |         | ***+***     |
+In the following sections we outline the primary steps required when generating programmatic and reproducible human brain visualizations, and provide tools and heuristics to guide this process.
 
-Note: The tools listed contain functionality required to generate (at least close-to) publication-ready neuroimaging figures via user-entered code within R, MATLAB and Python environments.
-This list does not include cross-platform general purpose visualization software.\
-***\**** Cortex only
+## Selecting your programming language
 
-# Examples of Neuroimaging Visualization Packages available in *R and Python*
+The first step (Figure 1) in generating code-based visualization is selecting the coding language you want to use.
+Three of the most popular languages are R, Python and MATLAB, all of which have many options for generating brain visualizations (see Table 1).
+This decision can be made based on what language you have prior experience with, or if you used one of these languages for the analysis part of your project, there can be advantages in using the same language for visualizations.
+Switching from your analysis environment to separate visualization environment, or a GUI-based visualization software, can be a cumbersome deviation from the scientific workflow.
+This can make debugging errors more difficult, as you must regularly switch program to visually examine the results of any modifications or adjustments to prior analyses.
+Using the brain visualization tools that already exist within your chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analysis.
+Alternatively, the choice of programming language may be dictated by visualization and data-type you require.
+For instance, visualizing streamlines from tractography may not be currently available in the R environment (Table 1), and therefore requires the use of Python or MATLAB.
+Other constraints may include access to proprietary software like MATLAB, which would necessitate the use of open-source options such as R or Python.
 
-The following two figures provide examples of voxel, vertex, ROI and edge-level figures generated within R (Figure 1) and Python (Figure 2) using open source, well documented and beginner-friendly packages.
-These are not an exhaustive representation of packages available for visualizing brain data in R and Python (See Table 1), rather, the figures aim to give the reader a sense of the available options, and an entry point to using them.
-All code used to compile the figures, as well as the contents of each panel are provided an accompanying [online repository.](https://github.com/sidchop/RepoNeuroVis)
+\newpage
+
+```{r Figure1, echo=FALSE, out.height = '70%', out.width = '100%', fig.align='center'}
+  knitr::include_graphics('images/Figure1.png')
+```
+
+**Figure 1** Primary steps to generating programmatic and reproducible human brain visualizations.
+Each step is outlined in Each step is outlined in the corresponding sub-section of Section \@ref(generating-code-based-visualizations).
+\newpage
+
+## Identify your visualization type
+
+Neuroimaging data and its derivatives can be visualized in multiple forms, which have different associated file-types and visualization requirements (see Section \@ref(identify-your-input-file-format.)).
+A brief description of more popular visualization types is provided below:
+
+***Voxel*****.** A voxel is the 3D equivalent of a pixel.
+In neuroimaging, voxels are used to represent the intensity values of a 3D scan, such as an MRI or CT scan.
+The voxels can be rendered in different colors to indicate different tissue types or other features of interest.
+For example, in functional MRI (fMRI) scans, voxels can be colored based on their level of activation, to show which areas of the brain are more active during a specific task or for MRI, parts of the brain showing statistically significant anatomical alterations.
+These visualizations are often displayed as either slices in axial, sagittal or coronal planes (Fig2A-B; Fig3A-B; or a 3D rendering of the whole brain.
+Statistical values are displayed as overlays on template anatomical images, which follow a common stereotaxic coordinate system (e.g., MNI152), or on individual-specific anatomical images ('native-space').
+
+***Vertex.*** A vertex is a point in 3D space that represents a location on the surface.
+In neuroimaging, vertices are used to create a mesh representation of a brain structure, such as the cerebral cortex or a subcortical structure.
+Each vertex has a set of coordinates that specify its location in 3D space and is connected to other vertices to form triangles, which make up the mesh.
+Vertices can be used to create a 3D visualization of the brain surface and color-coded based on different attributes such as sulcal depth, thickness, or functional activation.
+These visualizations are often displayed as 3D rendering of each hemisphere from medial and lateral views (Fig2C,G-H; Fig3C-D).
+Statistical values are displayed as overlays on template surfaces which follow a common stereotaxic coordinate system and fixed number of vertices (e.g., average; Fig2C,G; Fig3C-D), or on individual-specific surfaces that have been reconstructed using an anatomical image (Fig2H).
+
+***Regions of Interest (ROI).*** In neuroimaging, ROIs are used to identify specific brain structures or areas that are relevant to the research question.
+ROIs can be defined using various methods, such as manual tracing, atlas-based parcellation, or functional activation patterns.
+Visualizing ROIs can be done by grouping and assigning the same statistical value or color to sets of voxels or vertices (Fig2B-C; Fig3B-C), or can be done through polygons.
+Polygonal brain visualizations are simple 2D or 3D shapes which graphically represent the brain or specific structures, but do not carry any additional information on spatial coordinates and only roughly estimate the shape of the brain and its structures.
+Each region of 2D (Fig 2D,F) and 3D (Fig 2E; Fig3E) polygon visualizations can be filled in with colors representing a region label or a statistical value.
+
+***Edge.*** In the context of neuroimaging, an edge often represents a connection between two brain regions, which can be thought of as a line connecting two nodes (brain regions) in a network.
+To visualize edges, one common method is to use a matrix representation, often called a 'connectivity' or 'adjacency' matrix, which shows the strength of the connections between all pairs of brain regions.
+These matrices can be displayed as a heat map, where different colors represent different levels of connectivity.
+Another way to represent edges is using a 2D graph, where the vertices represent brain regions and the edges represent the connections between those regions, which displayed in a variety of ways, such as a 2D or 3D representation of the brain (Fig2I; Fig3F), with edges represented as lines connecting the vertices.
+The edges and nodes can also be color-coded based on properties such as the strength of the connection.
+
+***Streamlines.*** Streamlines are a representation of the white matter fibers in the brain and are usually generated using tractography applied to diffusion MRI (dMRI).
+Streamlines are typically visualized as 3D lines that connect different points of the brain and can be used to create 3D visualizations of specific white matter tracts, or all streamlines between regions.
+They can be color-coded based on the direction of the fibers or properties of the tract, such as myelination.
+These visualizations can be overlaid on voxel or vertex level anatomical brain representations to provide an anatomical reference point.
+
+The following two figures provide examples of voxel, vertex, ROI and edge-level visualizations generated within R (Figure 2) and Python (Figure 3) using open source, well documented and beginner-friendly packages.
+These are not an exhaustive representation of packages available for visualizing brain data in R and Python (see Table 1).
+Rather, the figures aim to give the reader a sense of the many options available, and an entry point to choosing the type of brain visualization needed (also see Section 4).
+All code used to compile the figures, as well as the contents of each panel are provided in the accompanying [online repository](https://github.com/sidchop/RepoNeuroVis).
 
@@
 ```
-```{r Figure1, echo=FALSE, out.height = '60%', out.width = '70%', fig.align='center'}
+```{r Figure2, echo=FALSE, out.height = '60%', out.width = '70%', fig.align='center'}
 
@@
 
-***Figure 1.
+***Figure 2.
 Examples of brain imaging visualizations made using R.*** A) Voxel-level statistical map thresholded and overlaid over a T1-weighted template image, with a single axial slice shown.
@@
 ```
-```{r Figure2, echo=FALSE, out.height = '60%', out.width = '80%', fig.align='center', dev='cairo_pdf'}
+```{r Figure3, echo=FALSE, out.height = '70%', out.width = '90%', fig.align='center', dev='cairo_pdf'}
 knitr::include_graphics("./data/python/Figure2.pdf")
@@
 
-***Figure 2.
+***Figure 3.
 Examples of brain imaging visualizations made using Python.*** A) Voxel-based statistical map thresholded and overlaid over a T1-weighted template data, with a single axial slice shown.
@@
 Made using `plotting.plot_roi` function from `nilearn`.
-C) Medial and lateral views of an cortex-wide ROI atlas, displayed on an inflated template surface where a statistical value has been randomly assigned to each ROI.
+C) Medial and lateral views of a cortex-wide ROI atlas, displayed on an inflated template surface where a statistical value has been randomly assigned to each ROI.
 Made using `Plot` function from `surfplot`.
@@
 
+## Identify your input file-format
+
+Human neuroimaging data and its derivatives come in many different file formats, and code-based visualization packages may have specific requirements about which formats they are designed to work with.
+Table 1 provides a key, listing which data-type can be used as inputs for each package.
+Some of the most common MRI file-formats used for visualizations are briefly described below:
+
+***Plain text format.*** The simplest input formats are a scalar, vector, and matrix, which represent a single data-point, one-dimensional array of data (e.g., a single column or row), and a two-dimensional array of data (e.g.
+multiple columns or rows), respectively.
+These data are often stored in plain text formats such as `.txt`, `.csv`, and `.tsv` and are often generated by neuroimaging analyses software such as SPM, FSL or FreeSurfer.
+These files can also contain rows and columns of region names and/or spatial coordinates.
+All programming languages have functions to read these plain text formats into your environment as data structures such as matrices and arrays.
+If you have conducted your imaging analyses using your chosen programming language, rather than stand-alone software, the data you want to visualize will likely already be available to you within your coding environment.
+These formats are often used in region-level visualizations, where groups of voxels or vertices share the same value or color (Fig2B-F; Fig3B-E), or in edge-level visualizations, where a matrix is used to identify which two brain regions are connected by an edge (Fig2F; Fig3F).
+
+***NIfTI*****.** NIfTI (`.nii`) files store 3D or 4D image data, which is often a matrix of voxel intensities or a series of 3D matrices for 4D data such as fMRI and dMRI.
+The image data is stored as a 3D matrix of voxel intensities, and the header contains information such as the image dimensions, voxel size, and data type.
+Additional information such as patient demographics and scanner parameters can also be stored in the header in the form of metadata.
+
+***GIfTI.*** GIFTI (`.gii`) is an extension of the NifTI format, which stores data in a surface-based format, with the data represented as a set of vertices, edges, and faces that define a surface mesh.
+The format also includes support for storing data such as curvature, thickness, and functional activity maps on the surface mesh.
+Often `.gii` files will have a pre-indicator of what information the file contains, such as `.surf.gii`, which would contain only vertices, edges and faces to define a surface mesh.
+Or `.func.gii` files which contain data values for every vertex, which are essentially data arrays whose indices correspond to a surface file and need a matching surface file to know where in the brain to assign the data values.
+GIFTI files can also store multiple surfaces in a single file, and can include information about the topology of the surfaces, such as the number of vertices, edges, and faces.
+Additional metadata can also be stored in the header.
+
+***CIfTI.*** CIFTI (`.cii`) files can store data from both surface-based and volume-based neuroimaging analyses.
+For surface-based data, the file contains vertex coordinates, and data values at each vertex.
+For volume-based data, the file contains a 3D matrix of voxel intensities.
+Often this datatype is used to represent the cortex as vertices, and subcortical structures as voxels.
+CIFTI files can be divided into three main types: `.dtseries` (store time-series data, such as fMRI data), `.dtscalar` (store scalar data, such as thickness or curvature maps), and `.dtlabel` (store label data, such as parcellations of the brain).
+CIFTI files can also include fields with additional information such as surface and volume registration information.
+
+***FreeSurfer.*** FreeSurfer is a commonly used image processing and analysis software that comes with a variety of proprietary formats to store outputs.
+The primary format is `.mgh` and its compressed version `.mgz`, which like `.nii` files, store voxel-level data.
+Vertex-level data is stored in multiple formats such as `.pial`, `.white`, and `.inflated`.
+FreeSurfer also uses `.label` files to store a list of vertices and associated labels for each brain structure, and `.annot` files FreeSurfer store annotation information such as vertices, labels and color information that can be overlaid on the surface reconstruction.
+
+***Tractography.*** Commonly used tractography file formats are `.trk` and `.tck` developed by the TrackVis and MRtrix software packages, respectively.
+Both formats store coordinates for streamlines as a series of 3D points, with each point represented by its x, y, and z coordinates, as well as a tract header with the number of points in the tract, the properties of the tract (e.g., mean diffusivity, fractional anisotropy), as well as the starting and ending indices of the tract.
+
+Critically, many of these datatypes are not immutable, and can be converted between each other.
+There are many ways to convert between format, for instance, the Connectome Workbench (Marcus et al., 2011), and corresponding R-packages like `ciftitools` (Pham, Muschelli, & Mejia, 2022) allow for conversion between NifTI, CifTI and GifTI formats.
+FreeSurfer functions such as `mri_convert`, NiBabel (Brett et al., 2023) library in Python and fsbrain (Schäfer & Ecker, 2020) in R, all allow for conversion between the propriety formats and open-source NifTI, CifTI and GifTI formats.
+Therefore, if the file format you have access to is not compatible with the visualization package, library, or toolbox you want to use, you may be able to convert your data into the desired format.
+Although, it is important to visualize and validate data after converting data-structures, such as converting between vectors and matrices, to ensure the data have not been misinterpreted, and also to be aware that unintended consequences of mapping between 2D surface and 3D volume formats may arise (e.g., Ciantar et al., 2022).
+
+## Select your package, library or toolbox
+
+The previous steps of deciding your programming environment, visualization type and input data-type will help you decide which package, library or toolbox is the right choice for you.
+Table 1 provides a list of tools in R, Python and MATLAB, characterized by whether they are able to generate voxel, vertex, ROI, edge and streamline based visualizations.
+While each tool may contain the ability to generate code-based visualizations, some tools are more beginner-friendly and better documented than others.
+Usually, this is the case for tools that are specifically designed for brain data visualization, as opposed to tools which are for brain data analysis, but also provide some limited visualization functionality.
+Some examples of well documented and beginner-friendly tools have been provided via a code template generator (see Section \@ref(brain-code-a-web-app-for-generating-code-templates-for-brain-visualizations)).
+
+An important consideration when selecting a tool can be whether it can generate publication-ready plots.
+Publication-ready plots are high resolution, labelled, contain all color bars and legends, and require no additional manual image manipulation.
+While all tools listed in Table 1 contain some of these features, some tools enable more precise control over publication-ready features such as legend placement, color bar placement, annotations, labeling, and multi-panel figures.
+These tools usually produce visualization that rely on mainstream general purpose plotting engines such as `ggplot` in R and `matplotlib` in Python.
+This allows users to leverage many additional features to make their brain visualizations publication ready.
+For example, the `ggseg` and `ggsed3d` packages (Mowinckel & Vidal-Piñeiro, 2020) in R generate plots compatible with the widely-used grammar of graphics (i.e., `ggplot2`) and `plotly` engines, respectively.
+Similarly, the `Nilearn` library in python allows plots to be generated using `matplotlib` or `plotly` engines.
+We note that many of the others listed in Table 1 also rely on common plotting engines to generate visualizations.\
+
+\newpage
+
+**Table1 Examples of code-based neuroimaging visualizations tools that can be accessed directly within R, MATLAB and Python environments.**
+
+|                                                                                                                 | Voxel                | Vertex              | ROI                   | Edge             | Streamlines    |
+|:-----------|:-----------|:-----------|:-----------|:-----------|:-----------|
+| **R**                                                                                                           |                      |                     |                       |                  |                |
+|   [ANTsR](https://antsx.github.io/ANTsR/articles/ANTsR.html)                                                    | ***+^2^***           | ***+^2^***          | ***+^2^***            | ***+^1^***       |                |
+|   [brainconn](https://github.com/sidchop/brainconn)                                                             |                      |                     |                       | ***+^1^***       |                |
+|   [brainR](https://github.com/muschellij2/brainR)                                                               | ***+^2,6,7^***       |                     | ***+^6,7^***          |                  |                |
+|   [ciftitools](https://github.com/mandymejia/ciftiTools)^**\^**^                                                | ***+^2,3^***         | ***+^2,3^***        | ***+\*^,2,3^***       |                  |                |
+|   [fsbrain](https://github.com/dfsp-spirit/fsbrain)^**\^**^                                                     | ***+^4^***           | ***+^3,4^***        | ***+\*^,6,3,4^***     |                  |                |
+|   [ggseg](https://github.com/ggseg/ggseg)^**\^**^                                                               |                      |                     | ***+^1,4^***          |                  |                |
+|   [neurobase](https://github.com/muschellij2/neurobase)                                                         | ***+^2^***           |                     |                       |                  |                |
+|   [oro.nifti](https://github.com/muschellij2/oro.nifti)                                                         | ***+^2^***           |                     |                       |                  |                |
+| **Python**                                                                                                      |                      |                     |                       |                  |                |
+|   [ANTsPy](https://github.com/ANTsX/ANTsPy)                                                                     | ***+^2^***           | ***+^2^***          | ***+^2^***            |                  |                |
+|   [brainiak](https://github.com/brainiak/brainiak)                                                              | ***+^2^***           |                     |                       |                  |                |
+|   [Brainplotlib](https://github.com/feilong/brainplotlib)                                                       |                      | ***+^1,11^***       | ***+\*^,1,11^***      |                  |                |
+|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)/[surfplot](https://github.com/danjgale/surfplot)^**\^**^ |                      | ***+^3,4,6,7.8^***  | ***+\*^,3,4,6,7,8^*** |                  |                |
+|   [DIPY](https://github.com/dipy/dipy)^**\^**^                                                                  | ***+^2^***           |                     |                       |                  | ***+^5,9^***   |
+|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)^**\^**^                                                  |                      |                     | ***+^1,3,4,6,8^***    |                  |                |
+|   [FSLeyes](https://github.com/pauldmccarthy/fsleyes)                                                           | ***+^2,3,4^***       | ***+^2,3,4^***      | ***+^2,3,4^***        |                  | ***+^2,5^***   |
+|   [ggseg](https://github.com/ggseg/python-ggseg)                                                                |                      |                     | ***+^1,4^***          |                  |                |
+|   [graphpype](https://neuropycon.github.io/graphpype/)                                                          |                      |                     |                       | ***+^1,10,11^*** |                |
+|   [MMVT](https://github.com/pelednoam/mmvt)                                                                     |                      | ***+2,4***          | ***+2,4***            | ***+^1,10,11^*** |                |
+|   [MNE](https://github.com/mne-tools/mne-python)                                                                | ***+^2,4^***         | ***+^4^***          | ***+^4^***            |                  |                |
+|   [mrivis](https://github.com/raamana/mrivis)                                                                   | ***+^2^***           |                     |                       |                  |                |
+|   [NaNSlice](https://github.com/spinicist/nanslice)                                                             | ***+^2^***           |                     |                       |                  |                |
+|   [netneurotools](https://github.com/netneurolab/netneurotools)                                                 |                      | ***+^4^***          | ***+\*^,3^***         |                  |                |
+|   [netplotbrain](https://github.com/wiheto/netplotbrain)^**\^**^                                                |                      |                     | ***+^1,2^***          | ***+^1,11^***    |                |
+|   [nilearn](https://github.com/nilearn/nilearn)^**\^**^                                                         | ***+^2^***           | ***+^2,3^***        | ***+^2,3^***          | ***+^1,11^***    |                |
+|   [niwidget](https://github.com/nipy/niwidgets)                                                                 | ***+^2,3,4^***       | ***+^2,3,4^***      |                       |                  | ***+^5^***     |
+|   [Pycortex](https://github.com/gallantlab/pycortex)                                                            | ***+^8,11,12^***     | ***+^4,8,11,12^***  | ***+\*^,4,8,11,12^*** |                  |                |
+|   [pySurfer](https://pysurfer.github.io/)^**\^**^                                                               |                      | ***+^4^***          | ***+\*^,4^***         |                  |                |
+|   [surfice](https://github.com/neurolabusc/surf-ice)                                                            | ***+^2,3,4^***       | ***+^2,3,4^***      | ***+^2,3,4,6^***      | ***+^1,4,6^***   | ***+^4,5,6^*** |
+|   [Visbrain](https://github.com/EtienneCmb/visbrain)                                                            | ***+^1,2,3,6^***     | ***+^1,2,3,6^***    | ***+^1,2,3,6^***      | ***+^1,2,3,6^*** |                |
+| **MATLAB**                                                                                                      |                      |                     |                       |                  |                |
+|   [BrainNetViewer](https://www.nitrc.org/projects/bnv/)                                                         | ***+^2,3,4,6,13^***  |                     | ***+^2,3,4,6,13^***   | ***+^1,10^***    |                |
+|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)^**\^**^                                                  |                      | ***+^3,4,6,7,8^***  | ***+\*^,3,4,6,7,8^*** |                  |                |
+|   [Brainstorm](https://github.com/brainstorm-tools/brainstorm3)                                                 | ***+^2^***           | ***+^4,6^***        | ***+^3,4,6^***        |                  |                |
+|   [bspmview](https://www.bobspunt.com/software/bspmview/)                                                       | ***+^2,13^***        |                     | ***+^2,13^***         |                  |                |
+|   [CandlabCore](https://github.com/canlab/CanlabCore)                                                           | ***+^2,10,13^***     |                     | ***+^2,10,13^***      |                  |                |
+|   [ECoG/fMRI Vis toolbox](https://github.com/Immiora/matlab-ecog-visualization)                                 |                      | ***+^10^***         | ***+\*^,10^***        |                  |                |
+|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)^**\^**^                                                  |                      |                     | ***+^1,10^***         |                  |                |
+|   [FieldTrip](https://www.fieldtriptoolbox.org/)                                                                | ***+^2^***           | ***+^10^***         |                       |                  |                |
+|   [Lead-DBS](https://www.lead-dbs.org/)                                                                         | ***+^2^***           |                     | ***+^2,3,4^***        |                  |                |
+|   [mni2fs](https://github.com/dprice80/mni2fs)                                                                  |                      | ***+^2,3^***        |                       |                  |                |
+|   [mrtools](https://github.com/justingardner/mrTools)                                                           | ***+^2^***           | ***+^4,12^***       | ***+\*^,2,4,12^***    |                  |                |
+|   [plotSurfaceROIBoundary](https://github.com/StuartJO/plotSurfaceROIBoundary)                                  |                      | ***+^4,10^***       | ***+\*^,4,10^***      |                  |                |
+|   [Vistasoft](http://web.stanford.edu/group/vista/cgi-bin/wiki/index.php/Visualization)                         | ***+^2^***           | ***+^2^***          | ***+^2^***            |                  | ***+^10^***    |
+
+Note: The tools listed contain functionality required to generate (at least close-to) publication-ready neuroimaging figures via user-entered code within R, MATLAB and Python environments.
+This list does not include cross-platform general purpose visualization software.
+**1** = .txt/.csv (scalar, vector, matrix as input); **2** = .nii/.nii.gz (nifti as input); **3** = .cii/.gii (cifti or gifti files as input, includes any subtypes e.g.
+dlabel, dtseries, .surf); **4** = FreeSurfer formats as input, including .mgz, .annot, .label, .curv, .wm etc); **5** = .trk/.tck (tractograms as input); **6** = .obj (3D object format); **7** = .ply (3D polygon format); **8** = .vtk (Visualization Toolkit format); **9** = .fib (Legacy vtk format); **10** = .mat (MATLAB format); 11 = .npy/.npz (Python numpy format); **12** = .off (object file format); ***\**** = Cortex only; **\^** = Well-documented, beginner friendly code-based visualization tool (note that this is a subjective criteria intended to guide new-users, and many other tools in this table may also meet this criteria).
+
+***\**** Cortex only \newpage
+
+## Generate and Share Visualization
+
+Once the plot has been generated, the visualization can be shared in multiple different ways, with differing levels of replicability (see Section \@ref(replicability)), interactivity (see Section \@ref(integrative-and-interactive-reporting)) and visual quality.
+One common way to share your visualization is to export it as a image raster format such as `.png`, `.tiff` or `.jpeg`.
+These commonly used file formats represent images as a grid of pixels, where each pixel in the grid contains information about the color and intensity of that specific point in the image.
+These formats are resolution dependent, which means that if you try to scale them larger, they will become pixelated and difficult to parse.
+Vector formats, such a `.svg` and `.eps` on the other hand, represent images using mathematical equations, which means that they can be scaled larger or smaller without losing quality.
+While all coding environments provide ways to export visualizations into raster formats, exporting using vector formats, while visually superior, can be dependent on the specific tool you have selected.
+These image files can the be inserted into outputs such as a manuscript and the associated code can be included in an online repository such as [GitHub](https://docs.github.com/en/get-started/quickstart/hello-world) or in supplementary materials associated with the output.
+While this is the simplest way to share visualizations and code, it is only reproducible if the input data, such as NIfTI files used to generate the figure, are also shared.
+In addition to repositories like GitHub, there now exist many platforms to share neuroimaging specific data, such as [OpenNeuro](https://openneuro.org/) and [NeuroVault](https://neurovault.org/).
+If the data used as input for the visualization cannot be shared, synthetic data can be provided (Quintana, 2020) to ensure the reproducibility of the brain visualizations.
+Alternatively, software such as *R Markdown*, *Quarto* and *Jupyter Notebook* allow for the mixing of prose and code in a single script, resulting in fully reproducible and publication ready papers (see Section \@ref(integrative-and-interactive-reporting)).
+While guidance on proper organization of the neuroimaging visualization code is beyond the scope of this guide, we point readers to other practical guides on this topic (Van Vliet, 2020; Gorgolewski & Poldrack, 2016)
+
+# Brain-Code: A web-app for generating code templates for brain visualizations
+
+To assist new users transition into generating code-based brain visualizations we have developed a web-app (<https://sidchop.shinyapps.io/braincode/>) which allows for interactive generation of code-templates for beginner friendly libraries/packages in R and Python.
+In the web-app, users can select R or Python as their coding environment, and choose between voxel, ROI, vertex, and edge-level visualizations.
+They can then manually adjust a limited set of visualization setting, such as color-scales and view, and are provided with a reactive code-template which can be copied and then used within their respective programming environment.
+The provided code templates require users to customize the code, such as alter file-paths.
+The available settings have been purposefully limited to allow users to explore and fine-tune additional visualization options within their own programming environment.
+The code-template also contains links and prompts to more detailed documentation, alternate packages/libraries and tutorials which allow for more complex and publication-ready brain visualizations.
+
+\newpage
+
+```{r Figure4, echo=FALSE, out.height = '70%', out.width = '100%', fig.align='center'}
+  knitr::include_graphics('images/image-17223477.png')
+```
+
+**Figure 4.** Interface for [BrainCode](https://sidchop.shinyapps.io/braincode/) web-app which generates simple code-based templates for brain visualizations.
+(1) Select your programming environment (R or Python).
+(2) Select your visualization type (Voxel, Region, Vertex and Edge).
+There is also a Frequent Asked Questions tab to aid users.
+(3) Manually adjust limited visualization settings and examine how it reactively changes the visualization (4) and code template (5).
+(5) Copy the code template into your programming environment, change file-paths to your own data and explore other visualization settings offered by the functions.
+\newpage
+
 # Limitations and Functionality Gaps
 
-While many code-based tools are well documented and do not require a strong knowledge of programming, there can still be a steeper learning curve for new users, compared to using a GUI.
-This is especially true for the purpose of a publication-ready figure, were fine adjustments to visual auxiliary such as legend placement, font size and multi-panel figure positioning are needed.
-While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability.
-Relatedly, while some interactive image viewers can be opened within an integrated development environment like R-Studio (e.g.
-Muschelli, 2016), for quick and interactive viewing of single images, GUI tool can be faster and more practical.
+While many code-based tools are well documented and do not require a strong knowledge of programming, there can still be a steep learning curve for new users, compared to using a GUI.
+We hope that the provided template generator gives jumping-off point for new users.
+Nonetheless, this is especially true for the purpose of a publication-ready figure, where fine adjustments to visual features such as legend placement, font size and multi-panel figure positioning may be needed.
+While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability, with tools that use established graphic engines such as `ggplot2`, `matplotlib` and `plotly` providing the most versatile and well-documented features for visual auxiliary.
+Relatedly, while some interactive image viewers can be opened within an integrated development environment like R-Studio (e.g., Muschelli, 2016), for quick and interactive viewing of single images, in this context, GUI tool can be faster and more practical.
 
-Often cerebellar and brain-stem regions are not well represented in software (e.g.
-Figure 1), potentially mirroring the cortico-centric sentiment that has prevailed in human neuroimaging research (Chin, Chang, & Holmes, 2022).
+Often cerebellar and brain-stem regions are not well represented in software (e.g., Figure 2-3), potentially mirroring the cortico-centric sentiment that has prevailed in human neuroimaging research (Chin, Chang, & Holmes, 2022).
 Likewise, custom non-cortical atlases such as non-standard subcortical atlas schemes are not yet straightforward, and usually require multiple functions and packages to visualize.
-Finally, some neuroimaging derived data types, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools.
+Visualizing these structures often requires chaining together multiple GUI-based tools (see Madan, 2015).
+Alternatively users can convert neuroimaging spesific file-formats into domain general visualization or polygon formats, such as `.vtk`, `.ply` or `.obj`, which can be read, manipulated and visualized using general purpose code-based tools.
+Examples of such tools include PyVista and Mayavi in Python and rayshader and plotly in R.
+Moreover, some neuroimaging derived datatypes, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools and future development should focus on enhancing visualization capabilities using these data-types.
 
-As can be seen in Table 1, there are usually multiple packages within each programming environment which are capable of visualizing each data type.
-While this provide choice for advanced users, it can also lead to confusion for novice users who may not be familiar with the nuanced differences between tools.
-Future work should continue to consolidate brain visualization methods into unified beginner-friendly code-based tools which are capable of plotting multiple data types.
+As can be seen in Table 1, there are usually multiple packages within each programming environment which can visualize each data type.
+While this provides choice for advanced users, it can also lead to confusion for novice users who may not be familiar with the nuanced differences between tools.
+While the process we have outlined, and the table and web-app we have provided will help users decide the ideal package, future work should continue to consolidate brain visualization methods into unified beginner-friendly code-based tools which rely on established and well-documented graphic engines and can plot multiple data types.
 
@@
 
-Poldrack, R.
-A., Baker, C.
-I., Durnez, J., Gorgolewski, K.
-J., Matthews, P.
-M., Munafò, M.
-R., ...
-& Yarkoni, T.
-(2017).
-Scanning the horizon: towards transparent and reproducible neuroimaging research.
-Nature reviews neuroscience, 18(2), 115-126.
+Baker M. (2015). Social media: A network boost. Nature 518: 263--265. 10.1038/nj7538-263a
 
-Steel, G.
-(2013, September 3).
-Publishing research without data is simply advertising, not science.
-Open Knowledge Foundation.
-<https://blog.okfn.org/2013/09/03/publishing-research-without-data-is-simply-advertising-not-science/>
+Brett, Matthew, Markiewicz, Christopher J., Hanke, Michael, Côté, Marc-Alexandre, Cipollini, Ben, McCarthy, Paul, Jarecka, Dorota, Cheng, Christopher P., Halchenko, Yaroslav O., Cottaar, Michiel, Larson, Eric, Ghosh, Satrajit, Wassermann, Demian, Gerhard, Stephan, Lee, Gregory R., Wang, Hao-Ting, Kastman, Erik, Kaczmarzyk, Jakub, Guidotti, Roberto, ...
+freec84.
+(2023).
+nipy/nibabel: 5.0.0 (5.0.0).
+Zenodo.
+<https://doi.org/10.5281/zenodo.7516526>
 
-The Comprehensive R Archive Network. (n.d.). Retrieved April 21, 2022, from <https://cran.r-project.org/>
+Chin, R., Chang, S. W., & Holmes, A. J. (2022). Beyond cortex: The evolution of the human brain. *Psychological Review*.
 
-Muschelli, J., Gherman, A., Fortin, J.
-P., Avants, B., Whitcher, B., Clayden, J.
-D., ...
-& Crainiceanu, C.
-M.
-(2019).
-Neuroconductor: an R platform for medical imaging analysis.
-*Biostatistics*, *20*(2), 218-239.
-
-Pham, D., Muschelli, J., & Mejia, A. (2022). ciftiTools: A package for reading, writing, visualizing, and manipulating CIFTI files in R. *NeuroImage*, 118877.
-
-Mowinckel, A. M., & Vidal-Piñeiro, D. (2020). Visualization of brain statistics with R packages ggseg and ggseg3d. *Advances in Methods and Practices in Psychological Science*, *3*(4), 466-483.
-
-Huntenburg, J., Abraham, A., Loula, J., Liem, F., Dadi, K., & Varoquaux, G. (2017). Loading and plotting of cortical surface representations in Nilearn. *Research Ideas and Outcomes*, *3*, e12342.
-
-Wickham H. (2016). *ggplot2: Elegant Graphics for Data Analysis.* Springer-Verlag New York. ISBN 978-3-319-24277-4, <https://ggplot2.tidyverse.org>.
-
-Muschelli, J. (2016). *papayar*. GitHub repository. <https://github.com/muschellij2/papayar>
+Ciantar, K.
+G., Farrugia, C., Galdi, P., Scerri, K., Xu, T., & Bajada, C.
+J.
+(2022).
+Geometric effects of volume-to-surface mapping of fMRI data.
+*Brain Structure and Function*, *227*(7), 2457-2464.
+Code Ocean.
+(2021, October 27).
+Computational Research Platform on.
+Retrieved April 21, 2022, from <https://codeocean.com/>
 
@@
 NeuroImage 80(2013):62-79.
-
-Sudlow, C., Gallacher, J., Allen, N., Beral, V., Burton, P., Danesh, J., ...
-& Collins, R.
-(2015).
-UK biobank: an open access resource for identifying the causes of a wide range of complex diseases of middle and old age.
-PLoS medicine, 12(3), e1001779.
 
@@
 
-Pernet, C., & Madan, C. R. (2019). Data visualization for inference in tomographic brain imaging.
+Gorgolewski, K. J., & Poldrack, R. A. (2016). A practical guide for improving transparency and reproducibility in neuroimaging research. PLoS biology, 14(7), e1002506.
 
-Code Ocean. (2021, October 27). Computational Research Platform on. Retrieved April 21, 2022, from <https://codeocean.com/>
+Huber, B., Barnidge, M., Gil de Zúñiga, H., & Liu, J. (2019). Fostering public trust in science: The role of social media. *Public understanding of science*, *28*(7), 759-777.
 
-Madan, C. R. (2015). Creating 3D visualizations of MRI data: A brief guide. F1000Research, 4.
-
-Baker M. (2015). Social media: A network boost. Nature 518: 263--265. 10.1038/nj7538-263a
+Huntenburg, J., Abraham, A., Loula, J., Liem, F., Dadi, K., & Varoquaux, G. (2017). Loading and plotting of cortical surface representations in Nilearn. *Research Ideas and Outcomes*, *3*, e12342.
 
@@
 
+Luc, J.
+G., Archer, M.
+A., Arora, R.
+C., Bender, E.
+M., Blitz, A., Cooke, D.
+T., ...
+& Antonoff, M.
+B.
+(2021).
+Does tweeting improve citations?
+One-year results from the TSSMN prospective randomized trial.
+The Annals of thoracic surgery, 111(1), 296-300.
+
+Madan, C. R. (2015). Creating 3D visualizations of MRI data: A brief guide. F1000Research, 4.
+
+Marcus, D.
+S., Harwell, J., Olsen, T., Hodge, M., Glasser, M.
+F., Prior, F., \...
+& Van Essen, D.
+C.
+(2011).
+Informatics and data mining tools and strategies for the human connectome project.
+*Frontiers in neuroinformatics*, *5*, 4.
+
+Mowinckel, A. M., & Vidal-Piñeiro, D. (2020). Visualization of brain statistics with R packages ggseg and ggseg3d. *Advances in Methods and Practices in Psychological Science*, *3*(4), 466-483.
+
+Muschelli, J. (2016). *papayar*. GitHub repository. <https://github.com/muschellij2/papayar>
+
+Muschelli, J., Gherman, A., Fortin, J.
+P., Avants, B., Whitcher, B., Clayden, J.
+D., ...
+& Crainiceanu, C.
+M.
+(2019).
+Neuroconductor: an R platform for medical imaging analysis.
+*Biostatistics*, *20*(2), 218-239.
+
+Mueller-Herbst, J.
+M., Xenos, M.
+A., Scheufele, D.
+A., & Brossard, D.
+(2020).
+Saw it on Facebook: The role of social media in facilitating science issue awareness.
+*Social Media+ Society*, *6*(2), 2056305120930412.
+
+Pernet, C., & Madan, C. R. (2019). Data visualization for inference in tomographic brain imaging.
+
+Pham, D., Muschelli, J., & Mejia, A. (2022). ciftiTools: A package for reading, writing, visualizing, and manipulating CIFTI files in R. *NeuroImage*, 118877.
+
+Poldrack, R. A., Baker, C. I., Durnez, J., Gorgolewski, K. J., Matthews, P. M., Munafò, M. R., ... &
+
+Yarkoni, T. (2017). Scanning the horizon: towards transparent and reproducible neuroimaging research. Nature reviews neuroscience, 18(2), 115-126.
+
+Quintana, D.S. (2020a). *Twitter for Scientists* [eBook edition]. Retrieved from <https://t4scientists.com/>. DOI: 10.5281/zenodo.3707741
+
+Quintana, D. S. (2020b). A synthetic dataset primer for the biobehavioural sciences to promote reproducibility and hypothesis generation. *Elife*, *9*, e53275.
+
 Schäfer, T., & Ecker, C. (2020). fsbrain: an R package for the visualization of structural neuroimaging data. bioRxiv.
 
-Chin, R., Chang, S. W., & Holmes, A. J. (2022). Beyond cortex: The evolution of the human brain. *Psychological Review*.
+Smith, C. N., & Seitz, H. H. (2019). Correcting misinformation about neuroscience via social media. *Science Communication*, *41*(6), 790-819.
+
+Steel, G.
+(2013, September 3).
+Publishing research without data is simply advertising, not science.
+Open Knowledge Foundation.
+<https://blog.okfn.org/2013/09/03/publishing-research-without-data-is-simply-advertising-not-science/>
+
+Sudlow, C., Gallacher, J., Allen, N., Beral, V., Burton, P., Danesh, J., ...
+& Collins, R.
+(2015).
+UK biobank: an open access resource for identifying the causes of a wide range of complex diseases of middle and old age.
+PLoS medicine, 12(3), e1001779.
+
+The Comprehensive R Archive Network. (n.d.). Retrieved April 21, 2022, from <https://cran.r-project.org/>
+
+Van Vliet, M. (2020). Seven quick tips for analysis scripts in neuroimaging. PLoS computational biology, 16(3), e1007358.
+
+Wickham H. (2016). *ggplot2: Elegant Graphics for Data Analysis.* Springer-Verlag New York. ISBN 978-3-319-24277-4, <https://ggplot2.tidyverse.org>.
 

### 2023-02-10 manuscript.Rmd
fig1 shift

@@
 output: 
-  bookdown::word_document2:
-    toc: false
   bookdown::pdf_book:
@@
     latex_engine: lualatex
+  bookdown::word_document2:
+    toc: false
 linestretch: 1.5

### 2023-02-10 manuscript.Rmd
sp fx

--- 2023-02-10 manuscript.Rmd
@@
 These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code, especially within the R, Python and MATLAB environments.
-Here, we first discuss the rational for the wide-spread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
-We then provide a practical guide outlining the steps required to make code-based brain visualizations and provide a web-app (<https://sidchop.shinyapps.io/braincode/>; Figure 4) which can generate simple code-templates as starting points for these visualizationds.
+Here, we first discuss the rationale for the widespread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
+We then provide a practical guide outlining the steps required to make code-based brain visualizations and provide a web-app (<https://sidchop.shinyapps.io/braincode/>; Figure 4) that can generate simple code-templates as starting points for these visualizations.
 We also present a comprehensive table of tools currently available for programmatic brain visualizations (Table 1) and provide didactic examples of visualizations and associated code as a point of reference (Figure 2-3).
-The focus of this guide is on human brain magnetic resonance imaging (MRI) data, but many of the principals discussed and tools provided will equally apply to visualizing data from other organs and imaging modalities such as EEG, MEG, PET and CT.
+The focus of this guide is on human brain magnetic resonance imaging (MRI) data, but many of the principles discussed and tools provided will equally apply to visualizing data from other organs and imaging modalities such as EEG, MEG, PET and CT.
 

### 2023-06-20 manuscript.Rmd
rdme sp fc
revised2

@@
   - name: Elvisha Dhamala
-    department: Department of Psychology
-    affiliation: Yale University
-    location: CT, USA, 06511
+    department: Institute of Behavioral Sciences
+    affiliation: Feinstein Institutes for Medical Research,
+    location: NY, USA 11030
   - name: Edwina R Orchard
-    department: Department of Psychology
+    department: Yale Child Study Center
     affiliation: Yale University
@@
   - name: Avram Holmes
-    department: Department of Psychology
-    affiliation: Yale University
-    location: CT, USA, 06511
+    department: Department of Psychiatry, Brain Health Institute
+    affiliation: Rutgers University
+    location: NJ, USA, 08854
 keywords:
@@
 We then provide a practical guide outlining the steps required to generate these code-based brain visualizations.
-We also present a comprehensive table of tools currently available for programmatic brain visualizations and provide didactic examples of visualizations and associated code as a point of reference.
-Finally, we provide a web-app (https://sidchop.shinyapps.io/braincode/) which can generate simple code-templates as starting points for these visualizations.
+We also present a comprehensive table of tools currently available for programmatic brain visualizations (https://sidchop.shinyapps.io/braincode_selector/) and provide didactic examples of visualizations and associated code as a point of reference.
+Finally, we provide a web-app (https://sidchop.shinyapps.io/braincode/) that can generate simple code-templates as starting points for these visualizations.
 output: 
-  bookdown::word_document2:
-    toc: false
   bookdown::pdf_book:
@@
 - \usepackage[left]{lineno}
+- \usepackage[normalem]{ulem} 
 - \modulolinenumbers[5]
@@
 In large part, the choice to use GUI-based software has been driven by a perception of convenience, flexibility, and accessibility.
-However, there now exist many code-based software packages which are well-documented and often do not require high-level knowledge of programming, making them more accessible to the neuroimaging community.
+However, there now exist code-based software packages that are well-documented and do not require high-level and comprehensive knowledge of programming, making them accessible to the neuroimaging community.
 These tools are flexible and allow for the generation of reproducible, high-quality, and publication-ready brain visualizations in only a few lines of code, especially within the R, Python and MATLAB environments.
 Here, we first discuss the rationale for the widespread adoption of code-generated visualizations by highlighting major advantages in replicability, flexibility, and integration over GUI based tools.
-We then provide a practical guide outlining the steps required to make code-based brain visualizations and provide a web-app (<https://sidchop.shinyapps.io/braincode/>; Figure 4) that can generate simple code-templates as starting points for these visualizations.
-We also present a comprehensive table of tools currently available for programmatic brain visualizations (Table 1) and provide didactic examples of visualizations and associated code as a point of reference (Figure 2-3).
+We then provide a practical guide outlining the steps required to make code-based brain visualizations and provide a web-app (<https://sidchop.shinyapps.io/braincode/>; see Section\@ref(brain-code-a-web-app-for-generating-code-templates-for-brain-visualizations)) that can generate simple code-templates as starting points for these visualizations.
+We also present a comprehensive table of tools currently available for programmatic brain visualizations (Table 1) and provide instructive examples of visualizations and associated code as a point of reference (Figure 2-3).
 The focus of this guide is on human brain magnetic resonance imaging (MRI) data, but many of the principles discussed and tools provided will equally apply to visualizing data from other organs and imaging modalities such as EEG, MEG, PET and CT.
@@
 Notably, some GUI-based tools have historically offered command-line access to generate replicable visualizations (e.g., FreeView, FSLeyes, surfice), making their use potentially equally replicable to purely code-based tools.
-Use of these specialized command-line interfaces can provide a useful middle ground of those who have little experience with coding environments.
-Nonetheless, these interfaces often still have a learning curve, but can lack other advantages, such as iteration, provided by the preferred programming environment (see Sections \@ref(flexibility) and \@ref(integrative-and-interactive-reporting)).
-Likewise, other GUI-based tools offer replicability in the form of automatically generated batch scripts (text files containing lines of specialized commands that can be re-executed) or in-built terminals, which are often idiosyncratic and may lack documentation to make them easily usable or replicable by those not familiar with the specific software.
+Use of these specialized command-line interfaces can provide a useful middle ground for those who have little experience with coding environments.
+Nonetheless, these interfaces often still have a learning curve, but can lack other advantages, such as iteration, provided by the preferred programming environment (see Sections \@ref(flexibility-and-scalability) and \@ref(integrative-and-interactive-reporting)).
+Likewise, other GUI-based tools offer replicability in the form of automatically generated batch scripts (text files containing lines of specialized commands that can be re-executed) or in-built terminals, which can be idiosyncratic and may lack documentation to make them easily usable or replicable by those not familiar with the specific software.
 
-## Flexibility
+## Flexibility and scalability
 
@@
 In particular, the ability to reprogram inputs (such as statistical maps) and settings (such as color schemes, thresholds, and visual orientations) can streamline the entire scientific workflow.
-Changing inputs and settings via code allows for the easy production of multiple figures, such as those resulting from multiple analyses which require similar visualizations.
-A simple for-loop or copying and pasting the code with altered input and/or settings-of-interest can be a powerful method for exploring visualization options or rapidly creating multi-panel figures.
+Changing inputs and settings via code allows for the easy production of multiple figures, such as those resulting from multiple analyses that require similar visualizations.
+A simple for-loop or plotting function with altered input and/or settings-of-interest can be a powerful method for exploring visualization options or rapidly creating multi-panel figures.
 Likewise, an arduous request from a reviewer or collaborator to alter the image processing or analysis becomes less of a burden when the associated figures can be re-generated with a few lines of code, as opposed to re-pasting and re-illustrating them manually.
@@
 Most code-based tools also come with documentation, with R-packages on the *CRAN* or *Neuroconductor* (Muschelli et al., 2019) repositories strictly requiring detailed guidance.
-Recent packages have started to include detailed beginner-friendly documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, & Mejia, 2022; Mowinckel & Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer & Ecker, 2020) which provide examples of figures that can be used as starting points or templates for new users (also see Section\@ref(brain-code-a-web-app-for-generating-code-templates-for-brain-visualizations)).
+Recent packages have started to include detailed beginner-friendly documentation in GitHub repositories, or even entire papers (e.g., Pham, Muschelli, & Mejia, 2022; Mowinckel & Vidal-Piñeiro, 2020; Huntenburg et at., 2017; Schäfer & Ecker, 2020) that provide examples of figures that can be used as starting points or templates for new users (also see Section\@ref(brain-code-a-web-app-for-generating-code-templates-for-brain-visualizations)).
 As the popularity of code-sharing for figure increases, there will be a cornucopia of templates that can be used as the basis for new figures.
@@
 The time it takes to open a single file and achieve the desired visualization settings vastly compounds when working with large datasets.
-Knowing how to programmatically generate brain visualizations can allow for iteration of visualization code over each image of a large datasets making checking the quality of each data processing step achievable.
-The visual outputs of each iteration can be complied into accessible documents which can be easily scrolled, with more advanced usage allowing for the creation of interactive HTML reports (see Section \@ref(generate-and-share-visualization)), similar to those created by standardized data processing tools like *fmriprep* (Esteban et al., 2019).
+Knowing how to programmatically generate brain visualizations can allow for iteration of visualization code over each image of a large datasets making quality checks of each data processing step achievable.
+The visual outputs of each iteration can be complied into accessible documents which can be easily scrolled, with more advanced usage allowing for the creation of interactive HTML reports (see Section \@ref(generate-visualization)), similar to those created by standardized data processing tools like *fmriprep* (Esteban et al., 2019).
 Increasing capacity to conduct visual quality control on larger datasets will improve the identification of processing errors and result in more reliable and valid findings.
@@
 Some journals that publish neuroimaging studies are moving towards allowing the submission of reproducible manuscripts, including reproducible figures (e.g.
-*e-Life*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code linked to the associated article via cloud-based platforms like *Code Ocean* (Code Ocean, 2021).
+*eLife*, *Aperture Neuro*), with other journals like *F1000Research* and *GigaScience* even allowing on-demand re-running of code linked to the associated article via cloud-based platforms like *Code Ocean* (Code Ocean, 2021).
 
@@
 Linking to or even embedding these videos or interactive figures in papers can greatly enhance the communication of findings and make papers more engaging for the reader.
-Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers - such as presentations, websites and social media - all of which can promote the communication of research with peers and reach larger audiences (Li and Xie, 2020).
+Such rich brain visualizations lend themselves to being embedded or shared on science communication mediums beyond academic papers---such as presentations, websites and social media---all of which can promote the communication of research with peers and reach larger audiences (Li and Xie, 2020).
 This last point is becoming increasingly salient as promoting science on social media has become a core medium for spreading discoveries via science communication to the public (Mueller-Herbst, et al., 2020; Smith & Seitz, 2019; Huber et al., 2019), to the research community (Luc et al., 2021; Quintana, 2020a), and even a primary avenue for employment opportunities for early-career researchers (Baker, 2015; Lee, 2019).
@@
 
-## Selecting your programming language
+## Selecting a programming language
 
-The first step (Figure 1) in generating code-based visualization is selecting the coding language you want to use.
+The first step (Figure 1) in generating code-based visualization can be selecting the coding language.
 Three of the most popular languages are R, Python and MATLAB, all of which have many options for generating brain visualizations (see Table 1).
-This decision can be made based on what language you have prior experience with, or if you used one of these languages for the analysis part of your project, there can be advantages in using the same language for visualizations.
-Switching from your analysis environment to separate visualization environment, or a GUI-based visualization software, can be a cumbersome deviation from the scientific workflow.
-This can make debugging errors more difficult, as you must regularly switch program to visually examine the results of any modifications or adjustments to prior analyses.
-Using the brain visualization tools that already exist within your chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analysis.
-Alternatively, the choice of programming language may be dictated by visualization and data-type you require.
+This decision can be made based on what language the user has prior experience with, or if one of these languages was for the analysis part of a project.
+There can be advantages in using the same language for visualizations and analyses, as switching from analysis environment to separate visualization environment, or a GUI-based visualization software, can be a cumbersome deviation from the scientific workflow.
+This can make debugging errors more difficult, as the user must regularly switch program to visually examine the results of any modifications or adjustments to prior analyses.
+Using the brain visualization tools that already exist within a chosen programming environment can provide instant visual feedback on the impact of modifications to processing or analysis.
+Alternatively, the choice of programming language may be dictated by visualization and data-type required.
 For instance, visualizing streamlines from tractography may not be currently available in the R environment (Table 1), and therefore requires the use of Python or MATLAB.
-Other constraints may include access to proprietary software like MATLAB, which would necessitate the use of open-source options such as R or Python.
+Other constraints may include limited access to proprietary software like MATLAB, which would necessitate the use of open-source options such as R, Python or Octave.
 
@@
 
-## Identify your visualization type
+## Identify a visualization type
 
-Neuroimaging data and its derivatives can be visualized in multiple forms, which have different associated file-types and visualization requirements (see Section \@ref(identify-your-input-file-format.)).
+Neuroimaging data and its derivatives can be visualized in multiple forms, that have different associated file-types and visualization requirements (see Section \@ref(identify-input-file-format)).
 A brief description of more popular visualization types is provided below:
 
-***Voxel*****.** A voxel is the 3D equivalent of a pixel.
-In neuroimaging, voxels are used to represent the intensity values of a 3D scan, such as an MRI or CT scan.
+***Voxel*****.** In neuroimaging, voxels are used to represent the intensity values of a 3D scan, such as an MRI or CT scan.
 The voxels can be rendered in different colors to indicate different tissue types or other features of interest.
-For example, in functional MRI (fMRI) scans, voxels can be colored based on their level of activation, to show which areas of the brain are more active during a specific task or for MRI, parts of the brain showing statistically significant anatomical alterations.
+For example, in functional MRI scans, voxels can be colored based on their level of activation, to show which areas of the brain are more active during a specific task.
 These visualizations are often displayed as either slices in axial, sagittal or coronal planes (Fig2A-B; Fig3A-B; or a 3D rendering of the whole brain.
-Statistical values are displayed as overlays on template anatomical images, which follow a common stereotaxic coordinate system (e.g., MNI152), or on individual-specific anatomical images ('native-space').
+Statistical values are displayed as overlays on template anatomical images, which follow a common stereotaxic coordinate system (e.g., MNI152), or on individual-specific anatomical images.
 
-***Vertex.*** A vertex is a point in 3D space that represents a location on the surface.
-In neuroimaging, vertices are used to create a mesh representation of a brain structure, such as the cerebral cortex or a subcortical structure.
+***Vertex.*** In neuroimaging, vertices are used to create a mesh representation of a brain structure, such as the cerebral cortex or a subcortical regions.
 Each vertex has a set of coordinates that specify its location in 3D space and is connected to other vertices to form triangles, which make up the mesh.
@@
 These visualizations are often displayed as 3D rendering of each hemisphere from medial and lateral views (Fig2C,G-H; Fig3C-D).
-Statistical values are displayed as overlays on template surfaces which follow a common stereotaxic coordinate system and fixed number of vertices (e.g., average; Fig2C,G; Fig3C-D), or on individual-specific surfaces that have been reconstructed using an anatomical image (Fig2H).
+Statistical values are displayed as overlays on template surfaces which follow a common stereotaxic coordinate system and fixed number of vertices (e.g., fsaverage; Fig2C,G; Fig3C-D), or on individual-specific surfaces that have been reconstructed using an anatomical image (Fig2H).
 
@@
 Polygonal brain visualizations are simple 2D or 3D shapes which graphically represent the brain or specific structures, but do not carry any additional information on spatial coordinates and only roughly estimate the shape of the brain and its structures.
-Each region of 2D (Fig 2D,F) and 3D (Fig 2E; Fig3E) polygon visualizations can be filled in with colors representing a region label or a statistical value.
+Each region of 2D (Fig 2D,F) and 3D (Fig 2E; Fig3E) polygon visualizations is filled in with colors indicating a region label or a statistical value.
 
-***Edge.*** In the context of neuroimaging, an edge often represents a connection between two brain regions, which can be thought of as a line connecting two nodes (brain regions) in a network.
-To visualize edges, one common method is to use a matrix representation, often called a 'connectivity' or 'adjacency' matrix, which shows the strength of the connections between all pairs of brain regions.
-These matrices can be displayed as a heat map, where different colors represent different levels of connectivity.
-Another way to represent edges is using a 2D graph, where the vertices represent brain regions and the edges represent the connections between those regions, which displayed in a variety of ways, such as a 2D or 3D representation of the brain (Fig2I; Fig3F), with edges represented as lines connecting the vertices.
-The edges and nodes can also be color-coded based on properties such as the strength of the connection.
+***Edge.*** In the context of neuroimaging, an edge often represents a connection between two brain regions, which can be visualised as a straight or curved line connecting two nodes (brain regions) in a network.
+One common way data is organized for edge-level visualization is a matrix, often called a 'connectivity' or 'adjacency' matrix, which indexes the presence and strength of the connections between pairs of brain regions.
+Visualitaion tools often convert these matrices into network graphs, where the vertices represent brain regions and the edges represent the connections between those regions, and can be displayed in a variety of ways, such as overlayed on a 2D or 3D representation of the brain (Fig2I; Fig3F), with edges represented as straight lines connecting the vertices.
+Often, visual properties of edges and nodes can also be adjusted to convey information, such as sized or color-coded based the strength or number of connections.
 
-***Streamlines.*** Streamlines are a representation of the white matter fibers in the brain and are usually generated using tractography applied to diffusion MRI (dMRI).
-Streamlines are typically visualized as 3D lines that connect different points of the brain and can be used to create 3D visualizations of specific white matter tracts, or all streamlines between regions.
+***Streamlines.*** While similar to edges, streamlines are specifically used to representation of the white matter fibers in the brain and are usually an out put of applying tractography applied to diffusion-weighted MRI.
+Streamlines are typically visualized as curved 3D lines that connect different points of the brain and can be used to create 3D visualizations of specific white matter tracts, or all streamlines between regions.
 They can be color-coded based on the direction of the fibers or properties of the tract, such as myelination.
@@
 These are not an exhaustive representation of packages available for visualizing brain data in R and Python (see Table 1).
-Rather, the figures aim to give the reader a sense of the many options available, and an entry point to choosing the type of brain visualization needed (also see Section 4).
+Rather, the figures aim to give the reader a sense of the many options available, and an entry point to choosing the type of brain visualization needed (also see Section\@ref(brain-code-a-web-app-for-generating-code-templates-for-brain-visualizations)).
 All code used to compile the figures, as well as the contents of each panel are provided in the accompanying [online repository](https://github.com/sidchop/RepoNeuroVis).
@@
 ```
-```{r Figure3, echo=FALSE, out.height = '70%', out.width = '90%', fig.align='center', dev='cairo_pdf'}
+```{r Figure3, echo=FALSE, out.height = '60%', out.width = '85%', fig.align='center', dev='cairo_pdf'}
 knitr::include_graphics("./data/python/Figure2.pdf")
@@
 
-## Identify your input file-format
+## Identify input file-formats
 
-Human neuroimaging data and its derivatives come in many different file formats, and code-based visualization packages may have specific requirements about which formats they are designed to work with.
+Neuroimaging data and its derivatives come in many different file formats, and code-based visualization packages have specific formats they are designed to work with.
 Table 1 provides a key, listing which data-type can be used as inputs for each package.
-Some of the most common MRI file-formats used for visualizations are briefly described below:
+Some of the most common MRI file-formats used as inputs for visualizations are briefly described below:
 
@@
 multiple columns or rows), respectively.
-These data are often stored in plain text formats such as `.txt`, `.csv`, and `.tsv` and are often generated by neuroimaging analyses software such as SPM, FSL or FreeSurfer.
+These data are often stored in plain text formats such as `.txt`, `.csv`, and `.tsv` and can be generated by neuroimaging analyses software such as SPM, FSL or FreeSurfer.
 These files can also contain rows and columns of region names and/or spatial coordinates.
-All programming languages have functions to read these plain text formats into your environment as data structures such as matrices and arrays.
-If you have conducted your imaging analyses using your chosen programming language, rather than stand-alone software, the data you want to visualize will likely already be available to you within your coding environment.
-These formats are often used in region-level visualizations, where groups of voxels or vertices share the same value or color (Fig2B-F; Fig3B-E), or in edge-level visualizations, where a matrix is used to identify which two brain regions are connected by an edge (Fig2F; Fig3F).
+All programming languages have functions to read these plain text formats into the coding environment.
+These formats are often used in region-level visualizations, where groups of voxels or vertices share the same value or color (Fig2B-F; Fig3B-E), or in edge-level visualizations, where a matrix is used to identify regions are connected by an edge (Fig2F; Fig3F).
 
 ***NIfTI*****.** NIfTI (`.nii`) files store 3D or 4D image data, which is often a matrix of voxel intensities or a series of 3D matrices for 4D data such as fMRI and dMRI.
-The image data is stored as a 3D matrix of voxel intensities, and the header contains information such as the image dimensions, voxel size, and data type.
-Additional information such as patient demographics and scanner parameters can also be stored in the header in the form of metadata.
+The image data is stored as a 3D matrix of voxel intensities, and the header contains additional information, such as the image dimensions and voxel size.
+Additional information such as the subjects demographics and scanner parameters can also be stored in the header in the form of metadata.
 
-***GIfTI.*** GIFTI (`.gii`) is an extension of the NifTI format, which stores data in a surface-based format, with the data represented as a set of vertices, edges, and faces that define a surface mesh.
+***GIfTI.*** GIFTI (`.gii`) is an extension of the NIFTI format, which stores data in a surface-based format, where data is represented as a set of vertices, edges, and faces that define a surface mesh.
 The format also includes support for storing data such as curvature, thickness, and functional activity maps on the surface mesh.
-Often `.gii` files will have a pre-indicator of what information the file contains, such as `.surf.gii`, which would contain only vertices, edges and faces to define a surface mesh.
+Often `.gii` file names will have a pre-indicator of what information the file contains, such as `.surf.gii`, which would contain only vertices, edges and faces to define a surface mesh.
 Or `.func.gii` files which contain data values for every vertex, which are essentially data arrays whose indices correspond to a surface file and need a matching surface file to know where in the brain to assign the data values.
@@
 
-***CIfTI.*** CIFTI (`.cii`) files can store data from both surface-based and volume-based neuroimaging analyses.
+***CIfTI.*** CIFTI (`.cii`) files can store data from both surface-based and volume-based neuroimaging analyses, combining aspects of NIFTI and GIFTI files.
 For surface-based data, the file contains vertex coordinates, and data values at each vertex.
@@
 CIFTI files can be divided into three main types: `.dtseries` (store time-series data, such as fMRI data), `.dtscalar` (store scalar data, such as thickness or curvature maps), and `.dtlabel` (store label data, such as parcellations of the brain).
-CIFTI files can also include fields with additional information such as surface and volume registration information.
+CIFTI files can also include fields with additional metadata such as surface and volume registration information.
 
 ***FreeSurfer.*** FreeSurfer is a commonly used image processing and analysis software that comes with a variety of proprietary formats to store outputs.
-The primary format is `.mgh` and its compressed version `.mgz`, which like `.nii` files, store voxel-level data.
+The primary format is `.mgh` and its compressed version `.mgz`, which like `.nii` files, genrally stores voxel-level data.
 Vertex-level data is stored in multiple formats such as `.pial`, `.white`, and `.inflated`.
@@
 
-***Tractography.*** Commonly used tractography file formats are `.trk` and `.tck` developed by the TrackVis and MRtrix software packages, respectively.
+***Tractography.*** Commonly used tractography file formats include `.trk` and `.tck` developed by the TrackVis and MRtrix software packages, respectively.
 Both formats store coordinates for streamlines as a series of 3D points, with each point represented by its x, y, and z coordinates, as well as a tract header with the number of points in the tract, the properties of the tract (e.g., mean diffusivity, fractional anisotropy), as well as the starting and ending indices of the tract.
 
-Critically, many of these datatypes are not immutable, and can be converted between each other.
-There are many ways to convert between format, for instance, the Connectome Workbench (Marcus et al., 2011), and corresponding R-packages like `ciftitools` (Pham, Muschelli, & Mejia, 2022) allow for conversion between NifTI, CifTI and GifTI formats.
-FreeSurfer functions such as `mri_convert`, NiBabel (Brett et al., 2023) library in Python and fsbrain (Schäfer & Ecker, 2020) in R, all allow for conversion between the propriety formats and open-source NifTI, CifTI and GifTI formats.
-Therefore, if the file format you have access to is not compatible with the visualization package, library, or toolbox you want to use, you may be able to convert your data into the desired format.
-Although, it is important to visualize and validate data after converting data-structures, such as converting between vectors and matrices, to ensure the data have not been misinterpreted, and also to be aware that unintended consequences of mapping between 2D surface and 3D volume formats may arise (e.g., Ciantar et al., 2022).
+Critically, many of these datatypes are interchangeable, and can be converted between each other.
+There are many ways to convert between format, for instance, the Connectome Workbench (Marcus et al., 2011), and corresponding R-packages like `ciftitools` (Pham, Muschelli, & Mejia, 2022) allow for conversion between NIFTI, CIFTI and GIFTI formats.
+FreeSurfer functions such as `mri_convert`, NiBabel (Brett et al., 2023) library in *Python* and `fsbrain` (Schäfer & Ecker, 2020) in R, all allow for conversion between the propriety formats and open-source NIFTI, CIFTI and GIFTI formats.
+Therefore, if a given file format to is not compatible with a visualization package, library, or toolbox, the user may be able to convert data into the desired format.
+Although, it is important to visualize and validate data after converting between data-structures, to ensure the data have not been misinterpreted, and also to be aware that unintended consequences of mapping between 2D surface and 3D volume formats may arise (e.g., Ciantar et al., 2022).
 
-## Select your package, library or toolbox
+## Select a package, library or toolbox
 
-The previous steps of deciding your programming environment, visualization type and input data-type will help you decide which package, library or toolbox is the right choice for you.
+The previous steps of deciding a programming environment, visualization type and input data-type will help users decide which package, library or toolbox is the right choice.
 Table 1 provides a list of tools in R, Python and MATLAB, characterized by whether they are able to generate voxel, vertex, ROI, edge and streamline based visualizations.
 While each tool may contain the ability to generate code-based visualizations, some tools are more beginner-friendly and better documented than others.
-Usually, this is the case for tools that are specifically designed for brain data visualization, as opposed to tools which are for brain data analysis, but also provide some limited visualization functionality.
+Usually, this is the case for tools that are specifically designed for brain data visualization, as opposed to tools which are designed for brain data analysis but also provide some limited visualization functionality.
 Some examples of well documented and beginner-friendly tools have been provided via a code template generator (see Section \@ref(brain-code-a-web-app-for-generating-code-templates-for-brain-visualizations)).
@@
 Similarly, the `Nilearn` library in python allows plots to be generated using `matplotlib` or `plotly` engines.
-We note that many of the others listed in Table 1 also rely on common plotting engines to generate visualizations.\
+We note that many of the others listed in Table 1 also rely on common plotting engines to generate visualizations.
 
@@
 |   [brainR](https://github.com/muschellij2/brainR)                                                       | ***+^2,6,7^***      |                    | ***+^6,7^***          |                  |                |
-|   [ciftitools](https://github.com/mandymejia/ciftiTools)^**\^**^                                                | ***+^2,3^***        | ***+^2,3^***       | ***+\*^,2,3^***       |                  |                |
-|   [fsbrain](https://github.com/dfsp-spirit/fsbrain)^**\^**^                                                     | ***+^4^***          | ***+^3,4^***       | ***+\*^,6,3,4^***     |                  |                |
-|   [ggseg](https://github.com/ggseg/ggseg)^**\^**^                                                               |                     |                    | ***+^1,4^***          |                  |                |
+|   [ciftitools](https://github.com/mandymejia/ciftiTools)                                                | ***+^2,3^***        | ***+^2,3^***       | ***+\*^,2,3^***       |                  |                |
+|   [fsbrain](https://github.com/dfsp-spirit/fsbrain)                                                     | ***+^4^***          | ***+^3,4^***       | ***+\*^,6,3,4^***     |                  |                |
+|   [ggseg](https://github.com/ggseg/ggseg)                                                               |                     |                    | ***+^1,4^***          |                  |                |
 |   [neurobase](https://github.com/muschellij2/neurobase)                                                 | ***+^2^***          |                    |                       |                  |                |
@@
 |   [Brainplotlib](https://github.com/feilong/brainplotlib)                                               |                     | ***+^1,11^***      | ***+\*^,1,11^***      |                  |                |
-|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)/[surfplot](https://github.com/danjgale/surfplot)^**\^**^ |                     | ***+^3,4,6,7.8^*** | ***+\*^,3,4,6,7,8^*** |                  |                |
-|   [DIPY](https://github.com/dipy/dipy)^**\^**^                                                                  | ***+^2^***          |                    |                       |                  | ***+^5,9^***   |
-|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)^**\^**^                                                  |                     |                    | ***+^1,3,4,6,8^***    |                  |                |
+|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)/[surfplot](https://github.com/danjgale/surfplot) |                     | ***+^3,4,6,7.8^*** | ***+\*^,3,4,6,7,8^*** |                  |                |
+|   [DIPY](https://github.com/dipy/dipy)                                                                  | ***+^2^***          |                    |                       |                  | ***+^5,9^***   |
+|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)                                                  |                     |                    | ***+^1,3,4,6,8^***    |                  |                |
 |   [FSLeyes](https://github.com/pauldmccarthy/fsleyes)                                                   | ***+^2,3,4^***      | ***+^2,3,4^***     | ***+^2,3,4^***        |                  | ***+^2,5^***   |
@@
 |   [netneurotools](https://github.com/netneurolab/netneurotools)                                         |                     | ***+^4^***         | ***+\*^,3^***         |                  |                |
-|   [netplotbrain](https://github.com/wiheto/netplotbrain)^**\^**^                                                |                     |                    | ***+^1,2^***          | ***+^1,11^***    |                |
-|   [nilearn](https://github.com/nilearn/nilearn)^**\^**^                                                         | ***+^2^***          | ***+^2,3^***       | ***+^2,3^***          | ***+^1,11^***    |                |
+|   [netplotbrain](https://github.com/wiheto/netplotbrain)                                                |                     |                    | ***+^1,2^***          | ***+^1,11^***    |                |
+|   [nilearn](https://github.com/nilearn/nilearn)                                                         | ***+^2^***          | ***+^2,3^***       | ***+^2,3^***          | ***+^1,11^***    |                |
 |   [niwidget](https://github.com/nipy/niwidgets)                                                         | ***+^2,3,4^***      | ***+^2,3,4^***     |                       |                  | ***+^5^***     |
 |   [Pycortex](https://github.com/gallantlab/pycortex)                                                    | ***+^8,11,12^***    | ***+^4,8,11,12^*** | ***+\*^,4,8,11,12^*** |                  |                |
-|   [pySurfer](https://pysurfer.github.io/)^**\^**^                                                               |                     | ***+^4^***         | ***+\*^,4^***         |                  |                |
+|   [pySurfer](https://pysurfer.github.io/)                                                               |                     | ***+^4^***         | ***+\*^,4^***         |                  |                |
 |   [surfice](https://github.com/neurolabusc/surf-ice)                                                    | ***+^2,3,4^***      | ***+^2,3,4^***     | ***+^2,3,4,6^***      | ***+^1,4,6^***   | ***+^4,5,6^*** |
@@
 |   [BrainNetViewer](https://www.nitrc.org/projects/bnv/)                                                 | ***+^2,3,4,6,13^*** |                    | ***+^2,3,4,6,13^***   | ***+^1,10^***    |                |
-|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)^**\^**^                                                  |                     | ***+^3,4,6,7,8^*** | ***+\*^,3,4,6,7,8^*** |                  |                |
+|   [Brainspace](https://github.com/MICA-MNI/BrainSpace)                                                  |                     | ***+^3,4,6,7,8^*** | ***+\*^,3,4,6,7,8^*** |                  |                |
 |   [Brainstorm](https://github.com/brainstorm-tools/brainstorm3)                                         | ***+^2^***          | ***+^4,6^***       | ***+^3,4,6^***        |                  |                |
@@
 |   [ECoG/fMRI Vis toolbox](https://github.com/Immiora/matlab-ecog-visualization)                         |                     | ***+^10^***        | ***+\*^,10^***        |                  |                |
-|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)^**\^**^                                                  |                     |                    | ***+^1,10^***         |                  |                |
+|   [ENIGMA TOOLBOX](https://github.com/MICA-MNI/ENIGMA)                                                  |                     |                    | ***+^1,10^***         |                  |                |
 |   [FieldTrip](https://www.fieldtriptoolbox.org/)                                                        | ***+^2^***          | ***+^10^***        |                       |                  |                |
@@
 Note: The tools listed contain functionality required to generate (at least close-to) publication-ready neuroimaging figures via user-entered code within R, MATLAB and Python environments.
+An interactive version of this table can be found here: <https://sidchop.shinyapps.io/braincode_selector/>.
 This list does not include cross-platform general purpose visualization software.
 **1** = .txt/.csv (scalar, vector, matrix as input); **2** = .nii/.nii.gz (nifti as input); **3** = .cii/.gii (cifti or gifti files as input, includes any subtypes e.g.
-dlabel, dtseries, .surf); **4** = FreeSurfer formats as input, including .mgz, .annot, .label, .curv, .wm etc); **5** = .trk/.tck (tractograms as input); **6** = .obj (3D object format); **7** = .ply (3D polygon format); **8** = .vtk (Visualization Toolkit format); **9** = .fib (Legacy vtk format); **10** = .mat (MATLAB format); 11 = .npy/.npz (Python numpy format); **12** = .off (object file format); ***\**** = Cortex only; **\^** = Well-documented, beginner friendly code-based visualization tool (note that this is a subjective criteria intended to guide new-users, and many other tools in this table may also meet this criteria).
+dlabel, dtseries, .surf); **4** = FreeSurfer formats as input, including .mgz, .annot, .label, .curv, .wm etc); **5** = .trk/.tck (tractograms as input); **6** = .obj (3D object format); **7** = .ply (3D polygon format); **8** = .vtk (Visualization Toolkit format); **9** = .fib (Legacy vtk format); **10** = .mat (MATLAB format); 11 = .npy/.npz (Python numpy format); **12** = .off (object file format); ***\**** = Cortex only.
 
-***\**** Cortex only \newpage
+\newpage
 
-## Generate and Share Visualization
+## Generate Visualization
 
-Once the plot has been generated, the visualization can be shared in multiple different ways, with differing levels of replicability (see Section \@ref(replicability)), interactivity (see Section \@ref(integrative-and-interactive-reporting)) and visual quality.
-One common way to share your visualization is to export it as a image raster format such as `.png`, `.tiff` or `.jpeg`.
-These commonly used file formats represent images as a grid of pixels, where each pixel in the grid contains information about the color and intensity of that specific point in the image.
-These formats are resolution dependent, which means that if you try to scale them larger, they will become pixelated and difficult to parse.
-Vector formats, such a `.svg` and `.eps` on the other hand, represent images using mathematical equations, which means that they can be scaled larger or smaller without losing quality.
-While all coding environments provide ways to export visualizations into raster formats, exporting using vector formats, while visually superior, can be dependent on the specific tool you have selected.
-These image files can the be inserted into outputs such as a manuscript and the associated code can be included in an online repository such as [GitHub](https://docs.github.com/en/get-started/quickstart/hello-world) or in supplementary materials associated with the output.
-While this is the simplest way to share visualizations and code, it is only reproducible if the input data, such as NIfTI files used to generate the figure, are also shared.
-In addition to repositories like GitHub, there now exist many platforms to share neuroimaging specific data, such as [OpenNeuro](https://openneuro.org/) and [NeuroVault](https://neurovault.org/).
-If the data used as input for the visualization cannot be shared, synthetic data can be provided (Quintana, 2020) to ensure the reproducibility of the brain visualizations.
-Alternatively, software such as *R Markdown*, *Quarto* and *Jupyter Notebook* allow for the mixing of prose and code in a single script, resulting in fully reproducible and publication ready papers (see Section \@ref(integrative-and-interactive-reporting)).
-While guidance on proper organization of the neuroimaging visualization code is beyond the scope of this guide, we point readers to other practical guides on this topic (Van Vliet, 2020; Gorgolewski & Poldrack, 2016)
+Popular integrated development environments (IDE) such as *RStudio*, *Visual Studio* and *Spyder*, come with the inbuilt ability to display and update figures as the code is executed.
+The resulting visualization can be shared or embedded in papers in multiple different ways, with differing levels of replicability (see Section \@ref(flexibility-and-scalability), interactivity (see Section \@ref(integrative-and-interactive-reporting)) and visual quality.
+One common way to share visualizations is to export it as a image raster format such as `.png`, `.tiff` or `.jpeg`, where images appear as a grid of pixels and each pixel in the grid contains information about the color and intensity of that specific point in the image.
+These formats are resolution dependent, and will become pixelated and difficult to parse when enlarged.
+Vector formats, such a `.svg`, `.eps` and `.pdf` on the other hand can be scaled larger or smaller without losing quality.
+While all coding environments provide ways to export visualizations into raster formats, exporting using vector formats, while visually superior, depends on the specific tool.
+Generally, only visualization tools that rely on mainstream general purpose plotting engines such as `ggplot` in R and `matplotlib` in Python allow for images to be exported as true vector formats.
+
+Increasingly popular software such as *R Markdown*, *Quarto*, *Jupyter Notebook*, and *Google Collab* allow for the creation of dynamic documents that combine code, text, and visualizations in a single file.
+This makes it easier to document workflows and share complete analyses, enhancing both collaboration and reproducibility.
+Including the code used to generate figures and other results alongside well formatted text-based explanations (see Section \@ref(integrative-and-interactive-reporting)), enables the user and others to replicate work accurately.
+These tools also offer a wide range of output formats, including PDF, HTML, Word and LaTeX.
+This versatility enables the generation of polished reports, presentations, manuscripts, or even interactive dashboards, all from a single source file.
+Currently, while *R Markdown* is primarily associated with R and *Google Collab* with Python, Jupyter Notebook and Quarto support a broader range of programming languages, including Python, Julia, R, and others.
+
+## Share visualization code
+
+Images that are generated using code can then be inserted into outputs such as a manuscript and the associated code can be included in supplementary materials.
+While this is the simplest way to share visualizations and code, it may not be as accessible as uploading code to dedicated code-sharing and version control platforms such as [GitHub](https://github.com/) or [GitLab](https://about.gitlab.com/).
+These services are currently widely used and allow for code to maintain formatting, version control and search functions.
+
+Alternate options are widely used general-purpose research data repositories, such as [Open Science Framework](https://osf.io/), [Zenodo](https://zenodo.org/) or [FigShare](https://figshare.com/), that enable researchers to publish and share their datasets, software, code and other research outputs at no cost.
+One advantage of these platforms is that users can assign a Digital Object Identifier (DOI) number to shared materials, making them independently citable and enhancing visibility since DOIs are indexed by various repositories and search engines.
+Some platforms even provide usage metrics, enabling users to gather insight on how often code and materials are accessed, cited or reused by others, which can be valuable information for evaluating the impact and reach of work.
+
+An important consideration when both writing and sharing code for figure generation is the long-term preservation of code and resistance to software collapse.
+While a discussion of code and software preservation is outside the scope of the current paper, readers can refer to Hinsen (2019), and initiatives such as [Software Heritage](https://www.softwareheritage.org/) which aim to preserve and archive all the software source code available worldwide, ensuring that valuable software source code is not lost over time.
+
+In addition to the platform used, how accessible the code is to both the user and others depends on how clearly the code is written, formatted and commented.
+While guidance on proper organization of the neuroimaging visualization code is beyond the scope of this guide, we point readers to other practical guides on this topic (Van Vliet, 2020; Gorgolewski & Poldrack, 2016).
+
+While sharing code is necessary for replicability, it is often not sufficient, as the underlying data source being visualized may be needed for the code to function correctly.
+While many of the same sharing platforms listed above can also host source data alongside code, there now exist specialized platforms, such as [OpenNeuro](https://openneuro.org/) and [NeuroVault](https://neurovault.org/), which to allow sharing neuroimaging specific datasets, such as those containing NIFTI and CIFTI images.
+If the source data for visualizations cannot be shared, synthetic data can also be generated and provided alongside the code be provided (Quintana, 2020).
 
@@
 The code-template also contains links and prompts to more detailed documentation, alternate packages/libraries and tutorials which allow for more complex and publication-ready brain visualizations.
+Users can download bundled version of the web-apps via a [GitHub](https://github.com/sidchop/RepoNeuroVis) repository.
 
@@
 **Figure 4.** Interface for [BrainCode](https://sidchop.shinyapps.io/braincode/) web-app which generates simple code-based templates for brain visualizations.
-(1) Select your programming environment (R or Python).
-(2) Select your visualization type (Voxel, Region, Vertex and Edge).
+(1) Select a programming environment (R or Python).
+(2) Select a visualization type (Voxel, Region, Vertex and Edge).
 There is also a Frequent Asked Questions tab to aid users.
 (3) Manually adjust limited visualization settings and examine how it reactively changes the visualization (4) and code template (5).
-(5) Copy the code template into your programming environment, change file-paths to your own data and explore other visualization settings offered by the functions.
+(5) Copy the code template into selected programming environment, change file-paths to data and explore other visualization settings offered by the functions.
 \newpage
@@
 We hope that the provided template generator gives jumping-off point for new users.
-Nonetheless, this is especially true for the purpose of a publication-ready figure, where fine adjustments to visual features such as legend placement, font size and multi-panel figure positioning may be needed.
+This is especially true for the purpose of a publication-ready figure, where fine adjustments to visual features such as legend placement, font size and multi-panel figure positioning may be needed.
 While most code-based tools offer some control over these finer steps, there are differences between them in feature availability and usability, with tools that use established graphic engines such as `ggplot2`, `matplotlib` and `plotly` providing the most versatile and well-documented features for visual auxiliary.
@@
 Visualizing these structures often requires chaining together multiple GUI-based tools (see Madan, 2015).
-Alternatively users can convert neuroimaging spesific file-formats into domain general visualization or polygon formats, such as `.vtk`, `.ply` or `.obj`, which can be read, manipulated and visualized using general purpose code-based tools.
-Examples of such tools include PyVista and Mayavi in Python and rayshader and plotly in R.
+Alternatively users can convert neuroimaging specific file-formats into domain general visualization or polygon formats, such as `.vtk`, `.ply` or `.obj`, which can be read, manipulated and visualized using general purpose code-based tools.
+Examples of such tools include PyVista and Mayavi in Python and `rayshader` and `plotly` in R.
 Moreover, some neuroimaging derived datatypes, such as streamlines resulting from DWI-based tractography, are still not well represented in code-based visualization tools and future development should focus on enhancing visualization capabilities using these data-types.
 
-As can be seen in Table 1, there are usually multiple packages within each programming environment which can visualize each data type.
+As can be seen in Table 1 (<https://sidchop.shinyapps.io/braincode_selector/>), there are usually multiple packages within each programming environment which can visualize each data type.
 While this provides choice for advanced users, it can also lead to confusion for novice users who may not be familiar with the nuanced differences between tools.
@@
 Gorgolewski, K. J., & Poldrack, R. A. (2016). A practical guide for improving transparency and reproducibility in neuroimaging research. PLoS biology, 14(7), e1002506.
+
+Hinsen, K. (2019). Dealing with software collapse. *Computing in Science & Engineering*, *21*(3), 104-108.
 
